### 2019-04-02 pSplines.tex
Initial commit

\documentclass[aps,reprint,amsmath,amssymb,showpacs,showkeys]{revtex4-1}% APS journal style
\usepackage{graphicx}% Include figure files
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
\usepackage{hyperref}% add hypertext capabilities
\usepackage{natbib}  % Package for Bibtex
%\usepackage{amsmath}  % Package for typesetting formulae
%\usepackage{amssymb}  % Package for some math symbols and fonts
\usepackage[justification=justified]{caption}
%\usepackage[caption=false]{subfig}
\usepackage{subcaption} % subfigure
\usepackage{amsopn}
\usepackage{float}
\usepackage{tikz}
\usetikzlibrary{shapes.geometric, arrows}
\tikzstyle{sqr2} = [rectangle, minimum width=3.5cm, minimum height=0.9cm, text centered, draw=black, fill=yellow!30]

\usepackage{ulem} % \sout function
\usepackage{booktabs} % Top and bottom rules for tables

%\usepackage{breqn}

\newcommand{\bt}[1]{\ensuremath{\mathbf{#1}}}  % bold math
%\newcommand{\bm}[1]{\mbox{\boldmath $#1$}}
\newcommand{\cb}{ \color{blue}}
\newcommand{\cred}{\color{red}}
\newcommand{\jv}[1]{\textcolor{cyan}{JV: #1}}
\newcommand{\Msun}{$\text{M}_{\odot}$}

\begin{document}
\graphicspath{{figures/}}

% Define author, title, and date
\title[Spectral density estimation using P-spline priors]{Spectral density estimation using P-spline priors}

\author{Patricio Maturana-Russel$^{1}$ and Renate Meyer$^1$}
\affiliation{$^1$ Department of
  Statistics, University of Auckland, Auckland 1142, New Zealand}

\begin{abstract}
This article proposes a P-spline prior to estimating the spectral density of a stationary time series.  Our proposal is a particular case of the B-spline prior algorithm.  We assess our proposal via a simulation study and two real data sets.
\end{abstract}

\pacs{04.30.-w, 02.50.-r, 05.45.Tp, 97.60.Bw}

\keywords{P-spline prior, B-spline prior, spectral density estimation, Whittle likelihood}

\maketitle

\section{Introduction}
The power spectral density (PSD), or simply spectral density, describes the distribution of the power or variance for each individual frequency composing a time series.  It embodies useful information for the study of stationary time series.  An absolutely summable autocovariance function, that is $\sum_{h=-\infty}^{\infty} |\gamma(h)| <\infty$, guaranties the existence of the spectral density function of a zero-mean weakly stationary time series, being continuous, bounded and given by
\begin{align*}
f(\lambda) = \dfrac{1}{2\pi} \sum_{h = -\infty}^{\infty}\gamma(h)\exp\left(-i h \lambda \right), 
\end{align*}
where $-\pi < \lambda \leq \pi$ is the angular frequency.

Methods to estimate this function go from parametric to nonparametric approaches.  The former are basically based on autoregressive moving average (ARMA) models.  As any parametric model, the results tend to be biased if the model approximation to the underlying time series is poor.

Most of the nonparametric methods work by smoothing the periodogram
\begin{align*}
I_n(\lambda) = \dfrac{1}{2 \pi n} \left| \sum_{t=1}^{n} Y_t \exp \left( -i t \lambda\right)\right|^2,
\end{align*}
where $-\pi < \lambda \leq \pi$ is the angular frequency and $Y_t$ is a stationary time series of length $n$, because this fluctuates around the true spectral density function.  This function has interesting features, for instance, it is asymptotically unbiased and follows an exponential distribution with mean $f(\lambda)$.  However, it is not consistent \citep{Brockwell:1986}.

Bayesian nonparametric approaches for spectral density estimation allow the data to determine the complexity of the model.  These approaches avoid the comparison of models which vary in complexity and adapt a single model according to the data. \cite{Gangopadhyay:1999} fitted a fixed low order piecewise polynomials to the log-spectral density function, allowing free known number and location of knots.  For this, a reversible jump Markov chain Monte Carlo (RJMCMC) \citep[RJMCMC;][]{Green:1995} algorithm was used.

\cite{Choudhuri:2004} allocated a Bernstein polynomial as prior on the spectral density function.  This prior is a mixture of beta densities weighted by a Dirichlet process.  The number of components, which is a smoothing parameter, has a discrete prior.  \cite{Edwards2018} extended the method replacing the beta densities for B-spline functions.  These functions have the advantage of having local support, unlike the beta densities, which allow more flexibility and have the potential of modelling spectral densities with sharp and abrupt changes.  Thus, the B-spline prior outperforms the Bernstein polynomial in estimating complex PSDs \citep{Edwards2018}.  However, this flexibility comes with a much higher computational cost.

The Bernstein polynomial and the B-spline methods make use of Whittle's approximation to the Gaussian likelihood, known simply as Whittle likelihood \citep{Whittle:1957}.  One of the advantages of this approximation is that involves directly the spectral density, unlike the true likelihood function.

Taking advantage of the limiting independent exponential distribution of the periodogram, the Whittle likelihood for a mean centred weakly stationary time series $\bm{Y}$ of length $n$ is defined as
\begin{align}
 L(\bm{Y} | f) &= \prod_{l=1}^{\nu}\dfrac{1}{f(\lambda_{l})} e^{-I_n(\lambda_l)/f(\lambda_l)}  \nonumber \\
 &\propto \exp \left( - \sum_{l=1}^{\nu} \log f(\lambda_l) + \dfrac{I_{n}(\lambda_l)}{f(\lambda_l)} \right), \label{eq:whittle_like}
\end{align}
where $\lambda_l = 2\pi l / n$ are the Fourier frequencies, $f(\cdot)$ spectral density of $\bm{Y}$, and $\nu = \left\lfloor (n-1)/2 \right\rfloor$ is the greatest integer value less than or equal to $(n-1)/2$.

Motivated by the flexibility of the B-spline prior and its associated high computational cost, in this work, we implement and test a particular case of the B-spline prior.  For this, we fix the number of B-spline densities and place a P-spline prior on the spectral density instead.  This approach allows to penalize automatically the inclusion of B-spline densities and save significantly computational time without sacrificing accuracy.  Actually, as will show, it outperforms the B-spline prior in our tests.  In addition, we propose a knot location scheme based on the periodogram, which aims to deal with PSD shapes that possess  abrupt peaks, without increasing computational time.  We show in an example how this scheme improves significantly the estimates.

We start defining the B-spline densities and the B-spline prior. Then, we define the P-spline prior and discuss its implementation.  Finally, we test our approach and compare it to the B-spline prior in a simulation study, and two real problems.

\section{B-spline prior}

A mixture of beta densities with only integer parameters can approximate uniformly any continuous density on $[0,1]$.  Let $G$ be a cumulative distribution function (cdf), with continuous density $g$, then 
\begin{align*}
\widehat{G}(\omega) &= \sum_{k=1}^{K} G \left( \dfrac{k-1}{K} , \dfrac{k}{K} \right] \beta(\omega; k, K-k+1)\\
&= \sum_{k=1}^{K} w_k \beta(\omega; k, K-k+1)
\end{align*}	  
converges uniformly to $G(\omega)$, where $G(u,v] = G(v) - G(u)$ and $\beta(\omega; a,b)$ is the beta density with parameters $a$ and $b$.
The Bernstein polynomial prior, which is used to describe a nonparametric prior for probability densities in the unit interval, is based on this approximation \citep{Petrone:1999a,Petrone:199b}.  \cite{Choudhuri:2004} proposed it as a prior on the spectral density function.

As shown by \cite{Perron:2001}, the mixture of beta distributions is unable to provide a full coverage of the space of cdfs on $[0,1]$.  They defined a loss function and showed that this cannot be arbitrarily small under this approach.  Alternatively, they showed that if the beta distributions are replaced by B-spline distributions of fixed order (shown for order~2) with variable knots, the loss function can be arbitrarily small by increasing the number of knots.  The adequacy of the approximation is a function of the number of knots.  This work motivated \cite{Edwards2018} to propose the B-spline prior on the spectral density function.

\subsection*{B-splines}

A spline of degree $r+1$ is a function defined piecewise by polynomials of degree $\leq r$, which meet at points called \textit{knots} where the function is continuous.  Any of these spline functions can be described by basis functions, known as \textit{B-splines}, in other words, all spline functions can be represented as a unique combination of B-splines with the same order and over the same partition.  Without loss of generality, the global domain of interest is assumed to be the unit interval $[0,1]$.

The B-spline basic functions of any order can be recursively defined as
\begin{align*}
	B_{k,0}(\omega, \bm{\xi})=&	
	\left\{
	\begin{array}{ll}
		1, & \omega \in [\xi_{k-1}, \xi_{k}] \\
		0, & \mbox{otherwise} 
	\end{array}
	\right.\\
	B_{k,r}(\omega, \bm{\xi}) = & \:v_{k,r}B_{k,r-1}(\omega; \bm{\xi}) + \\
	   & \: (1-v_{k+1,r})B_{k+1,r-1}(\omega; \bm{\xi}),	
\end{align*}
where	   
\begin{align*}	   
	v_{k,r}&=	
	\left\{
	\begin{array}{ll}
		\dfrac{\omega - \xi_{k-1}}{\xi_{k+r-1} - \xi_{k-1}}, & \xi_{k-1} \neq \xi_{k+r-1}\\
		0, & \mbox{otherwise} 
	\end{array}
	\right.
\end{align*}
and 
\begin{align*}
	\bm{\xi} = \{0 = \xi_0 = \xi_1 = \cdots = \xi_r \leq \xi_{r+1} \leq \cdots \leq \xi_{K}\\ = \xi_{K+1} = \cdots = \xi_{K+r} = 1 \}
\end{align*}
is the knot vector.  This vector is a non-decreasing sequence that contains $K+r+1$ knots, which can be divided in $2(r+1)$ external and $K-r-1$ internal knots.  The latter must be $\geq r$.

The \textit{B-spline densities}, normalized B-spline functions, are defined by
\begin{align*}
	b_{k,r}(\omega; \bm{\xi}) = \left(r+1\right) \dfrac{B_{k,r}(\omega;\bm{\xi})}{\xi_{k+r} - \xi_{k-1}}.
\end{align*}

% include reference for the normalizing constant

\subsection*{Prior description}
The B-spline prior is a mixture of B-spline densities defined as
\begin{align}
s_r(\omega; K,\bm{w}_K, \bm{\xi}) = \sum_{k=1}^{K} w_{k,K} b_{k,r}(\omega;\bm{\xi}),
\label{eq:bs_prior}	
\end{align}
where $K$ is the number of B-spline densities  $b_{k,r}(\cdot)$ of fixed degree $r$, $\bm{w}_K=(w_{1,K},\cdots, w_{K,K})$ the weight vector, and $\bm{\xi}$ is the knot sequence.

If the number of B-spline densities $K$ is considered as another parameter, the $\bm{w}_K$ dimension changes according to it.  A natural way to deal with it is the RJMCMC~\citep{Green:1995} algorithm, which was implemented by \cite{Gangopadhyay:1999} in the similar context of low-order polynomials.  However, the method often has implementation difficulties which vary from problem to problem, such as trans-dimensional jumps.

Assuming that the weights are induced by a cdf $G$ on $[0,1]$, \cite{Choudhuri:2004} proposed a Dirichlet process as a prior for the latter.  This approach avoids the potential difficulties caused by the parameter dimension changes associated to RJMCMC.  \cite{Edwards2018} followed this approach, but in addition they assumed that the $K-r$ internal knot differences are induced by a cdf $H$ on $[0,1]$, for which was allocated another Dirichlet prior, that is
\begin{align*}
\Delta_{j} = \xi_{j+r} - \xi_{j+r-1} = H\left(\dfrac{j-1}{K-r}, \dfrac{j}{K-r} \right],
\end{align*}
for $j=1,\dots,K-r$.

Therefore, the B-spline prior on the spectral density function parametrized in terms of $K$, $G$, and $H$ is given by
\begin{align*}
s_r(\omega;K,G,H) = \sum_{k=1}^{K} G \left( \dfrac{k-1}{K} , \dfrac{k}{K} \right] b_{k,r}(\omega;H).
\end{align*}

The B-spline prior is a mixture of B-spline densities with local support, unlike the Bernstein polynomial prior which uses beta densities with full support on the unit interval.  This feature allows the B-spline method more flexibility for spectral density estimation \citep{Edwards2018}.

When there are no internal knots, the B-spline basis turn into the Bernstein polynomial basis.  Therefore, the B-spline prior can be considered as a generalization of the Bernstein polynomial prior.

%\subsection*{Prior for the spectral density}

The prior discussed above has been defined for the unit interval $[0,1]$.  However, the spectral density function $f(\cdot)$ is defined on the interval $[0,\pi]$, therefore its prior is reparametrized and defined as
\begin{align*}
f(\pi \omega) = \tau \times s_r(\omega;K,G,H), \quad \omega \in [0,1],
\end{align*}
where $\tau = \int_{0}^{1}f(\tau \omega)\text{d}\omega$ is the normalizing constant.

To sum up, the prior has the following components:
\begin{itemize}
	\item $G \sim \text{DP}(M_G, G_0)$, where $M_G>0$ is the concentration parameter and $G_0$ is the base distribution.  It produces the weights for the $K$ B-spline densities.  
	\item $H \sim \text{DP}(M_H, H_0)$, where $M_H>0$ is the concentration parameter and $H_0$ is the base distribution.  It produces the knot locations determining the shape and location of the B-spline densities.
	\item $p(K) \propto \text{exp}(-\theta_K K^2)$ for $K=1,\dots,K_{\text{max}}$.  $K$ is the number of B-spline densities used in Eq. \eqref{eq:bs_prior}.  It plays the role of smoothing parameter, that is smaller values implies smoother spectral density estimates.  $K_{\text{max}}$ must be limited  for computational reasons.  In practice, pilot runs are used to determine it.
	\item $\tau \sim \text{Inverse-Gamma}(\alpha_{\tau},\beta_{\tau})$ is the normalizing constant.  $\alpha_{\tau}$ and $\beta_{\tau}$ are the shape and scale parameters, respectively.
\end{itemize}

\subsection*{Implementation discussion}

The B-spline prior proposed by \cite{Edwards2018} involves 2 Dirichlet processes placed on $G$ and $H$, the weights of the B-spline densities and the knot locations, respectively.  The sampling from them is carried out by the stick-breaking construction \citep{Sethuraman:1994}.  This method consists basically in a representation of the Dirichlet process as a sequential division of the unit interval infinite times according to a beta distribution.  For computational reasons, the number of components of this representation of $G$ and $H$ is limited to a large but finite positive integer, that is $L_G$ and $L_H$, respectively.  These parameters control the quality of the approximations.  Larger values improve the approximation, but increase the number of calculations and consequently the computation time.

The stick-breaking approximation requires the re-parametrization of $G$ to $(Z_0,Z_1,\dots, Z_{L_G},V_1,\dots,V_{L_G})$, which are all independent, such that
\begin{align*}
G&=\sum_{l=1}^{L_G} p_l \delta_{Z_l} + \left(1- \sum_{l=1}^{L_G} p_l\right)\delta_{ Z_{0}},
%p1 &= V_1, \\
%p_l&=\left( \prod_{j=1}^{l-1} (1 - V_j) \right) V_l, \quad l \geq 2\\
%p_0&=1-\sum_{l=1}^{L_G}p_l,\\
%V_l &\sim \text{Beta}(1,M_G), \quad l=1,\dots,L_G,\\
%Z_l &\sim G_0, \quad l = 0,1\dots,L_G.
\end{align*}
where $p_1 = V_1$, 
\begin{align*}
p_l &= V_l \prod_{j=1}^{l-1} (1 - V_j), \:\: \text{for } l \geq 2, \\
p_0 &= 1-\sum_{l=1}^{L_G}p_l,
\end{align*}
$\delta_{Z_{0}}$ is a degenerate probability density, that is $\delta_{Z_0}=1$ at $Z_0$ and 0 otherwise, $V_l \sim \text{Beta}(1,M_G)$, for $l=1,\dots,L_G,$ and $Z_l \sim G_0$, for $l = 0,1\dots,L_G$.  Conditional on $K$, the weights of the B-spline densities are defined as
\begin{align*}
	w_{k,K} = \sum_{l=0}^{L_G}p_l I\left\{ \dfrac{k-1}{K} < Z_l < \dfrac{k}{K}\right\}.
\end{align*}

Analogously, $H$ is re-parametrized as $(X_0,\dots,X_{L_H}, U_1,\dots,U_{L_H})$ by the stick-breaking method truncated at $L_H$.  Thus, the knot differences and consequently the knot locations are defined as
\begin{align*}
\Delta_{j} &= \xi_{j+r} - \xi_{j+r-1} \\
&= \sum_{l=0}^{L_H}q_l I\left\{ \dfrac{j-1}{K-r} < X_l < \dfrac{j}{K-r} \right\},		
\end{align*}
where $q_l$ is defined similarly to $p_l$ that was described above.

As a result, the joint prior for the parameter vector $\bm{\theta} = (\textbf{v},\textbf{z},\textbf{u},\textbf{x},K,\tau)$ is given by
\begin{align}
p(\bm{\theta}) \propto & \left( \prod_{l=1}^{L_G} M_G(1-v_l)^{M_G -1} \right) \left( \prod_{l=0}^{L_G}g_0(z_l) \right) \nonumber \\
&\times \left( \prod_{l=1}^{L_H} M_H(1-u_l)^{M_H -1} \right) \left( \prod_{l=0}^{L_H}h_0(x_l) \right) \nonumber \\
&\times p(K) p(\tau) \label{eq:prior}
\end{align}
An unnormalized joint pseudo-posterior is obtained by multiplying this prior density and the Whittle likelihood defined in Eq. \eqref{eq:whittle_like}.

The sampling is performed via the Gibbs sampling technique.  More details about the MCMC implementation of the B-spline prior method can be found in \cite{Edwards2018}.\\

As shown by \cite{Edwards2018}, the B-spline prior outperforms the Bernstein polynomial prior in estimating spectral densities with sharp and abrupt changes. In the case of simple ones, there is no a significant difference.  However, in practice, this function is unknown, which makes necessary the use of general methods, such as the B-spline prior algorithm.

The Bernstein polynomial prior requires the calculation of the beta densities only once and they can be re-utilized across MCMC steps.  On the other hand, the B-spline prior method requires the calculation of the B-spline densities at each iteration, since these vary according to the knot locations and the number of B-spline densities.  Moreover, the Dirichlet processes approximated via the stick-breaking method require $2\!\times\!(L_G + L_H)$ calculations (see Eq. \eqref{eq:prior}) at each iteration.  All of this makes the B-spline prior algorithm much more complex in comparison to the Bernstein polynomial prior.

The complexity of the B-spline prior algorithm is mainly because it allows variable number of B-spline densities $K$.  This has prompted us to the implementation and assessment of the P-spline prior algorithm, which uses a fixed number of B-spline densities and knot locations, avoiding the Dirichlet processes and the recalculation of the B-spline densities.  One of its advantages is that this method keeps the local support of the B-spline densities, unlike the Bernstein polynomials, which can potentially allow to estimate spectral densities with sharp and abrupt changes.

%%%%%%%%%%%%%%%%%
%%% P-splines %%%
%%%%%%%%%%%%%%%%%

\section{P-spline prior}

Fixing the number of B-spline densities $K$ in Eq. \eqref{eq:bs_prior}, the prior for the unit interval $[0,1]$ is defined as
\begin{align*}
	s_r(\omega;\bm{w}, \bm{\xi}) = \sum_{k=1}^{K} w_{k} b_{k,r}(\omega;\bm{\xi}),	
\end{align*}
where $\bm{w} = (w_1,\dots,w_K)$, and $\bm{\xi}$ contains equidistant internal knots on $[0,1]$.  Therefore, the P-spline prior on the spectral density function is given by
\begin{align*}
f(\pi \omega) = \tau \times s_r(\omega;\bm{w}, \bm{\xi}), \quad \omega \in [0,1],
\end{align*}
where $\tau = \int_{0}^{1}f(\tau \omega)\text{d}\omega$ is the normalizing constant.

The roughness penalty in a frequentist context is carried out by penalizing the likelihood function.  In a Bayesian context, this is translated into the prior distribution \citep{Lang:2004} for the $d^{\text{th}}$ order difference of successive weights, known also as B-spline parameters.

In our approach, we place an indirect prior on the weights
\begin{align*}
\bm{v}&\sim\text{N}_{K-1}(\bm{0}, (\phi \bm{P})^{-1})\\
\phi|\delta &\sim \text{Gamma}(\alpha_{\phi}, \delta \beta_{\phi})\\
\delta &\sim \text{Gamma}(\alpha_{\delta}, \beta_{\delta})
\end{align*}
where $\bm{v} = (v_1,\dots,v_{K-1})^\top$ is a $K-1$ dimensional parameter vector, $\phi$ is the smoothing or penalty parameter, $\textbf{P} = \textbf{D}^\top \textbf{D} +\epsilon \textbf{I}_{K-1}$ is the penalty matrix (full matrix rank matrix for a small quantity $\epsilon$, for instance, $10^{-6}$) with $\textbf{D}$ the $d^{\text{th}}$ order difference matrix, $\alpha_{\phi}$ and $\alpha_{\delta}$ are shape parameters, $\delta \beta_{\phi}$ and $\beta_{\delta}$ are rate parameters, and $\text{Gamma}(a,b)$ denotes a gamma distribution with mean $a/b$ and variance $a/b^2$.

The $1^{\text{st}}$ order difference matrix \textbf{D} is defined as
\begin{align*}
\begin{bmatrix}
-1 & 1   & 0  & 0 & \cdots & & & 0 \\
0  &  -1 & 1  & 0 & \cdots & & & 0 \\
\vdots  &    & \ddots & \ddots   & &&& \vdots \\   
0 &    &    &   & \cdots & 0 &  -1 & 1 \\
\end{bmatrix}
\in \mathbb{R}^{(K-2)\times K-1},
\end{align*}
for the $2^{\text{nd}}$ order is
\begin{align*}
\begin{bmatrix}
1 & -2 &  1 & 0 & 0 & \cdots & & & & 0 \\
0 & 1  & -2 & 1 & 0 & \cdots & & & & 0 \\
\vdots & & \ddots & \ddots & \ddots &&&&& \vdots \\
0 & &&&& \cdots & 0 & 1 & -2 & 1 
\end{bmatrix}
\in \mathbb{R}^{(K-3)\times K-1},
\end{align*}
and for the $3^{\text{rd}}$ order is
\begin{align*}
\setcounter{MaxMatrixCols}{12} % to include more than 10 columns in bmatrix enviroment
\begin{bmatrix}
1 & -3 &  3 & -1 &  0 & 0 & \cdots & & & & & 0 \\
0 &  1 & -3 &  3 & -1 & 0 & \cdots & & & & & 0 \\
\vdots &    & \ddots &  \ddots & \ddots & \ddots & &  & & & & \vdots \\
0 &   &  &   &  &  & \cdots & 0 & 1 & -3 & 3 & -1  
\end{bmatrix}
\end{align*}
$\in \mathbb{R}^{(K-4)\times K-1}$.  These are the order penalties used in the application section.

If we define each element of $\bm{v}$ as
\begin{align*}
v_k = \log \left( \dfrac{w_k}{1-\sum_{k=1}^{K-1}w_k} \right),
\end{align*}
after some calculations, it can be shown that the weights are given by
\begin{align*}
w_k = \dfrac{e^{v_k}}{1+ \sum^{K-1}_{k=1}e^{v_k}}.
\end{align*}
The last weight is defined as $w_K = 1 - \sum_{k=1}^{K-1}w_k$, thus they all sum to 1.

We take a robust specification for the penalty prior distribution as suggested by \cite{Jullion:2007}.  Choosing small values for $\alpha_{\delta}$ and $\beta_{\delta}$, for instance, $10^{-4}$, the choice of $\alpha_{\phi}$ and $\beta_{\phi}$ do not affect the spectral density estimate. Here we set $\alpha_{\phi}$ and $\beta_{\phi}$ equal to 1 as used by \cite{Bremhorst:2016}.  These are the prior specifications used in Section \ref{sec:application}.

The parameter vector for the P-spline prior is $\bm{\theta} = (\bm{v}^\top, \phi, \delta, \tau)^\top$ and its joint posterior distribution
\begin{align*}
p(\bm{\theta}|\bm{Y}) & \propto L(\bm{Y}|f) \times p(\bm{\theta})  \\
&= L(\bm{Y}|f) \times p(\bm{v}|\phi, \delta) \times p(\phi|\delta) \times p(\delta) \times p(\tau).
\end{align*}
This posterior is proper since all the prior distributions are proper.  Using the Whittle likelihood defined in \eqref{eq:whittle_like} we obtain a pseudo-posterior.

The conditional posterior of $\phi$, $\delta$, and $\tau$ belong to known families of distributions, that is
\begin{align*}
\phi|\bm{Y},\delta &\sim \text{Gamma}\left(\dfrac{K-1}{2} + \alpha_{\phi},\: \dfrac{1}{2} \bm{v}^\top \textbf{P} \bm{v} + \delta \beta_{\phi} \right), \\
\delta|\bm{Y},\phi &\sim \text{Gamma}\left(\alpha_{\phi} + \alpha_{\delta}, \beta_{\phi} \phi + \beta_{\delta}\right),\:\: \text{and}\\
\tau|\bm{Y} &\sim \text{Inverse-Gamma}\left(\alpha_{\tau}+ \nu,\: \sum_{l=1}^{\nu} \dfrac{I_n(\lambda_l)}{f(\lambda_l)} + \beta_{\tau} \right).
\end{align*}
The posterior distribution of $\tau$ is the same one obtained under the B-spline prior approach.

The influence of the penalty is set by $\phi$.  The larger this value, the smoother the results, whereas the lower the value, the wiggler the results.  It is interesting to note the role of the penalty matrix  \textbf{P} in the rate parameter on the posterior distribution of $\phi$.  When the quadratic form $\bm{v}^\top \textbf{P} \bm{v}$ tends to infinity, the distribution becomes concentrated at $0^{+}$.  This limiting behaviour implies wiggler results.  Note that in this case the prior rate $\delta\beta_{\phi}$ becomes irrelevant in the posterior.  On the other hand, when $\bm{v}^\top \textbf{P} \bm{v}$ tends to $0^{+}$, the distribution favours large $\phi$ values, implying smoother results.

The magnitude of this quadratic form is controlled by the penalty order and the number of B-spline densities.  In general, large order penalties produce lower values for this quantity, causing smoother results.  In this way the penalty matrix penalizes the inclusion of B-spline densities.

\subsection*{Implementation discussion}

The sampling from the posterior distribution for $\phi$, $\delta$, and $\tau$ can be performed directly since its distributions are known as was discussed above.  However, for $\bm{v}$ we require a sampling technique.  For this purpose, we consider the Gibbs sampling algorithm.

In order to improve the mixing of the chains, we apply the Metropolis algorithm on a reparametrized posterior distribution \citep{Lambert:2007}.  For this, we need a pilot posterior sample for $\bm{v}$ from which we calculate its mean vector $\bm{\overline{v}}$ and covariance matrix $\textbf{S}$.  Then we define the following re-parametrization
\begin{align*}
\bm{v} = \textbf{S}^{1/2} \bm{\beta} + \overline{\bm{v}},
\end{align*}
where $\bm{\beta}=(\beta_1, \dots, \beta_{K-1})^\top$.

We can update $\bm{v}$ by modifying $\bm{\beta}$ according to a proposal distribution.  In this work, we propose a univariate proposal value $\beta_{k}^{*}$ to update $\bm{v}$ given by
\begin{align*}
\beta_{k}^{*} &= \beta_{k} + \sigma z,\\
\bm{v}^{*} &= \textbf{S}^{1/2}\bm{\beta}^{*} + \overline{\bm{v}}, 
\end{align*}
where $\bm{\beta}^{*}= (\beta_1, \dots, \beta_{k}^{*} ,\dots, \beta_{K-1})^\top$, $z \sim \text{Normal}(0,1)$, and $\sigma$ controls the length of the proposals.  We vary $\sigma$ across iterations in order to get an acceptance rate between 0.3 and 0.5.  Note that even though the proposal is univariate on $\bm{\beta}$, it is multivariate on $\bm{v}^*$.

The $m^{\text{th}}$ iteration of our approach is given by
\begin{itemize}
	\item Draw $\bm{v}^{m}$ from $p(\bm{v}|\bm{\beta}^{m-1},\phi^{m-1},\delta^{m-1},\tau^{m-1})$ using $K-1$ univariate Metropolis steps for $\beta_k$, with $k=1,\dots,K-1$, according to the reparametrization described above;
	\item Draw $\phi^{m}$ from 
	\begin{align*}
	\text{Gamma}\!\left(\dfrac{K-1}{2}\!+\!\alpha_{\phi},\: \dfrac{1}{2} \bm{v}^{m^\top} \textbf{P} \bm{v}^{m}\!+\!\delta^{m-1} \beta_{\phi} \right)
	\end{align*}
	 in a Gibbs step;
	\item Draw $\delta^{m}$ from $\text{Gamma}\left(\alpha_{\phi} + \alpha_{\delta}, \beta_{\phi} \phi^{m} + \beta_{\delta}\right)$ in a Gibbs step;
	\item Draw $\tau^{m}$ from 
	\begin{align*}
	\text{Inverse-Gamma}\!\left(\alpha_{\tau}\!+\!\nu, \sum_{l=1}^{\nu} \dfrac{I_n(\lambda_l)}{f(\lambda_l)} + \beta_{\tau} \right) 
	\end{align*}
	in a Gibbs step.
\end{itemize}

\cite{Ruppert2002} discussed about the dependence of the amount of smoothness allowed by a particular prior on the scale of the responses $\bm{Y}$.  To avoid the problem, the author suggested to standardize $\bm{Y}$ before the sampling process and apply the inverse operation afterwards.  This also allows to avoid numerical problems in the MCMC process.  We follow his proposal.
%page 187

\subsection*{Choosing the number of B-splines}

The number of B-spline densities plays a critical role in the model fit.  Too few could cause an underfitting whereas too many an overfitting.  Even though the penalty parameter should control the smoothness of the fit, in general, a large number of B-splines produces wiggler results.  On the other hand, a low number produces smoother results.  There is a minimum adequate value which fits the features of the data, but that greater than it the contribution on the fit is not significant.  Unexpectedly, \cite{Ruppert2002} found some cases in which too many B-splines degrades the fitting in terms of mean square error.  Unfortunately, there is no a consensus in how to find the optimal value.

\cite{Eilers2015} consider the use of 100 B-splines a wise choice, unless computational constraints are evident.  Some algorithms that explore different number of knots $K^*$ in a trial sequence and find a relative optimal value according to an arbitrary criteria have been proposed. \cite{Ruppert2002} proposed the full-search and myopic algorithms which are based on the generalized cross-validation statistic.  However, their results are not conclusive and this criterion only applies if the penalty parameter is treated as a tuning constant~\cite{Kauermann2011}.  \cite{Likhachev2017} also considered a knot sequence and proposed the selection of the number of knots based on some statistical information criteria, namely, Akaike, corrected Akaike and Bayesian Information Criterion.

\cite{Ruppert2002} also discussed a heuristic rule of thumb $K^* = \min\left\{n/4, 40\right\}$, which is simpler and, in our experience, seems to work pretty well.  This criterion allocates a knot between $\max\left\{ 4, \lfloor n/40 \rfloor \right\}$ observations, where $\lfloor \cdot \rfloor$ stands for the floor function.  We consider this approach in the application Section for selecting the number of B-spline densities.

% comment: in PSD estimation the knots are alocated in the frequency domain. The knots are not allocated among the observations.  Maybe it should be commented.

\subsection*{Knot location}

The knots are often equally spaced in B-spline methods.  This scheme works quite well in the P-spline prior algorithm for PSDs in which there are no abrupt changes.  However, in the opposite situation, the algorithm will require a large number of knots.  The B-spline prior method is able to handle abrupt changes because the knot location is driven by the data.  We propose a method that allocates the knots according to the periodogram.  The idea is to concentrate more effort in those areas that have potential peaks, detected by the periodogram.

Our knot location proposal works as follow.  We calculate the periodogram and apply a square root transformation in order to have more regular magnitudes and eliminate a potential trend.  Second, we standardize these values and apply the absolute value function.  We treat this transformed periodogram as a density function $f$ and calculate its distribution function $F$.  Finally, the knots are allocated according to the quantiles of the inverse distribution function of $f$.  Thus, in those areas in which the periodogram has sharp and abrupt changes, our procedure will assign more knots in proportion to their magnitudes.

\section{Application}
\label{sec:application}

We assess our P-spline prior approach and compared it to B-spline prior via the simulation study, and the sunspot dataset used by \cite{Edwards2018}.  In addition, we assess the impact of the knot location on the  P-spline estimates using the S. Carinae data, case in which the PSD contains abrupt peaks.  For all the analyses, we use cubic B-splines ($r=3$).

\subsection{Simulation study}

This simulation study was performed by \cite{Edwards2018} in order to compare the B-spline to the Bernstein polynomial priors.  We perform the same analysis, varying only the number of iterations used in the analysis.  The study consists on two autoregressive (AR) time series models of order 1 and 4.  For AR(1), a first autocorrelation order $\rho_1 = 0.9$ is considered, which produces a relatively simple spectral density.  For AR(4), the parameters considered are $\rho_1~=~0.9$, $\rho_2~=~-~0.9$, $\rho_3~=~0.9$ and $\rho_4~=~-~0.9$, which produces a spectral density with two large peaks.  From these models, 300 time series with unit variance Gaussian innovations of length $n = \{128, 256, 512\}$ are generated.

We estimate the spectral density functions using the B-spline and P-spline priors.  The B-spline analysis is performed via the \texttt{gibbs\_bspline} function in the \textsf{R} package \texttt{bsplinePsd} \cite{Edwards:bsplinePsd:2018}.  The algorithm was executed for 100,000 iterations with a burn-in period of 25,000 and thinning factor of 10, resulting in 7,500 samples used for the inferences.  Pilot analyses showed that these specifications are suitable in these examples in terms of convergence (results not shown).

The P-spline analysis is performed using the \texttt{gibbs\_pspline} function implemented in the \textsf{R} package \texttt{psplinePsd}.  Equally spaced knots are assumed.  The algorithm was executed for a total of 100,000 iterations.  A pilot sample of 25,000 iterations with a burn-in period of 5,000 was used to calibrate the proposals.  The final sample consisted in 75,000 samples with a burn-in period of 5,000 and thinning factor of 10, resulting in 7,000 samples used in the inferences.  Therefore, the number of iterations equate the B-spline analysis.

The theoretical spectral density for an AR($p$) model is given by
\begin{align*}
	f(\lambda) = \dfrac{\sigma^2}{2 \pi} \dfrac{1}{|1-\sum_{j=1}^{p}\rho_j \exp(-i \lambda)|^2},
\end{align*}
where $\sigma^2$ is the variance of the white noise innovations and ($\rho_1, \dots, \rho_p$) are the model parameters.  Therefore, the estimates can be compared to the true function.  For this purpose we use the integrated absolute error (IAE) or $L_1$-error, which is a measure of closeness and accuracy and is defined as
\begin{align*}
	\text{IAE} = || \widehat{f} - f ||_1 = \int_{0}^{\pi} |\widehat{f}(\omega)-f(\omega)|\text{d}\omega,
\end{align*}
where $\widehat{f}(\cdot)$ is the Monte Carlo estimate of the spectral density $f(\cdot)$.  We consider the median as spectral density estimate $\widehat{f}(\cdot)$.  This measure is calculated for each of the 300 replications and its median is calculated.  The results are displayed in Table~\ref{table:sim_IAE}.

For the AR(1) cases, the P-spline prior yields more accurate results than the B-spline prior for $n=128$.  Slightly better results are obtained for $n=256$.  For $n=512$, only the first order penalty P-splines yield better results, whereas the second and third order penalty P-splines yield marginally higher median IAE values.

For the AR(4) cases, the P-spline prior yields more accurate results than the B-spline prior for $n=\{128, 256\}$.  For $n=512$, the third order penalty P-splines has a slightly higher median IAE.  Overall, P-spline method produces better results.

As shown by \cite{Edwards2018}, the B-spline prior produces better results than the Bernstein polynomial in terms of median IAE in this example.  Consequently, our P-spline algorithm outperforms the Bernstein method in this context.

% IAE
\begin{table}
	\centering
	\begin{ruledtabular}
	\begin{tabular}{lccc}
		\toprule
		          & $n=128$ & $n=256$ & $n=512$ \\ \hline
		AR(1)     &  &  &  \\ 
		B-spline &  1.257 & 0.738 & 0.605  \\
		P-spline $d=1$ &  0.729 & 0.632 & 0.517  \\
		P-spline $d=2$&  0.844 & 0.612 & 0.638  \\
		P-spline $d=3$&  0.727 & 0.728 & 0.769  \\ \hline
		AR(4)     &  &  &  \\ 
		B-spline & 4.069 & 3.016 & 2.401   \\
		P-spline $d=1$ & 3.014 & 2.462 & 2.335  \\
		P-spline $d=2$& 3.247 & 2.619 & 2.399 \\
		P-spline $d=3$& 3.438 & 2.786 & 2.510  \\		
		\bottomrule
	\end{tabular}
	\end{ruledtabular}
	\caption{Median IAE for the estimated spectral densities using the B-spline and P-spline priors on simulated AR(1) and AR(4) data.  Different penalty order values ($d$) are considered for the P-spline prior algorithm.}
	\label{table:sim_IAE}
\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Coverage probability %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The IAE measures the closeness and accuracy of the tested methods as point estimate, but we could be interested in testing interval estimates.  To assess this estimation, the uniform credible band is considered.  This approach allows to measure the coverage levels for the whole spectral density.  We define the $100(1-\alpha)\%$ uniform credible band as
\begin{align*}
 \widehat{f}(\lambda) \pm \zeta_\alpha \times \text{mad} \left( \widehat{f}_i (\lambda) \right), \:\: \lambda \in [0,\pi],
\end{align*}
where $\widehat{f}(\lambda)$ is the pointwise posterior median spectral density, $\text{mad}( \widehat{f}_i (\lambda))$ is the median absolute deviation of the posterior samples $\widehat{f}_i (\lambda)$ and $\zeta_\alpha$ is such that
\begin{align*}
 \text{P}\left( \max \left\{ \dfrac{|\widehat{f}_i (\lambda) - \widehat{f} (\lambda)|}{\text{mad}(\widehat{f}_i (\lambda))} \right\} \leq \zeta_{\alpha} \right) = 1 - \alpha.
\end{align*}

The idea is to measure the proportion of times that the true spectral density is contained by the uniform credible band.  This calculation is made for all 300 replications.  The results are displayed in Table~\ref{table:sim_covPbb}.

For AR(1) case, both methods cover the whole true spectral density function for all the frequencies $\lambda$.  For AR(4) case there are some variations.  For the first penalty order P-splines, the coverage probabilities are all higher.  For the second order penalty, our proposal has a better performance, except for $n=512$, where the values are equal.  For the third order penalty, we see a mix of results, that is equal, higher and lower values for the increasing sequence of $n$, respectively.  For the P-spline case, in particular for $n=512$, the coverage probability decreases as the penalty degree order increases, which can be explained by the forced smoothness imposed by the higher penalty order degree.  In general, P-spline prior have a better coverage probability than the B-spline method.  \cite{Edwards2018} showed that the Bernstein polynomial method has a coverage probability 0 for $n = \{128,256,512\}$ in the AR(4) case.

% Coverage probability Table
\begin{table}
	\centering
	\begin{ruledtabular}
	\begin{tabular}{lccc}
		\toprule
		          & $n=128$ & $n=256$ & $n=512$ \\ \hline
		AR(1)     &  &  &  \\ 
		B-spline &  1.000 & 1.000 & 1.000 \\
		P-spline $d=1$ & 1.000 & 1.000 & 1.000  \\
		P-spline $d=2$& 1.000 & 1.000 & 1.000 \\
		P-spline $d=3$& 1.000 & 1.000 & 1.000 \\ \hline		       
		AR(4)     &  &  &  \\ 
		B-spline &  0.968 & 0.945 & 0.980  \\
		P-spline $d=1$ & 0.984 & 0.992 & 0.984  \\
		P-spline $d=2$& 0.984 & 0.984 & 0.980 \\
		P-spline $d=3$& 0.968 & 0.976 & 0.969  \\		
		\bottomrule
	\end{tabular}
	\end{ruledtabular}
	\caption{Coverage probabilities based on $90\%$ uniform credible bands.}
	\label{table:sim_covPbb}
\end{table}

%%%%%%%%%%%%
%%% TIME %%%
%%%%%%%%%%%%

As \cite{Edwards2018} noted, one of the B-spline drawbacks is its computational complexity relative to the Bernstein method.  This is directly reflected on the run-time, which is approximately 2-3 times higher in this example (see Table~3 in \cite{Edwards2018}).  The median run-time in our analyses are displayed in Table~\ref{table:sim_time}.  The P-spline method is approximately 4-6 times faster than B-spline approach in these examples.  This is basically because it does not involve as many calculations as the latter per iteration, namely, the B-spline densities are calculated once at the start of the algorithm and there are not Dirichlet process approximations via the stick-breaking method.

% TIME Table
\begin{table}
	\centering
	\begin{ruledtabular}
	\begin{tabular}{lccc}
		\toprule
		          & $n=128$ & $n=256$ & $n=512$ \\ \hline
		AR(1)     &  &  &  \\ 
		B-spline &   30.872 & 34.873 & 42.738 \\
		P-spline $d=1$& 5.506 & 8.185 & 10.338 \\
		P-spline $d=2$& 5.490 & 8.208 & 10.300 \\
		P-spline $d=3$& 5.515 & 8.176 & 10.352 \\ 
		B-spline/P-spline & 5.609 & 4.258 & 4.137 \\ \hline
		AR(4)     &  &  & \\ 
		B-spline & 31.663 & 37.066 & 47.604  \\
		P-spline $d=1$ & 5.538 & 8.410 & 10.443  \\
		P-spline $d=2$&5.501 & 8.187 & 10.402 \\
		P-spline $d=3$& 5.590 & 8.249 & 10.330  \\	
		B-spline/P-spline & 5.712 & 4.475 & 4.581 \\	
		\bottomrule
	\end{tabular}
	\end{ruledtabular}
	\caption{Median run-times in minutes and relative run-times respect to an average of the P-spline times.}
	\label{table:sim_time}
\end{table}

Overall, the P-spline prior outperforms the B-spline prior in this example in terms of IAE, coverage probability and run-time.  Particularly, the first order penalty P-splines outperforms the B-spline method in all our tests.  The criterion which allows to choose automatically the number of B-spline densities in our approach, that is $\min\left\{n/4, 40\right\}$, works perfectly well in these examples.

\subsection{Sunspot data analysis}

Sunspots are darker and cooler regions of the sun's surface.  We analyse the average annual mean sunspot number for the year 1700-1987, that is 288 observations.  This is a classic dataset used to assess power posterior estimation methods.  A square root transformation is applied in order to make the observations more symmetrical and stationary.

As in \cite{Edwards2018}, the B-spline prior MCMC algorithm is run for 100,000 iterations, with a burn-in period of 50,000 and thinning factor of 10.  Thus, 5,000 samples are used for the estimation.  This took around 32 minutes.  The other specifications are the same as those used in the simulation study.

The P-spline prior MCMC algorithm is run in two stages.  The first one is used to calibrate the proposal in the final run.  This is run for 25,000 iterations with a burn-in period of 10,000 and thinning factor of 10, resulting in 1,500 samples.  The final analysis is run for 75,000 iterations with a burn-in period of 25,000 and thinning factor of 10, resulting in 5,000 samples, which is the same number of samples used for the B-spline analysis.  The second order penalty is considered in this example.  The analysis took around 6 minutes.  The rest of specifications are the same as those used in the simulation study.

\begin{figure}[]
	\centering
	\includegraphics[scale=0.5,clip=true,angle=0]{sunspot.pdf}
	\caption{Log-spectral density estimate for the sunspot data. The continuous gray line represents the periodogram.  The continuous and dashed lines stand for the posterior median log-PSD obtained using the B-spline and P-spline priors, respectively.  The dot lines represents the 90\% pointwise credible bands for the latter prior.}
	\label{fig:sunspot}
\end{figure}

The results are displayed in Figure~\ref{fig:sunspot}.  Both methods reveal a large peak at about 0.57. In other words, there is a periodic solar cycle of length $2 \pi / 0.57 \approx 11$ years.  These results are consistent with those obtained via the Bernstein polynomial prior \citep{Choudhuri:2004}.

The 90\% credible bands estimated via the P-spline prior algorithm cover 98\% of the times the posterior median log-PSD estimated via the B-spline prior method.  These results are excellent considering the saving in computational time: B-spline prior method takes approximately 32 minutes, whereas the P-spline prior method only takes around 6 minutes, i.e., approximately 5 times more efficient.

\subsection{Variable star S. Carinae data analysis}

This dataset contains 1189 visual observations of the S. Carinae, a variable star in the southern hemisphere sky.  These are daily observations collected by the Royal Astronomical Society of New Zealand, which correspond to 10 day averages of light intensities over several years.  This data set has been analysed in the literature previously (see for instance \cite{Cart:1997,Huerta:1999,Kirch:2018}).  It contains 40 missing observations that in this work have been replaced by the mean.  This technique does not affect the general features of the data.  In addition, the data has been squared root transformed and mean centred.

\cite{Kirch:2018} provided a PSD estimate via a nonparametrically corrected (NPC) approach, which is able to detect several peaks in the function.  These features makes this data set suitable to assess the impact of the knot location on the P-spline prior estimate. We test 2 knot location approaches: equally spaced internal knots and distributed according to the periodogram as defined above.

The P-spline algorithm is run in two stages.  First, a Markov chain of 5,000 samples with a burn-in period of 1,000 and thinning factor of 10, resulting in 400 points, is used to calibrate the proposals of the final run.  Then, a Markov chain of 10,000 samples with a burn-in period of 2,000 and thinning factor of 10, that is 800 points, is used for the PSD estimation.  We use 40 B-spline densities, number based on the criterion discussed above, and the second order penalty.  The whole process takes less than 2 minutes and is replicated for the 2 knot location approaches.  The results are displayed in Figure~\ref{fig:carinae}.

\begin{figure}[]
	\centering
	\includegraphics[scale=0.5,clip=true,angle=0]{carinae.pdf}
	\caption{Log-spectral density estimate for the variable star S. Carinae data via the P-spline prior algorithm. The continuous gray line represents the periodogram whereas the continuous and dashed black lines stand for the posterior median log-PSD obtained using the periodogram knot location scheme and equally spaced knots, respectively.}
	\label{fig:carinae}
\end{figure}

The periodogram shows several peaks, being the main one located at around 0.42.  When the knots are equally spaced, the algorithm is unable to capture properly this characteristic, unlike our knot location proposal.  This is because our approach allocates more knots in those areas which contain significant peaks detected via the periodogram.  The rest of peaks, which are smaller in comparison to the main one, are slightly better captured by the equally spaced knot approach, since the other scheme spend most of the effort in the main peak, but which can be corrected by increasing the number of knots.  Allowing the knots to vary according to the periodogram, the PSD estimate is more similar to the one obtained via the NPC approach \citep{Kirch:2018}.

\section{Discussion}

As shown by \cite{Edwards2018}, the B-spline prior outperforms the Bernstein polynomial prior in terms of IAE and uniform coverage probabilities for complicated PSDs.  This is explained by the local support of the B-splines and its better approximation properties \citep{Edwards2018}.  However, its performance involves a much higher computational cost.  These characteristics have motivated our approach.

The P-spline prior can be considered a particular case of the B-spline prior.  The latter works with a variable number and location of knots, which are driven by the data.  The P-spline prior assumes a fixed number of knots, with fixed locations.  These simplifications reduce the computational effort significantly, without sacrificing accuracy as shown in the application section.  To deal with abrupt peaks in the PSD, we propose to locate the knots according to the peaks detected via the periodogram.  This approach does not affect the computational time and improve the estimates dramatically.

In our simulation study, we showed that, in general, the P-spline prior outperforms the B-spline prior in all our assessments, that is in terms of IAE, coverage probabilities, and run-times.  From this example, it seems that a second order penalty P-spline works reasonably well.

We also assessed our proposal in the classic sunspot dataset, in which the B-spline prior algorithm is able to estimate correctly the solar cycle that occurs every 11.07 years.  The P-spline prior yields almost identical results, but in a significant less computational time.

Finally, we assess the P-spline prior algorithm under two different knot location schemes: equally spaced and distributed according to the periodogram.  For this, we estimated the PSD for the S. Carinae data, which function contains several abrupt peaks.  The equal spaced scheme fails in capturing the peaks of the function.  On the other hand, our proposal improved significantly the results without affecting the computational time.

The number of B-spline densities was selected according to the criterion proposed by \cite{Ruppert2002}.  It seems to work well, at least for the kind of PSDs treated in this work.  However, more complex functions could require larger number of B-spline densities.  This can be evaluated studying the peaks of the periodogram.  Our knot location proposal is very useful in these situations.

In conclusion, the P-spline prior represents a more viable alternative to the B-spline prior for spectral density estimation.  As shown in this work, it can handle quite well complex PSDs in a reasonable time-run.\\

\begin{acknowledgements}
  We thank the New Zealand eScience Infrastructure (NeSI) for their high performance computing facilities, and the Centre for eResearch at the University of Auckland for their technical support. PM's and RM's work is supported by Grant 3714568 from the University of Auckland Faculty Research Development Fund and the DFG Grant KI 1443/3-1.  All analysis was conducted in \textsf{R}, an open-source statistical software available on \textsf{CRAN} (cran.r-project.org).
\end{acknowledgements}

%\pagebreak

%%%%%%%%%%

\bibliographystyle{apsrev4-1}

\bibliography{pSplinesReferences}

\end{document}

%\begin{thebibliography}{10}

%\end{thebibliography}
%%%%%%%%%%

\end{document}

### 2019-04-02 R calculations/runTimeComparisonSplines.R

@@ -1,0 +1,47 @@
+
+# Table 3 - B-splines paper
+
+x = matrix(c(2.967,3.186,3.659,
+             1.423,1.572,1.844,
+             2.086,2.026,1.985,
+             4.044,4.422,5.174,
+             1.443,1.694,2.281,
+             2.802,2.610,2.268),ncol =3, byrow = TRUE)
+
+rownames(x) = c("Bspline AR(1)", "Bernstein AR(1)", "B-spline/Bernstein",
+                "Bspline AR(4)", "Bernstein AR(4)", "B-spline/Bernstein")
+
+colnames(x) = c("n=128","n=256","n=512")
+
+x
+60*x/4; # in minutes
+
+
+###############################
+### P-spline time estimates ###
+###############################
+
+y = matrix( c(30.872, 34.873, 42.738,
+              5.506, 8.185, 10.338, 
+              5.490, 8.208, 10.300,
+              5.515, 8.176, 10.352, 
+              31.663, 37.066, 47.604,  
+              5.538, 8.410, 10.443,  
+              5.501, 8.187, 10.402, 
+              5.590, 8.249, 10.330), ncol = 3, byrow = TRUE) 	
+
+# AR(1)
+y[1,]/y[2,]; # B-splines / Psplines d=1
+y[1,]/y[3,]; # B-splines / Psplines d=2
+y[1,]/y[4,]; # B-splines / Psplines d=3
+
+round(y[1,]/apply(y[2:4,],2,mean),3)
+
+# AR(4)
+y[5,]/y[6,]; # B-splines / Psplines d=1
+y[5,]/y[7,]; # B-splines / Psplines d=2
+y[5,]/y[8,]; # B-splines / Psplines d=3
+
+round(y[5,]/apply(y[6:8,],2,mean),3)
+
+

### 2019-04-18 pSplines.tex
format change

@@
-\documentclass[aps,reprint,amsmath,amssymb,showpacs,showkeys]{revtex4-1}% APS journal style
-\usepackage{graphicx}% Include figure files
-\usepackage{dcolumn}% Align table columns on decimal point
-\usepackage{bm}% bold math
-\usepackage{hyperref}% add hypertext capabilities
-\usepackage{natbib}  % Package for Bibtex
-%\usepackage{amsmath}  % Package for typesetting formulae
-%\usepackage{amssymb}  % Package for some math symbols and fonts
-\usepackage[justification=justified]{caption}
-%\usepackage[caption=false]{subfig}
-\usepackage{subcaption} % subfigure
-\usepackage{amsopn}
-\usepackage{float}
-\usepackage{tikz}
-\usetikzlibrary{shapes.geometric, arrows}
-\tikzstyle{sqr2} = [rectangle, minimum width=3.5cm, minimum height=0.9cm, text centered, draw=black, fill=yellow!30]
+%\documentclass[11pt,a4paper]{article} % Document specs
+\documentclass[twocolumn,final]{svjour3}
 
-\usepackage{ulem} % \sout function
+% Useful packages
+\usepackage{graphicx}  % Package for importing postscript images
+\usepackage[authoryear, round, numbers]{natbib}  % Package for Bibtex
+\usepackage{amsmath}  % Package for typesetting formulae
+\usepackage{hyperref}  % Package for hyperlinking
+%\usepackage{Sweave}  % Package for Sweave
+\usepackage[top=1.5in, bottom=1.5in, left=1in, right=1in]{geometry}
+%\usepackage{caption}
+\usepackage{amsfonts}
+
 \usepackage{booktabs} % Top and bottom rules for tables
+%\usepackage{lmodern}
 
-%\usepackage{breqn}
+\newcommand{\bm}[1]{\mbox{\boldmath $#1$}}
+\newcommand{\w}{{\bf w}}
+% Define author, title, and date
 
-\newcommand{\bt}[1]{\ensuremath{\mathbf{#1}}}  % bold math
-%\newcommand{\bm}[1]{\mbox{\boldmath $#1$}}
-\newcommand{\cb}{ \color{blue}}
-\newcommand{\cred}{\color{red}}
-\newcommand{\jv}[1]{\textcolor{cyan}{JV: #1}}
-\newcommand{\Msun}{$\text{M}_{\odot}$}
+\title{Spectral density estimation using P-spline priors}
+\author{Patricio Maturana-Russel \and Renate Meyer}
 
-\begin{document}
+\institute{Patricio Maturana-Russel \at Department of Statistics, University of Auckland, Auckland, New Zealand \\ \email{p.russel@auckland.ac.nz} 
+\and Renate Meyer  \at Department of Statistics, University of Auckland, Auckland, New Zealand 
+}
+
+%\date{} % Today's date appears if unspecified
+
 \graphicspath{{figures/}}
 
-% Define author, title, and date
-\title[Spectral density estimation using P-spline priors]{Spectral density estimation using P-spline priors}
-
-\author{Patricio Maturana-Russel$^{1}$ and Renate Meyer$^1$}
-\affiliation{$^1$ Department of
-  Statistics, University of Auckland, Auckland 1142, New Zealand}
-
-\begin{abstract}
-This article proposes a P-spline prior to estimating the spectral density of a stationary time series.
-Our proposal is a particular case of the B-spline prior algorithm.
-We assess our proposal via a simulation study and two real data sets.
-\end{abstract}
-
-\pacs{04.30.-w, 02.50.-r, 05.45.Tp, 97.60.Bw}
-
-\keywords{P-spline prior, B-spline prior, spectral density estimation, Whittle likelihood}
+\begin{document}\sloppy
 
@@
 
+\begin{abstract}
+
+This article proposes a Bayesian approach to estimating the spectral density of a stationary time series using a  P-spline prior.
+Our proposal is motivated by the B-spline Dirichlet process prior of \cite{Edwards2018} combined with Whittle's likelihood with the aim of reducing the computational complexity of posterior computations.
+The strength of the B-spline Dirichlet process prior
+over the Bernstein-Dirichlet process prior of \cite{Choudhuri:2004}  lies in its ability to estimate spectral densities with sharp peaks and abrupt changes due to the flexibility of B-splines with variable number and location of knots.
+Here, we fix the number and location of the knots and specify a P-spline prior.
+We demonstrate in a simulation study and two real  case studies that this approach retains the flexiblity of the B-splines, achieves the same ability
+to accurately estimate peaks due to a new data-driven knot allocation scheme but significantly reduces the computational costs.
+
+\end{abstract}
+
+\keywords{P-spline prior \and B-spline prior \and spectral density estimation \and Whittle likelihood}
+
 \section{Introduction}
-The power spectral density (PSD), or simply spectral density, describes the distribution of the power or variance for each individual frequency composing a time series.
+The power spectral density (psd), or simply spectral density, describes the distribution of the power or variance over the individual frequency components of a time series.
 It embodies useful information for the study of stationary time series.
-An absolutely summable autocovariance function, that is $\sum_{h=-\infty}^{\infty} |\gamma(h)| <\infty$, guaranties the existence of the spectral density function of a zero-mean weakly stationary time series, being continuous, bounded and given by
+For a zero-mean weakly stationary time series, an absolutely summable autocovariance function, i.e.\ $\sum_{h=-\infty}^{\infty} |\gamma(h)| <\infty$, guarantees the existence of a continuous and bounded spectral density function given by
 \begin{align*}
@@
 
-Methods to estimate this function go from parametric to nonparametric approaches.
-The former are basically based on autoregressive moving average (ARMA) models.
-As any parametric model, the results tend to be biased if the model approximation to the underlying time series is poor.
+Methods to estimate this function range from parametric to nonparametric approaches.
+The former are mainly based on fitting  autoregressive moving average (ARMA) models.
+However, the corresponding spectral density will be biased if the ARMA model does not adequately describe the dependence structure of the time series.
+This is a downside of all parametric approaches.
+They are efficient if the model fits well but because they make restrictive assumptions about the data-generating mechanism, their inference is sensitive to model misspecifications and may  be severely biased.
+Bayesian nonparametric approaches, on the other hand, put a prior distribution on the set of all possible densities that could have generated the data not just a small parametric subset thereof.
+They thus avoid the problem of choosing between different parametric models  by an intrinsic data-driven determination of the model complexity.
 
-Most of the nonparametric methods work by smoothing the periodogram
+As the periodogram fluctuates around the true spectral density, most  nonparametric methods for spectral density estimation hinge on  smoothing the periodogram
 \begin{align*}
@@
 \end{align*}
-where $-\pi < \lambda \leq \pi$ is the angular frequency and $Y_t$ is a stationary time series of length $n$, because this fluctuates around the true spectral density function.
-This function has interesting features, for instance, it is asymptotically unbiased and follows an exponential distribution with mean $f(\lambda)$.
+in one way or another, where $-\pi < \lambda \leq \pi$ is the angular frequency and $\bm{Y}=(Y_1,\ldots,Y_n)$ is a stationary time series of length $n$.
+The periodogram has interesting features, for instance, it is asymptotically unbiased and follows an exponential distribution with mean $f(\lambda)$.
 However, it is not consistent \citep{Brockwell:1986}.
 
-Bayesian nonparametric approaches for spectral density estimation allow the data to determine the complexity of the model.
-These approaches avoid the comparison of models which vary in complexity and adapt a single model according to the data.
-\cite{Gangopadhyay:1999} fitted a fixed low order piecewise polynomials to the log-spectral density function, allowing free known number and location of knots.
-For this, a reversible jump Markov chain Monte Carlo (RJMCMC) \citep[RJMCMC;][]{Green:1995} algorithm was used.
-
-\cite{Choudhuri:2004} allocated a Bernstein polynomial as prior on the spectral density function.
-This prior is a mixture of beta densities weighted by a Dirichlet process.
-The number of components, which is a smoothing parameter, has a discrete prior.
-\cite{Edwards2018} extended the method replacing the beta densities for B-spline functions.
-These functions have the advantage of having local support, unlike the beta densities, which allow more flexibility and have the potential of modelling spectral densities with sharp and abrupt changes.
-Thus, the B-spline prior outperforms the Bernstein polynomial in estimating complex PSDs \citep{Edwards2018}.
-However, this flexibility comes with a much higher computational cost.
-
-The Bernstein polynomial and the B-spline methods make use of Whittle's approximation to the Gaussian likelihood, known simply as Whittle likelihood \citep{Whittle:1957}.
-One of the advantages of this approximation is that involves directly the spectral density, unlike the true likelihood function.
-
-Taking advantage of the limiting independent exponential distribution of the periodogram, the Whittle likelihood for a mean centred weakly stationary time series $\bm{Y}$ of length $n$ is defined as
+Most Bayesian approaches to psd estimation use the so-called Whittle likelihood, introduced by \citep{Whittle:1957} as an approximation to the likelihood of a Gaussian stationary time series.
+Based on the  limiting independent exponential distribution of the periodogram, the Whittle likelihood for a mean centered weakly stationary time series $\bm{Y}$ of length $n$ is defined as
 \begin{align}
 L(\bm{Y} | f) &= \prod_{l=1}^{\nu}\dfrac{1}{f(\lambda_{l})} e^{-I_n(\lambda_l)/f(\lambda_l)}  \nonumber \\
- &\propto \exp \left( - \sum_{l=1}^{\nu} \log f(\lambda_l) + \dfrac{I_{n}(\lambda_l)}{f(\lambda_l)} \right), \label{eq:whittle_like}
+&\propto \exp \left( - \sum_{l=1}^{\nu} \log f(\lambda_l) + \dfrac{I_{n}(\lambda_l)}{f(\lambda_l)} \right)\label{eq:whittle_like}
 \end{align}
 where $\lambda_l = 2\pi l / n$ are the Fourier frequencies, $f(\cdot)$ spectral density of $\bm{Y}$, and $\nu = \left\lfloor (n-1)/2 \right\rfloor$ is the greatest integer value less than or equal to $(n-1)/2$.
+Advantages of this approximation is that it depends directly on the spectral density, unlike the true Gaussian likelihood function, and gives a good approximation for non-Gaussian time series under certain conditions as shown by \cite{ShaoXiaofeng2007ASTf}.
 
-Motivated by the flexibility of the B-spline prior and its associated high computational cost, in this work, we implement and test a particular case of the B-spline prior.
-For this, we fix the number of B-spline densities and place a P-spline prior on the spectral density instead.
-This approach allows to penalize automatically the inclusion of B-spline densities and save significantly computational time without sacrificing accuracy.
+Bayesian nonparametric prior specifications for the psd include \cite{Cart:1997}
+who used a smoothing prior on the log psd.
+\cite{Rosen:2007} decompose the log psd into a linear and nonlinear component and  put a Gaussian process prior on the nonlinear part, \cite{Pensky:2007} propose Bayesian wavelet-smoothing of the log psd and recently, and \cite{Cadonna2017} modeled the log psd  as a mixture
+of Gaussian distributions with frequency-dependent weights and mean functions.
+%
+\cite{Gangopadhyay:1999} modelled the log-spectral density function by  piecewise polynomials of low order between strategically placed knots on the support of the psd,  allowing the data to select the number and location of the knots.
+A reversible jump Markov chain Monte Carlo (RJMCMC) \citep[RJMCMC;][]{Green:1995} algorithm was implemented to sample from the posterior distribution.
+
+The P-spline prior on the spectral density function that we are going to propose is related to the 
+Bernstein-Dirichlet process (BDP) prior  introduced by \cite{Choudhuri:2004} and combined with the Whittle likelihood.
+The BDP prior  is a mixture of beta densities with weights induced by a Dirichlet process.
+The number of components, which is a smoothing parameter, is given a discrete prior.
+\cite{Edwards2018} extended the method by replacing the beta densities with B-spline densities.
+Similar to
+\cite{Gangopadhyay:1999}, the specification of the number and location of the B-spline densities is data-driven, however it avoids a RJMCMC algorithm by putting a second Dirichlet process prior  on the cdf that induces the inter-knot distances.
+Unlike the beta densities  which have global support on the interval [0,1], B-spline densities have local support, thereby increasing their flexibility and allowing better modelling of spectral densities with sharp peaks and abrupt changes.
+As demonstrated in \citep{Edwards2018}, the B-spline--Dirichlet process (BspDP) prior outperforms the BDP prior in estimating complex psd's.
+However, this flexibility comes at a high computational price.
+
+%The Bernstein polynomial and the B-spline methods make use of Whittle's approximation to the Gaussian likelihood, known simply as Whittle likelihood \citep{Whittle:1957}.
+
+With the aim of reducing the computational cost of the BspDP prior approach,  we suggest to use the idea of P-splines \cite{Eilers:1996} that combine a large but fixed number of  B-splines with a simple difference penalty that controls the degree smoothness of the spectral density.
+P-splines, i.e.\  equally-spaced B-splines,  have been successfully implemented for frequentist nonparametric regression but to the best of our knowledge not yet for Bayesian spectral density estimation.
+We suggest
+two novel Bayesian  approaches in this paper, both based on the Whittle likelihood and on a mixture of B-spline densities. Both approaches fix the  number of B-spline densities. 
+The first approach uses equally spaced B-spline densities with a P-spline prior which we define in Section III.
+We demonstrate that this approach yields significant savings in computational time without sacrificing accuracy over the BspDP prior approach.
+The second approach has a particular  focus on spectral densities with sharp peaks.
+Instead of an equidistant spacing, the knots are spread based on the distribution of peaks in the periodogram.
+Thus the knot allocation is also data-driven as with the BspDP prior but the knots  remains fixed once
+allocated.
+This novel knot allocation scheme avoids the computational complexity of the BspDP prior.
+We show that with this knot allocation scheme, the estimates of spectral densities with spikes and abrupt changes improves significantly without an increase in computing time.
+
+%Motivated by the flexibility of the B-spline prior and its associated high computational cost, in this work, we implement and test a particular case of the B-spline prior.
+For this, we fix the %number of B-spline densities and place a P-spline prior on the spectral density instead.
+This approach allows to penalize automatically the inclusion of B-spline densities and save significantly %computational time without sacrificing accuracy.
 Actually, as will show, it outperforms the B-spline prior in our tests.
-In addition, we propose a knot location scheme based on the periodogram, which aims to deal with PSD shapes that possess  abrupt peaks, without increasing computational time.
-We show in an example how this scheme improves significantly the estimates.
+In addition, we propose a knot location scheme based on the %periodogram, which aims to deal with psd shapes that possess  abrupt peaks, without increasing computational time.
+We show in an example how this scheme improves significantly the %estimates.
 
-We start defining the B-spline densities and the B-spline prior.
-Then, we define the P-spline prior and discuss its implementation.
-Finally, we test our approach and compare it to the B-spline prior in a simulation study, and two real problems.
+The paper is organized as follows: In order to set notation and allow for a direct comparison, we start by
+reviewing the B-spline densities and the BspDP prior in Section II.
+Section III specifies the P-spline prior for the psd, the novel knot allocations scheme and the Gibbs sampler used to sample from the posterior distribution based on the P-spline prior and Whittle likelihood.
+In Section IV, we test our P-spline-based approaches to spectral density estimation and compare results to those based on the BspDP prior in a simulation study and the application to two real time series.
+The paper concludes with a discussion of the relative merits of the new approaches.
 
-\section{B-spline prior}
+\section{B-spline--Dirichlet process prior}
 
-A mixture of beta densities with only integer parameters can approximate uniformly any continuous density on $[0,1]$.
+By an application of the Weierstrass theorem it can be seen that a mixture of beta densities with only integer parameters can uniformly approximate any continuous density  on the interval $[0,1]$ \citep{Choudhuri:2004}.
 Let $G$ be a cumulative distribution function (cdf), with continuous density $g$, then
@@
 
-As shown by \cite{Perron:2001}, the mixture of beta distributions is unable to provide a full coverage of the space of cdfs on $[0,1]$.
-They defined a loss function and showed that this cannot be arbitrarily small under this approach.
-Alternatively, they showed that if the beta distributions are replaced by B-spline distributions of fixed order (shown for order~2) with variable knots, the loss function can be arbitrarily small by increasing the number of knots.
+As shown by \cite{Perron:2001} for distributions on the interval [0,1], the mixture of B-spline distributions has better approximation properties than the mixture of beta distributions for a fixed number of components.
+In particular, the approximation error of the mixture of B-spline distributions
+can be made arbitrarily small by finding  suitable knot locations and increasing the number of knots.
+%the mixture of beta distributions is unable to provide a full coverage of the space of cdfs on $[0,1]$.
+They defined a loss function and showed that this cannot be arbitrarily small under this %approach.
+Alternatively, they showed that if the beta distributions are replaced by B-spline distributions of fixed order (shown for order~2) with variable knots, the loss function can be %arbitrarily small by increasing the number of knots.
 The adequacy of the approximation is a function of the number of knots.
-This work motivated \cite{Edwards2018} to propose the B-spline prior on the spectral density function.
+This work motivated \cite{Edwards2018} to propose the BspDP prior on the spectral density function.
 
@@
 
-The \textit{B-spline densities}, normalized B-spline functions, are defined by
+The \textit{B-spline densities}, i.e.\ normalized B-spline functions, are defined by
 \begin{align*}
@@
 \subsection*{Prior description}
-The B-spline prior is a mixture of B-spline densities defined as
+The BspDP prior is a mixture of B-spline densities defined as
 \begin{align}
@@
 
-Assuming that the weights are induced by a cdf $G$ on $[0,1]$, \cite{Choudhuri:2004} proposed a Dirichlet process as a prior for the latter.
-This approach avoids the potential difficulties caused by the parameter dimension changes associated to RJMCMC.
-\cite{Edwards2018} followed this approach, but in addition they assumed that the $K-r$ internal knot differences are induced by a cdf $H$ on $[0,1]$, for which was allocated another Dirichlet prior, that is
+Assuming that the weights are induced by a cdf $G$ on $[0,1]$, \cite{Choudhuri:2004} proposed a Dirichlet process as a prior for $G$.
+This approach avoids the potential difficulties caused by  jumps between the parameter spaces of different dimensions in  RJMCMC.
+\cite{Edwards2018} followed this approach, but in addition they assumed that the $K-r$ internal knot differences are induced by a cdf $H$ on $[0,1]$ with further Dirichlet process prior, that is
 \begin{align*}
@@
 
-Therefore, the B-spline prior on the spectral density function parametrized in terms of $K$, $G$, and $H$ is given by
+Therefore, the BspDP prior on the spectral density function parametrized in terms of $K$, $G$, and $H$ is given by
 \begin{align*}
@@
 
-The B-spline prior is a mixture of B-spline densities with local support, unlike the Bernstein polynomial prior which uses beta densities with full support on the unit interval.
+The BspDP prior is a mixture of B-spline densities with local support, unlike the BDP prior which uses beta densities with  support on the whole unit interval.
 This feature allows the B-spline method more flexibility for spectral density estimation \citep{Edwards2018}.
 
-When there are no internal knots, the B-spline basis turn into the Bernstein polynomial basis.  Therefore, the B-spline prior can be considered as a generalization of the Bernstein polynomial prior.
+When there are no internal knots, the B-spline basis coincides with the Bernstein polynomial basis.  Therefore, the BspDP  prior can be considered  a generalization of the BDP prior.
 
@@
 
-The prior discussed above has been defined for the unit interval $[0,1]$.
+The BspDP prior discussed above has been defined for the unit interval $[0,1]$.
 However, the spectral density function $f(\cdot)$ is defined on the interval $[0,\pi]$, therefore its prior is reparametrized and defined as
@@
 
-To sum up, the prior has the following components:
+To sum up, the BspDP prior has the following components:
 \begin{itemize}
@@
 
-\subsection*{Implementation discussion}
+%\subsection*{Implementation discussion}
 
-The B-spline prior proposed by \cite{Edwards2018} involves 2 Dirichlet processes placed on $G$ and $H$, the weights of the B-spline densities and the knot locations, respectively.
-The sampling from them is carried out by the stick-breaking construction \citep{Sethuraman:1994}.
-This method consists basically in a representation of the Dirichlet process as a sequential division of the unit interval infinite times according to a beta distribution.
-For computational reasons, the number of components of this representation of $G$ and $H$ is limited to a large but finite positive integer, that is $L_G$ and $L_H$, respectively.
+The BspDP prior proposed by \cite{Edwards2018} involves two  Dirichlet processes, one  placed on $G$ inducing  the weights of the B-spline densities and the other on  $H$ for the knot locations.
+In practice, these are implemented using   the  stick-breaking representation \citep{Sethuraman:1994}, for details see \cite{Edwards2018}.
+% This method consists basically in a representation of the Dirichlet process as a sequential division of the unit interval infinite times according to a beta distribution. 
+For computational reasons, the number of components of this infinite series representation of $G$ and $H$ is limited to a large but finite positive integer, that is $L_G$ and $L_H$, respectively.
 These parameters control the quality of the approximations.
@@
 
-The stick-breaking approximation requires the re-parametrization of $G$ to $(Z_0,Z_1,\dots, Z_{L_G},V_1,\dots,V_{L_G})$, which are all independent, such that
-\begin{align*}
-G&=\sum_{l=1}^{L_G} p_l \delta_{Z_l} + \left(1- \sum_{l=1}^{L_G} p_l\right)\delta_{ Z_{0}},
-%p1 &= V_1, \\
-%p_l&=\left( \prod_{j=1}^{l-1} (1 - V_j) \right) V_l, \quad l \geq 2\\
-%p_0&=1-\sum_{l=1}^{L_G}p_l,\\
-%V_l &\sim \text{Beta}(1,M_G), \quad l=1,\dots,L_G,\\
-%Z_l &\sim G_0, \quad l = 0,1\dots,L_G.
-\end{align*}
-where $p_1 = V_1$, 
-\begin{align*}
-p_l &= V_l \prod_{j=1}^{l-1} (1 - V_j), \:\: \text{for } l \geq 2, \\
-p_0 &= 1-\sum_{l=1}^{L_G}p_l,
-\end{align*}
-$\delta_{Z_{0}}$ is a degenerate probability density, that is $\delta_{Z_0}=1$ at $Z_0$ and 0 otherwise, $V_l \sim \text{Beta}(1,M_G)$, for $l=1,\dots,L_G,$ and $Z_l \sim G_0$, for $l = 0,1\dots,L_G$.
-Conditional on $K$, the weights of the B-spline densities are defined as
-\begin{align*}
-	w_{k,K} = \sum_{l=0}^{L_G}p_l I\left\{ \dfrac{k-1}{K} < Z_l < \dfrac{k}{K}\right\}.
-\end{align*}
+As shown by \cite{Edwards2018}, the approach based on the  BspDP prior outperforms the BDP prior in estimating spiky spectral densities.  In the case of smooth
+spectral densities, there is no a significant difference in the performance of these two approaches.
+However, in practice, the true psd function is unknown, which makes it necessary to use   methods such as the BspDP that will work well in all conceivable cases.
 
-Analogously, $H$ is re-parametrized as $(X_0,\dots,X_{L_H}, U_1,\dots,U_{L_H})$ by the stick-breaking method truncated at $L_H$.
-Thus, the knot differences and consequently the knot locations are defined as
-\begin{align*}
-\Delta_{j} &= \xi_{j+r} - \xi_{j+r-1} \\
-&= \sum_{l=0}^{L_H}q_l I\left\{ \dfrac{j-1}{K-r} < X_l < \dfrac{j}{K-r} \right\},		
-\end{align*}
-where $q_l$ is defined similarly to $p_l$ that was described above.
+To implement the BDP prior, one only has to  calculate the beta densities once and they can be stored and re-utilized across MCMC steps.
+On the other hand, the BspDP prior requires the calculation of the B-spline densities at each iteration, since these vary due to varying knot locations.
+Moreover,  $2\!\times\!(L_G + L_H)$ calculations are needed in the truncated stick-breaking representation of the two Dirichlet processes  at each iteration %(see Eq.
+\eqref{eq:prior})
+resulting in a significantly increased complexity of  the  BspDP algorithm.
 
-As a result, the joint prior for the parameter vector $\bm{\theta} = (\textbf{v},\textbf{z},\textbf{u},\textbf{x},K,\tau)$ is given by
-\begin{align}
-p(\bm{\theta}) \propto & \left( \prod_{l=1}^{L_G} M_G(1-v_l)^{M_G -1} \right) \left( \prod_{l=0}^{L_G}g_0(z_l) \right) \nonumber \\
-&\times \left( \prod_{l=1}^{L_H} M_H(1-u_l)^{M_H -1} \right) \left( \prod_{l=0}^{L_H}h_0(x_l) \right) \nonumber \\
-&\times p(K) p(\tau) \label{eq:prior}
-\end{align}
-An unnormalized joint pseudo-posterior is obtained by multiplying this prior density and the Whittle likelihood defined in Eq. \eqref{eq:whittle_like}.
-
-The sampling is performed via the Gibbs sampling technique.  More details about the MCMC implementation of the B-spline prior method can be found in \cite{Edwards2018}.\\
-
-As shown by \cite{Edwards2018}, the B-spline prior outperforms the Bernstein polynomial prior in estimating spectral densities with sharp and abrupt changes.
-In the case of simple ones, there is no a significant difference.
-However, in practice, this function is unknown, which makes necessary the use of general methods, such as the B-spline prior algorithm.
-
-The Bernstein polynomial prior requires the calculation of the beta densities only once and they can be re-utilized across MCMC steps.
-On the other hand, the B-spline prior method requires the calculation of the B-spline densities at each iteration, since these vary according to the knot locations and the number of B-spline densities.
-Moreover, the Dirichlet processes approximated via the stick-breaking method require $2\!\times\!(L_G + L_H)$ calculations (see Eq.
-\eqref{eq:prior}) at each iteration.
-All of this makes the B-spline prior algorithm much more complex in comparison to the Bernstein polynomial prior.
-
-The complexity of the B-spline prior algorithm is mainly because it allows variable number of B-spline densities $K$.
-This has prompted us to the implementation and assessment of the P-spline prior algorithm, which uses a fixed number of B-spline densities and knot locations, avoiding the Dirichlet processes and the recalculation of the B-spline densities.
-One of its advantages is that this method keeps the local support of the B-spline densities, unlike the Bernstein polynomials, which can potentially allow to estimate spectral densities with sharp and abrupt changes.
+This has prompted us to explore the performance of two new algorithms, both based on the P-spline prior algorithm.
+Both use a fixed number of B-spline densities and knot locations, avoiding the Dirichlet processes and the recalculation of the B-spline densities.
+Whereas the first approach uses  equidistant knots, the second algorithm  proposes a new fast knot allocation strategy.
+Both methods preserve the flexibility of the B-spline densities and their potential ability to estimate spectral densities with sharp peaks and abrupt changes.
 
@@
 \end{align*}
-where $\bm{w} = (w_1,\dots,w_K)$, and $\bm{\xi}$ contains equidistant internal knots on $[0,1]$.  Therefore, the P-spline prior on the spectral density function is given by
+where $\bm{w} = (w_1,\dots,w_K)$.
+
+$\bm{\xi}$ contains fixed equidistant internal knots on $[0,1]$ in our basic P-spline prior. A different  allotment of knots is described in  section ``knot allocation strategy" .
+Therefore, the P-spline prior on the spectral density function is given by
 \begin{align*}
@@
 
-The roughness penalty in a frequentist context is carried out by penalizing the likelihood function.
-In a Bayesian context, this is translated into the prior distribution \citep{Lang:2004} for the $d^{\text{th}}$ order difference of successive weights, known also as B-spline parameters.
+Smoothing splines in the frequentist context are based on approximating a given function by a linear combination of B-splines and
+putting a penalty on the integrated squared second derivative  as a measure of roughness.
+P-splines avoid derivatives by expressing the penalty as the sum of squares of differences of the B-psline coefficients.
+%The roughness penalty in a frequentist context is carried out by penalizing the likelihood function.  
+In a Bayesian context, this penalty can be translated into a prior distribution for the $d^{\text{th}}$ order difference of successive coefficients, see e.g. \citep{Lang:2004}.
+%, known also as B-spline parameters.
 
-In our approach, we place an indirect prior on the weights
+In our approach, the B-spline coefficients $w_k$ are weights, so positive and sum to one. 
+We therefore reparametrize to the vector $\bm{v}$ with
+\begin{align*}
+v_k = \log \left( \dfrac{w_k}{1-\sum_{k=1}^{K-1}w_k} \right).
+\end{align*}
+After some calculations, it can be shown that the weights are given by
+\begin{align*}
+w_k = \dfrac{e^{v_k}}{1+ \sum^{K-1}_{k=1}e^{v_k}}.
+\end{align*}
+The last weight is defined as $w_K = 1 - \sum_{k=1}^{K-1}w_k$, thus they all sum to 1.
+Then, an indirect prior is placed on the weights via
 \begin{align*}
@@
 \end{align*}
-where $\bm{v} = (v_1,\dots,v_{K-1})^\top$ is a $K-1$ dimensional parameter vector, $\phi$ is the smoothing or penalty parameter, $\textbf{P} = \textbf{D}^\top \textbf{D} +\epsilon \textbf{I}_{K-1}$ is the penalty matrix (full matrix rank matrix for a small quantity $\epsilon$, for instance, $10^{-6}$) with $\textbf{D}$ the $d^{\text{th}}$ order difference matrix, $\alpha_{\phi}$ and $\alpha_{\delta}$ are shape parameters, $\delta \beta_{\phi}$ and $\beta_{\delta}$ are rate parameters, and $\text{Gamma}(a,b)$ denotes a gamma distribution with mean $a/b$ and variance $a/b^2$.
-
+where $\bm{v} = (v_1,\dots,v_{K-1})^\top$ is a $K-1$ dimensional parameter vector, $\phi$ is the smoothing or penalty parameter, $\textbf{P} = \textbf{D}^\top \textbf{D} +\epsilon \textbf{I}_{K-1}$ is the penalty matrix (which is a full matrix rank matrix for any small quantity $\epsilon$, for instance, $10^{-6}$) with $\textbf{D}$ the $d^{\text{th}}$ order difference matrix ($d=1,2,3,\ldots)$, $\alpha_{\phi}$ and $\alpha_{\delta}$ are shape parameters, $\delta \beta_{\phi}$ and $\beta_{\delta}$ are rate parameters, and $\text{Gamma}(a,b)$ denotes a gamma distribution with mean $a/b$ and variance $a/b^2$.
+We used first, second and third order difference penalties in the simulation and case study.
 The $1^{\text{st}}$ order difference matrix \textbf{D} is defined as
@@
 \end{bmatrix}
-\in \mathbb{R}^{(K-2)\times K-1},
+\in \mathbb{R}^{(K-2)\times (K-1)},
 \end{align*}
@@
 \end{bmatrix}
-\in \mathbb{R}^{(K-3)\times K-1},
+\in \mathbb{R}^{(K-3)\times (K-1)},
 \end{align*}
@@
 \end{align*}
-$\in \mathbb{R}^{(K-4)\times K-1}$.  These are the order penalties used in the application section.
-
-If we define each element of $\bm{v}$ as
-\begin{align*}
-v_k = \log \left( \dfrac{w_k}{1-\sum_{k=1}^{K-1}w_k} \right),
-\end{align*}
-after some calculations, it can be shown that the weights are given by
-\begin{align*}
-w_k = \dfrac{e^{v_k}}{1+ \sum^{K-1}_{k=1}e^{v_k}}.
-\end{align*}
-The last weight is defined as $w_K = 1 - \sum_{k=1}^{K-1}w_k$, thus they all sum to 1.
+$\in \mathbb{R}^{(K-4)\times (K-1)}$.
 
@@
 
-The parameter vector for the P-spline prior is $\bm{\theta} = (\bm{v}^\top, \phi, \delta, \tau)^\top$ and its joint posterior distribution
+The parameter vector for the P-spline prior is $\bm{\theta} = (\bm{v}^\top, \phi, \delta, \tau)^\top$.
+
+\subsection*{Posterior Computation}
+
+The  joint posterior distribution of the parameter vector $\bm{\theta} = (\bm{v}^\top, \phi, \delta, \tau)^\top$ is given by
 \begin{align*}
 p(\bm{\theta}|\bm{Y}) & \propto L(\bm{Y}|f) \times p(\bm{\theta})  \\
-&= L(\bm{Y}|f) \times p(\bm{v}|\phi, \delta) \times p(\phi|\delta) \times p(\delta) \times p(\tau).
+&= L(\bm{Y}|f) \times p(\bm{v}|\phi, \delta) \times p(\phi|\delta) \times p(\delta) \times p(\tau)
 \end{align*}
-This posterior is proper since all the prior distributions are proper.  Using the Whittle likelihood defined in \eqref{eq:whittle_like} we obtain a pseudo-posterior.
+where $ L(\bm{Y}|f) $ denotes the Whittle likelihood defined in \eqref{eq:whittle_like}.
+This posterior is proper since all the prior distributions are proper.  
+We use the Gibbs sampler to sample from the joint posterior distribution.
 
-The conditional posterior of $\phi$, $\delta$, and $\tau$ belong to known families of distributions, that is
+Sampling from the full conditional posterior distributions of $\phi$, $\delta$, and $\tau$ can be performed directly. These are given by
+%The conditional posterior of $\phi$, $\delta$, and $\tau$ belong to known families of distributions, that is
 \begin{align*}
-\phi|\bm{Y},\delta &\sim \text{Gamma}\left(\dfrac{K-1}{2} + \alpha_{\phi},\: \dfrac{1}{2} \bm{v}^\top \textbf{P} \bm{v} + \delta \beta_{\phi} \right), \\
+\phi|\bm{Y},\delta,\bm{v} &\sim \text{Gamma}\left(\tfrac{K-1}{2} + \alpha_{\phi},\: \tfrac{1}{2} \bm{v}^\top \textbf{P} \bm{v} + \delta \beta_{\phi} \right), \\
 \delta|\bm{Y},\phi  &\sim \text{Gamma}\left(\alpha_{\phi} + \alpha_{\delta}, \beta_{\phi} \phi + \beta_{\delta}\right),\:\: \text{and}\\
-\tau|\bm{Y} &\sim \text{Inverse-Gamma}\left(\alpha_{\tau}+ \nu,\: \sum_{l=1}^{\nu} \dfrac{I_n(\lambda_l)}{f(\lambda_l)} + \beta_{\tau} \right).
+\tau|\bm{Y},\delta, \phi,\bm{v}   &\sim \text{Inv-Gamma}\!\left(\alpha_{\tau}+ \nu,\:  {\textstyle\sum\limits_{l=1}^{\nu}} \tfrac{I_n(\lambda_l)}{f(\lambda_l)} + \beta_{\tau} \right)\!.
 \end{align*}
-The posterior distribution of $\tau$ is the same one obtained under the B-spline prior approach.
+%The posterior distribution of $\tau$ is the same one obtained under the B-spline prior approach.
 
-The influence of the penalty is set by $\phi$.
-The larger this value, the smoother the results, whereas the lower the value, the wiggler the results.
-It is interesting to note the role of the penalty matrix  \textbf{P} in the rate parameter on the posterior distribution of $\phi$.
-When the quadratic form $\bm{v}^\top \textbf{P} \bm{v}$ tends to infinity, the distribution becomes concentrated at $0^{+}$.
-This limiting behaviour implies wiggler results.
-Note that in this case the prior rate $\delta\beta_{\phi}$ becomes irrelevant in the posterior.
-On the other hand, when $\bm{v}^\top \textbf{P} \bm{v}$ tends to $0^{+}$, the distribution favours large $\phi$ values, implying smoother results.
-
-The magnitude of this quadratic form is controlled by the penalty order and the number of B-spline densities.
-In general, large order penalties produce lower values for this quantity, causing smoother results.
-In this way the penalty matrix penalizes the inclusion of B-spline densities.
-
-\subsection*{Implementation discussion}
-
-The sampling from the posterior distribution for $\phi$, $\delta$, and $\tau$ can be performed directly since its distributions are known as was discussed above.
-However, for $\bm{v}$ we require a sampling technique.
-For this purpose, we consider the Gibbs sampling algorithm.
-
+However, to sample from the full conditional distribution of $\bm{v}$, we use the Metropolis algorithm.
 In order to improve the mixing of the chains, we apply the Metropolis algorithm on a reparametrized posterior distribution \citep{Lambert:2007}.
@@
 
-The $m^{\text{th}}$ iteration of our approach is given by
+The $m^{\text{th}}$ cycle of the Gibbs sampler is given by
 \begin{itemize}
@@
 	\begin{align*}
-	\text{Gamma}\!\left(\dfrac{K-1}{2}\!+\!\alpha_{\phi},\: \dfrac{1}{2} \bm{v}^{m^\top} \textbf{P} \bm{v}^{m}\!+\!\delta^{m-1} \beta_{\phi} \right)
+	\text{Gamma}\!\left(\tfrac{K-1}{2}\!+\!\alpha_{\phi},\: \tfrac{1}{2} \bm{v}^{m^\top} \textbf{P} \bm{v}^{m}\!+\!\delta^{m-1} \beta_{\phi} \right);
 	\end{align*}
-	 in a Gibbs step;
-	\item Draw $\delta^{m}$ from $\text{Gamma}\left(\alpha_{\phi} + \alpha_{\delta}, \beta_{\phi} \phi^{m} + \beta_{\delta}\right)$ in a Gibbs step;
+
+	\item Draw $\delta^{m}$ from $\text{Gamma}\left(\alpha_{\phi} + \alpha_{\delta}, \beta_{\phi} \phi^{m} + \beta_{\delta}\right);$ 
 	\item Draw $\tau^{m}$ from 
 	\begin{align*}
-	\text{Inverse-Gamma}\!\left(\alpha_{\tau}\!+\!\nu, \sum_{l=1}^{\nu} \dfrac{I_n(\lambda_l)}{f(\lambda_l)} + \beta_{\tau} \right) 
+	\text{Inv-Gamma}\!\left(\alpha_{\tau}\!+\!\nu, {\textstyle\sum\limits_{l=1}^{\nu}} \tfrac{I_n(\lambda_l)}{f(\lambda_l)} + \beta_{\tau} \right) .
 	\end{align*}
-	in a Gibbs step.
 \end{itemize}
 
-\cite{Ruppert2002} discussed about the dependence of the amount of smoothness allowed by a particular prior on the scale of the responses $\bm{Y}$.
+The influence of the penalty is determined by $\phi$.  The larger this value is, the smoother is the resulting estimate. %, whereas the lower the value, the rougher the results.
+It is interesting to note the role of the penalty matrix  \textbf{P} in the rate parameter on the full conditional posterior distribution of $\phi$.
+When the quadratic form $\bm{v}^\top \textbf{P} \bm{v}$ tends to infinity, this distribution becomes concentrated at $0^{+}$.
+This limiting behaviour yields rougher results.
+Note that in this case the prior rate $\delta\beta_{\phi}$ becomes irrelevant in the full conditional posterior density.
+On the other hand, when $\bm{v}^\top \textbf{P} \bm{v}$ tends to $0^{+}$, the distribution favours large $\phi$ values, yielding smoother estimates.
+%
+The magnitude of this quadratic form is controlled by the penalty order and the number of B-spline densities.
+In general, large order penalties produce lower values for this quantity, causing smoother results.
+In this way the penalty matrix penalizes the inclusion of B-spline densities.
+
+\cite{Ruppert2002} explained that the degree of smoothness implied by a certain choice of prior parameters also depends on the scale of the responses $\bm{Y}$.
 To avoid the problem, the author suggested to standardize $\bm{Y}$ before the sampling process and apply the inverse operation afterwards.
@@
 
-The number of B-spline densities plays a critical role in the model fit.
-Too few could cause an underfitting whereas too many an overfitting.
-Even though the penalty parameter should control the smoothness of the fit, in general, a large number of B-splines produces wiggler results.
-On the other hand, a low number produces smoother results.
-There is a minimum adequate value which fits the features of the data, but that greater than it the contribution on the fit is not significant.
+The number of B-spline densities plays a critical role in the model fit and may result in under- or oversmoothing.
+%Too few could cause an underfitting whereas too many an overfitting.  
+Even though the penalty parameter controls the smoothness of the fit, a large number of B-splines produces rougher estimates and  a low number produces smoother spectral densities.
+There is a smallest sufficient number which results in a psd estimate that fits the features of the data whereas choosing a larger number will have an insignificant effect on the fit.
 Unexpectedly, \cite{Ruppert2002} found some cases in which too many B-splines degrades the fitting in terms of mean square error.
-Unfortunately, there is no a consensus in how to find the optimal value.
+Unfortunately, there is no  consensus in how to find the optimal value.
 
-\cite{Eilers2015} consider the use of 100 B-splines a wise choice, unless computational constraints are evident.
-Some algorithms that explore different number of knots $K^*$ in a trial sequence and find a relative optimal value according to an arbitrary criteria have been proposed.
-\cite{Ruppert2002} proposed the full-search and myopic algorithms which are based on the generalized cross-validation statistic.
+\cite{Eilers2015} consider the use of $K=100$ B-splines a wise choice, unless computational constraints are evident.
+Some algorithms that explore different number of knots $K^*$ in a trial sequence and find a relative optimal value according to certain goodness-of-fit criteria have been proposed.
+\cite{Ruppert2002} suggested a full-search and myopic algorithms which are based on the generalized cross-validation statistic.
 However, their results are not conclusive and this criterion only applies if the penalty parameter is treated as a tuning constant~\cite{Kauermann2011}.
-\cite{Likhachev2017} also considered a knot sequence and proposed the selection of the number of knots based on some statistical information criteria, namely, Akaike, corrected Akaike and Bayesian Information Criterion.
+\cite{Likhachev2017} also considered a knot sequence and proposed the selection of the number of knots based on  Akaike's, corrected Akaike's and
+the Bayesian Information Criterion.
 
-\cite{Ruppert2002} also discussed a heuristic rule of thumb $K^* = \min\left\{n/4, 40\right\}$, which is simpler and, in our experience, seems to work pretty well.
+\cite{Ruppert2002} discussed a heuristic rule of thumb $K^* = \min\left\{n/4, 40\right\}$, which is simpler and, in our experience, seems to work very well in general.
 This criterion allocates a knot between $\max\left\{ 4, \lfloor n/40 \rfloor \right\}$ observations, where $\lfloor \cdot \rfloor$ stands for the floor function.
-We consider this approach in the application Section for selecting the number of B-spline densities.
+We employ this approach in the application Section for selecting the number of B-spline densities.
 
-% comment: in PSD estimation the knots are alocated in the frequency domain. The knots are not allocated among the observations.  Maybe it should be commented.
+% comment: in psd estimation the knots are alocated in the frequency domain. The knots are not allocated among the observations.  Maybe it should be commented.
 
-\subsection*{Knot location}
+\subsection*{Knot allocation strategy}
 
-The knots are often equally spaced in B-spline methods.
-This scheme works quite well in the P-spline prior algorithm for PSDs in which there are no abrupt changes.
-However, in the opposite situation, the algorithm will require a large number of knots.
-The B-spline prior method is able to handle abrupt changes because the knot location is driven by the data.
-We propose a method that allocates the knots according to the periodogram.
-The idea is to concentrate more effort in those areas that have potential peaks, detected by the periodogram.
+The knots are often equally spaced in P-spline methods.
+This scheme works quite well in the P-spline prior algorithm for psds in which there are no abrupt changes.
+However, for spectral densities with spikes, the algorithm will require a large number of knots to adequately estimate the peaks.
+The B-spline prior method is able to handle abrupt changes because the knot location is variable and  driven by the data.
+We propose a method that allocates the knots according to the periodogram but once allocated, the knots stay fixed.
+The idea is to concentrate more effort in those regions that have potential peaks as detected by the periodogram.
 
@@
 Second, we standardize these values and apply the absolute value function.
-We treat this transformed periodogram as a density function $f$ and calculate its distribution function $F$.
-Finally, the knots are allocated according to the quantiles of the inverse distribution function of $f$.
+We treat this transformed periodogram as a density function $\widetilde{f}$ and calculate its distribution function $F$.
+Finally, the knots are allocated according to the quantiles of the inverse distribution function of $\widetilde{f}$.
 Thus, in those areas in which the periodogram has sharp and abrupt changes, our procedure will assign more knots in proportion to their magnitudes.
+Hereafter, we will refer to this scheme as \textit{periodogram-spaced knots} or short \textit{P-spaced knots}.
+%REMARK: HOW ABOUT "PERIODOGRAM-SPACED KNOTS"	OR 'P-SPACED KNOTS'  INSTEAD OF DISTRIBUTED KNOTS?
+
+The procedure can be summarized as follows:
+\begin{enumerate}
+	\item Let $\bm{x} = \left (\sqrt{I_n(\lambda_1)}, \dots, \sqrt{I_n(\lambda_{\nu})} \right)$ be the square root of the periodogram values for each angular frequency.
+	\item Define $z_i = \left|\dfrac{x_i - \overline{x}}{S_x} \right|$ and calculate its distribution function $F_z = P(z \leq z_i)$.
+	\item Let $\bm{\xi}^* = \left(\xi_1^*,\dots,\xi_{K^*}^* \right)$ be a vector of $K^*$ equally spaced values in $[0,1]$.
+	\item The P-spaced knot vector is given by $\bm{\xi} = \left(F_{z}^{-1}(\xi^*_1),\dots, F_{z}^{-1}(\xi^*_{K^*}) \right)$.
+
+\end{enumerate}
+
+%REMARK: CAN YOU PLEASE INSERT EXPLICIT FORMULAE.
 
@@
 
-We assess our P-spline prior approach and compared it to B-spline prior via the simulation study, and the sunspot dataset used by \cite{Edwards2018}.
-In addition, we assess the impact of the knot location on the  P-spline estimates using the S.
-Carinae data, case in which the PSD contains abrupt peaks.
+We first assess the properties of our P-spline spectral density estimates in a simulation study and compare them to estimates based on the BspDP prior. We furthermore apply all approaches to the
+analysis of the classical sunspot dataset  and the {\it S.\ Carinae} variable star light intensities, previously analyzed by \cite{Cart:1997,Huerta:1999,Kirch:2018}).
+As the psd of the  {\it S.\ Carinae} time series contains sharp peaks, its analysis allows to assess the impact of the periodogram-spaced knots in contrast to the equidistant knots.
 For all the analyses, we use cubic B-splines ($r=3$).
@@
 
-This simulation study was performed by \cite{Edwards2018} in order to compare the B-spline to the Bernstein polynomial priors.
-We perform the same analysis, varying only the number of iterations used in the analysis.
-The study consists on two autoregressive (AR) time series models of order 1 and 4.
-For AR(1), a first autocorrelation order $\rho_1 = 0.9$ is considered, which produces a relatively simple spectral density.
-For AR(4), the parameters considered are $\rho_1~=~0.9$, $\rho_2~=~-~0.9$, $\rho_3~=~0.9$ and $\rho_4~=~-~0.9$, which produces a spectral density with two large peaks.
-From these models, 300 time series with unit variance Gaussian innovations of length $n = \{128, 256, 512\}$ are generated.
+The setup of our simulation study mirrors that  in \cite{Edwards2018} who  compared spectral density estimates based on the BspDP and the BDP priors. 
+% We perform the same analysis, varying only the number of iterations used in the analysis.  
+We generated  300 time series autoregressive time series of order 1 and 4 with unit variance Gaussian innovations of length $n = \{128, 256, 512\}$.
+For the AR(1) model, the  first  order autocorrelation $\rho_1 = 0.9$ was chosen, which has a relatively simple spectral density (see Figure~\ref{fig:ar1}).
+The AR(4) time series with $\rho_1~=~0.9$, $\rho_2~=~-~0.9$, $\rho_3~=~0.9$ and $\rho_4~=~-~0.9$ have a spectral density with two large peaks (see Figure~\ref{fig:ar4}).
 
-We estimate the spectral density functions using the B-spline and P-spline priors.
-The B-spline analysis is performed via the \texttt{gibbs\_bspline} function in the \textsf{R} package \texttt{bsplinePsd} \cite{Edwards:bsplinePsd:2018}.
-The algorithm was executed for 100,000 iterations with a burn-in period of 25,000 and thinning factor of 10, resulting in 7,500 samples used for the inferences.
+We estimated the spectral density functions using the BspDP and P-spline priors.
+The BspDP-based  analysis was performed via the \texttt{gibbs\_bspline} function in the \textsf{R} package \texttt{bsplinePsd} \cite{Edwards:bsplinePsd:2018}.
+The algorithm was executed for 100,000 iterations with a burn-in period of 25,000 and thinning factor of 10, resulting in 7,500 samples used for posterior inference.
 Pilot analyses showed that these specifications are suitable in these examples in terms of convergence (results not shown).
 
-The P-spline analysis is performed using the \texttt{gibbs\_pspline} function implemented in the \textsf{R} package \texttt{psplinePsd}.
-Equally spaced knots are assumed.
-The algorithm was executed for a total of 100,000 iterations.
-A pilot sample of 25,000 iterations with a burn-in period of 5,000 was used to calibrate the proposals.
-The final sample consisted in 75,000 samples with a burn-in period of 5,000 and thinning factor of 10, resulting in 7,000 samples used in the inferences.
-Therefore, the number of iterations equate the B-spline analysis.
+The P-spline analysis was performed using the \texttt{gibbs\_pspline} function implemented in the \textsf{R} package \texttt{psplinePsd}.  A total of 100,000 MCMC samples were generated
+using  a pilot run of 25,000 iterations with a burn-in period of 5,000 to calibrate the proposals.
+The final sample consisted of 75,000 samples with a burn-in period of 5,000 and thinning factor of 10, resulting in 7,000 samples used for posterior inferences.
+%Therefore, the number of iterations equate the B-spline analysis.  
+This procedure was replicated for both equidistant and P-spaced knots.
 
@@
 where $\sigma^2$ is the variance of the white noise innovations and ($\rho_1, \dots, \rho_p$) are the model parameters.
-Therefore, the estimates can be compared to the true function.
-For this purpose we use the integrated absolute error (IAE) or $L_1$-error, which is a measure of closeness and accuracy and is defined as
+In order to compare the psd estimates  to the true function, we use the integrated absolute error (IAE) or $L_1$-error which is defined as
 \begin{align*}
@@
 \end{align*}
-where $\widehat{f}(\cdot)$ is the Monte Carlo estimate of the spectral density $f(\cdot)$.
-We consider the median as spectral density estimate $\widehat{f}(\cdot)$.
-This measure is calculated for each of the 300 replications and its median is calculated.
+where $\widehat{f}(\cdot)$ is the pointwise posterior median of the spectral density $f(\cdot)$.   The IAE is computed for each of the 300 replications and its median is calculated.  
 The results are displayed in Table~\ref{table:sim_IAE}.
 
-For the AR(1) cases, the P-spline prior yields more accurate results than the B-spline prior for $n=128$.
-Slightly better results are obtained for $n=256$.
-For $n=512$, only the first order penalty P-splines yield better results, whereas the second and third order penalty P-splines yield marginally higher median IAE values.
+For the AR(1) time series, the P-spline prior with equidistant knots and penalty order $d=1$ yields better psd estimates than the BspDP prior.
+A larger penalty order does not necessarily result in more accurate psd estimates.
+In this case of a relatively smooth spectral density, the P-spaced knots do not yield an improvement over the equidistant knots but slightly better or similar accuracy than the BspDP prior.
+% yields more accurate results than the B-spline prior for $n=128$.
+Slightly better results are obtained for $n=256$ with equidistant knots.
+However, the distributed knot scheme yields %slightly higher values.
+For $n=512$ and both knot schemes, only the first order penalty P-splines yield better results, whereas the second and third order penalty P-splines yield marginally %higher median IAE values.
+In general, the equidistant knot scheme yield lower IAE values than the distributed knots.
 
-For the AR(4) cases, the P-spline prior yields more accurate results than the B-spline prior for $n=\{128, 256\}$.
-For $n=512$, the third order penalty P-splines has a slightly higher median IAE.
-Overall, P-spline method produces better results.
+For the AR(4) time series, the P-spline prior with P-spaced knots categorically outperforms  the BspDP algorithm.
+It also outperforms the equidistant knot scheme.
+The latter yields more accurate results than the BspDP prior for $n=\{128, 256\}$.
+For $n=512$, the third order penalty P-splines has a slightly higher median IAE, but overall, this approach achieves higher accuracy.
+As shown by \cite{Edwards2018}, the BspDP prior produces better results than the BDP in terms of median IAE in this example.  Consequently, our P-spline algorithm outperforms the BDP method as well.
 
-As shown by \cite{Edwards2018}, the B-spline prior produces better results than the Bernstein polynomial in terms of median IAE in this example.
-Consequently, our P-spline algorithm outperforms the Bernstein method in this context.
+%REMARK: PLEASE INSERT GRAPH WITH AR(1) AND AR(4) PSD ESTIMATES
 
@@
 	\centering
-	\begin{ruledtabular}
+%	\begin{ruledtabular}
 		\begin{tabular}{lccc}
@@
 			B-spline &  1.257 & 0.738 & 0.605  \\
-		P-spline $d=1$ &  0.729 & 0.632 & 0.517  \\
-		P-spline $d=2$&  0.844 & 0.612 & 0.638  \\
-		P-spline $d=3$&  0.727 & 0.728 & 0.769  \\ \hline
+			\textit{Equidistant knots}& & &   \\
+			\hspace{0.5em}P-spline $d=1$ &  0.729 & 0.632 & 0.517  \\
+			\hspace{0.5em}P-spline $d=2$&  0.844 & 0.612 & 0.638  \\
+			\hspace{0.5em}P-spline $d=3$&  0.727 & 0.728 & 0.769  \\
+			\textit{P-spaced knots}& & &   \\
+			\hspace{0.5em}P-spline $d=1$ & 0.910 & 0.769 & 0.595 \\
+			\hspace{0.5em}P-spline $d=2$ & 0.994 & 0.867 & 0.731 \\
+			\hspace{0.5em}P-spline $d=3$ & 1.059 & 0.829 & 0.610 \\ \hline
 			AR(4)     &  &  &  \\ 
 			B-spline & 4.069 & 3.016 & 2.401   \\
-		P-spline $d=1$ & 3.014 & 2.462 & 2.335  \\
-		P-spline $d=2$& 3.247 & 2.619 & 2.399 \\
-		P-spline $d=3$& 3.438 & 2.786 & 2.510  \\		
+			\textit{Equidistant knots}& & &   \\
+			\hspace{0.5em}P-spline $d=1$ & 3.014 & 2.462 & 2.335  \\
+			\hspace{0.5em}P-spline $d=2$& 3.247 & 2.619 & 2.399 \\
+			\hspace{0.5em}P-spline $d=3$& 3.438 & 2.786 & 2.510  \\	
+			\textit{P-spaced knots}& & &   \\
+			\hspace{0.5em}P-spline $d=1$ & 2.659 & 1.884 & 1.465 \\
+			\hspace{0.5em}P-spline $d=2$ & 2.627 & 1.849 & 1.411 \\
+			\hspace{0.5em}P-spline $d=3$ & 2.625 & 1.835 & 1.410 \\
 			\bottomrule
 		\end{tabular}
-	\end{ruledtabular}
-	\caption{Median IAE for the estimated spectral densities using the B-spline and P-spline priors on simulated AR(1) and AR(4) data.
-Different penalty order values ($d$) are considered for the P-spline prior algorithm.}
+%	\end{ruledtabular}
+	\caption{Median IAE for the psd estimates using the BspDP and the P-spline priors for simulated AR(1) and AR(4) time series with  different penalty orders ($d$) and knot location schemes for the P-splines.}
 	\label{table:sim_IAE}
 \end{table}
+
+\begin{figure}[]
+	\centering
+	\includegraphics[scale=0.4,clip=true,angle=0]{ar1.pdf}
+	\caption{Estimated log-spectral density for an AR(1) time series via the P-spline prior algorithm.  The solid line stands for the true psd, whereas gray area is the uniform 90\% credible band.}
+	\label{fig:ar1}
+\end{figure}
+
+\begin{figure}[]
+	\centering
+	\includegraphics[scale=0.4,clip=true,angle=0]{ar4.pdf}
+	\caption{Estimated log-spectral density for an AR(4) time series via the P-spline prior algorithm.  The solid line stands for the true psd, whereas gray area is the uniform 90\% credible band.}
+	\label{fig:ar4}
+\end{figure}
 
@@
 
-The IAE measures the closeness and accuracy of the tested methods as point estimate, but we could be interested in testing interval estimates.
+The IAE measures the accuracy of the  point estimates of the spectral density. In the following, we will also compare the coverage of posterior uniform credible bands.
+%but we could be interested in testing interval estimates.
 To assess this estimation, the uniform credible band is considered.
-This approach allows to measure the coverage levels for the whole spectral density.
+This approach allows to measure the coverage levels for the %whole spectral density.
 We define the $100(1-\alpha)\%$ uniform credible band as
@@
 
-For AR(1) case, both methods cover the whole true spectral density function for all the frequencies $\lambda$.
-For AR(4) case there are some variations.
-For the first penalty order P-splines, the coverage probabilities are all higher.
-For the second order penalty, our proposal has a better performance, except for $n=512$, where the values are equal.
+For the AR(1) model, all 90\% credible bands cover the whole true spectral density function, i.e.\ for all frequencies $\lambda$.
+For the AR(4) model, the P-spline prior with P-spaced knots has the best coverage accross the different specifications.
+% Only one of its coverage probabilities is not equal to 1 (for $d=3$), but it is close.  
+For the equidistant knots the coverage is always greater than 90\%.
+For the first penalty order P-splines, the coverage probabilities are all higher than the ones produced by the BspDP prior algorithm.
+For the second order penalty, it has a better performance, except for $n=512$, where the values are equal.
 For the third order penalty, we see a mix of results, that is equal, higher and lower values for the increasing sequence of $n$, respectively.
-For the P-spline case, in particular for $n=512$, the coverage probability decreases as the penalty degree order increases, which can be explained by the forced smoothness imposed by the higher penalty order degree.
-In general, P-spline prior have a better coverage probability than the B-spline method.
-\cite{Edwards2018} showed that the Bernstein polynomial method has a coverage probability 0 for $n = \{128,256,512\}$ in the AR(4) case.
+For the P-spline priors, in particular for $n=512$, the coverage probabilities decrease as the penalty degree order increases, which can be explained by the forced smoothness imposed by the higher penalty order.
+In general, P-spline priors have a better coverage probability than the BspDP method which
+as shown by  \cite{Edwards2018} have better coverage probabilities than the BDP prior in the AR(4) case.
 
@@
 			B-spline &  1.000 & 1.000 & 1.000 \\
-		P-spline $d=1$ & 1.000 & 1.000 & 1.000  \\
-		P-spline $d=2$& 1.000 & 1.000 & 1.000 \\
-		P-spline $d=3$& 1.000 & 1.000 & 1.000 \\ \hline		       
+			\textit{Equidistant \&}& & &   \\
+			\textit{P-spaced knots}& & &   \\
+			\hspace{0.5em}P-spline $d=1,2,3$ & 1.000 & 1.000 & 1.000  \\ \hline
+			%		P-spline $d=2$& 1.000 & 1.000 & 1.000 \\
+			%		P-spline $d=3$& 1.000 & 1.000 & 1.000 \\ \hline		       
 			AR(4)     &  &  &  \\ 
 			B-spline &  0.968 & 0.945 & 0.980  \\
-		P-spline $d=1$ & 0.984 & 0.992 & 0.984  \\
-		P-spline $d=2$& 0.984 & 0.984 & 0.980 \\
-		P-spline $d=3$& 0.968 & 0.976 & 0.969  \\		
+			\textit{Equidistant knots}& & &   \\
+			\hspace{0.5em}P-spline $d=1$ & 0.984 & 0.992 & 0.984  \\
+			\hspace{0.5em}P-spline $d=2$& 0.984 & 0.984 & 0.980 \\
+			\hspace{0.5em}P-spline $d=3$& 0.968 & 0.976 & 0.969  \\
+			\textit{P-spaced knots}& & &   \\	
+			\hspace{0.5em}P-spline $d=1$ & 1.000 &  1.000 &  1.000 \\
+			\hspace{0.5em}P-spline $d=2$ & 1.000 &  1.000 &  1.000 \\
+			\hspace{0.5em}P-spline $d=3$ & 0.984 &  1.000 &  1.000 \\	
 			\bottomrule
 		\end{tabular}
-	\end{ruledtabular}
-	\caption{Coverage probabilities based on $90\%$ uniform credible bands.}
+%	\end{ruledtabular}
+	\caption{Coverage probabilities of $90\%$ uniform credible bands.}
 	\label{table:sim_covPbb}
@@
 
-As \cite{Edwards2018} noted, one of the B-spline drawbacks is its computational complexity relative to the Bernstein method.
-This is directly reflected on the run-time, which is approximately 2-3 times higher in this example (see Table~3 in \cite{Edwards2018}).
-The median run-time in our analyses are displayed in Table~\ref{table:sim_time}.
-The P-spline method is approximately 4-6 times faster than B-spline approach in these examples.
-This is basically because it does not involve as many calculations as the latter per iteration, namely, the B-spline densities are calculated once at the start of the algorithm and there are not Dirichlet process approximations via the stick-breaking method.
+As \cite{Edwards2018} noted, one of the drawbacks of the BspDP prior  is its computational complexity relative to the BDP.
+This is directly reflected in the run-time, which is approximately 2-3 times higher in this example (see Table~3 in \cite{Edwards2018}).
+The median run-time in our analyses for the BspDP prior and the P-spline with P-spaced knots are displayed in Table~\ref{table:sim_time}.
+Note that the P-spline method based on equidistant knots has similar run-times to the P-spaced knots.
+The P-spline method is approximately 4-6 times faster than BspDP approach in these examples.
+This is basically because it does not involve as many calculations as the latter per iteration because the B-spline densities are calculated only  once at the start of the algorithm and there are no Dirichlet process approximations via the stick-breaking method.
 
@@
 			B-spline &   30.872 & 34.873 & 42.738 \\
-		P-spline $d=1$& 5.506 & 8.185 & 10.338 \\
-		P-spline $d=2$& 5.490 & 8.208 & 10.300 \\
-		P-spline $d=3$& 5.515 & 8.176 & 10.352 \\ 
-		B-spline/P-spline & 5.609 & 4.258 & 4.137 \\ \hline
+			\textit{P-spaced knots}& & &   \\
+			\hspace{0.5em}P-spline $d=1$& 5.629 & 8.291 & 10.507 \\
+			\hspace{0.5em}P-spline $d=2$& 5.650 & 8.374 & 10.512 \\
+			\hspace{0.5em}P-spline $d=3$& 5.587 & 8.301 & 10.427 \\ 
+			B-spline/P-spline &  5.491 & 4.190 & 4.077 \\ \hline
 			AR(4)     &  &  & \\ 
 			B-spline & 31.663 & 37.066 & 47.604  \\
-		P-spline $d=1$ & 5.538 & 8.410 & 10.443  \\
-		P-spline $d=2$&5.501 & 8.187 & 10.402 \\
-		P-spline $d=3$& 5.590 & 8.249 & 10.330  \\	
-		B-spline/P-spline & 5.712 & 4.475 & 4.581 \\	
+			\textit{P-spaced knots}& & &   \\
+			\hspace{0.5em}P-spline $d=1$ & 5.621 & 8.240 & 10.462 \\
+			\hspace{0.5em}P-spline $d=2$& 5.641 & 8.372 & 10.549 \\
+			\hspace{0.5em}P-spline $d=3$& 5.611 & 8.327 & 10.686 \\
+			B-spline/P-spline & 5.630 & 4.459 & 4.506 \\	
 			\bottomrule
 		\end{tabular}
-	\end{ruledtabular}
+%	\end{ruledtabular}
 	\caption{Median run-times in minutes and relative run-times respect to an average of the P-spline times.}
@@
 Overall, the P-spline prior outperforms the B-spline prior in this example in terms of IAE, coverage probability and run-time.
-Particularly, the first order penalty P-splines outperforms the B-spline method in all our tests.
-The criterion which allows to choose automatically the number of B-spline densities in our approach, that is $\min\left\{n/4, 40\right\}$, works perfectly well in these examples.
+Particularly, our novel knot allocation scheme, the periodogram-spaced knots, yield the best results in the case of peaked spectral densities.
+The  rule-of-thumb criterion,  $K=\min\left\{n/4, 40\right\}$, for the number of B-spline densities in the P-spline prior algorithm worked perfectly well in this example.
 
 \subsection{Sunspot data analysis}
-
-Sunspots are darker and cooler regions of the sun's surface.
-We analyse the average annual mean sunspot number for the year 1700-1987, that is 288 observations.
-This is a classic dataset used to assess power posterior estimation methods.
+Sunspots are regions of reduced surface temperature that are visible as darker spots on the sun's photosphere.
+We analyse the average annual mean sunspot numbers for the year 1700-1987 consisting of  288 observations.
+This is a classic dataset used to assess spectral density estimation methods.
 A square root transformation is applied in order to make the observations more symmetrical and stationary.
 
-As in \cite{Edwards2018}, the B-spline prior MCMC algorithm is run for 100,000 iterations, with a burn-in period of 50,000 and thinning factor of 10.
+As in \cite{Edwards2018}, the MCMC algorithm based on the BspDP prior  is run for 100,000 iterations, with a burn-in period of 50,000 and thinning factor of 10.
 Thus, 5,000 samples are used for the estimation.
@@
 
-The P-spline prior MCMC algorithm is run in two stages.
-The first one is used to calibrate the proposal in the final run.
-This is run for 25,000 iterations with a burn-in period of 10,000 and thinning factor of 10, resulting in 1,500 samples.
-The final analysis is run for 75,000 iterations with a burn-in period of 25,000 and thinning factor of 10, resulting in 5,000 samples, which is the same number of samples used for the B-spline analysis.
-The second order penalty is considered in this example.
-The analysis took around 6 minutes.
-The rest of specifications are the same as those used in the simulation study.
+The MCMC algorithm based on the P-spline prior is run in two stages.
+In  the first stage, the proposal  is calibrated for the final run.
+It consists of 25,000 iterations with a burn-in period of 10,000 and thinning factor of 10, resulting in 1,500 samples.
+Then the algorithm is run for 75,000 iterations with a burn-in period of 25,000 and thinning factor of 10, resulting in 5,000 samples.
+The analysis is replicated for the first, second and third order penalties and for equidistant and P-spaced knots.
+All other specifications are the same as those used in the simulation study.
+A single analysis takes around 6 minutes.
 
@@
 	\centering
-	\includegraphics[scale=0.5,clip=true,angle=0]{sunspot.pdf}
+	\includegraphics[scale=0.40,clip=true,angle=0]{sunspot.pdf}
 	\caption{Log-spectral density estimate for the sunspot data.
 The continuous gray line represents the periodogram.
-The continuous and dashed lines stand for the posterior median log-PSD obtained using the B-spline and P-spline priors, respectively.
-The dot lines represents the 90\% pointwise credible bands for the latter prior.}
+The continuous and dashed lines stand for the posterior median log psd obtained using the B-spline and P-spline priors, respectively.
+The dot lines represents the 90\% pointwise credible bands for the latter prior, which consider a first order penalty and P-spaced knots.}
 	\label{fig:sunspot}
@@
 
-The results are displayed in Figure~\ref{fig:sunspot}.
-Both methods reveal a large peak at about 0.57.
-In other words, there is a periodic solar cycle of length $2 \pi / 0.57 \approx 11$ years.
-These results are consistent with those obtained via the Bernstein polynomial prior \citep{Choudhuri:2004}.
+%REMARK: PLEASE CHANGE THE TICK MARKS ON X-AXIS IF THESE ARE CYCLES/YEAR NOT ANGULAR FREQUENCIES.\smallskip
 
-The 90\% credible bands estimated via the P-spline prior algorithm cover 98\% of the times the posterior median log-PSD estimated via the B-spline prior method.
-These results are excellent considering the saving in computational time: B-spline prior method takes approximately 32 minutes, whereas the P-spline prior method only takes around 6 minutes, i.e., approximately 5 times more efficient.
+The results for the BspDP and P-spline prior  with first order penalty and P-spaced knots are displayed in Figure~\ref{fig:sunspot}.
+Both methods reveal a large peak at about 0.09.
+In other words, there is a periodic solar cycle of length $1/ 0.09 \approx 11$ years.
+These results are consistent with those obtained via the BDP  prior~\citep{Choudhuri:2004}.
+However, the BspDP estimation takes approximately 32 minutes compared to only 6 minutes for the the P-spline estimation. Thus the P-spline estimation is approximately
+five times more efficient.
+
+%REMARK: I WOULD  DELETE THE NEXT PARAGRAPH AND TABLE AS THIS SUGGESTS THAT THE BSPDP ESTIMATE IS THE GOLD STANDARD. BUT THE AR(4) SIMULATION STUDY HAS SHOWN THAT THIS IS NOT THE CASE. 
+%We also estimate the 90\% credible bands obtained via the P-spline prior algorithm and calculate the proportion of times that these encapsulate the posterior median log psd obtained using the B-spline prior.
+The results are displayed in Table~\ref{table:sunsont_knots}.
+The coverage is quite high in general and decreases as the penalty order increases.
+%\begin{table}
+%	\centering
+%	\begin{ruledtabular}
+%		\begin{tabular}{lccc}
+%			\toprule
+%			                   & $d=1$ & $d=2$ & $d=3$ \\ \hline
+%			Equidistant knots  & 0.98 & 0.98 & 0.94 \\ 
+%			Distributed knots  & 1.00 & 0.94 & 0.85  \\			
+%			\bottomrule
+%		\end{tabular}
+%	\end{ruledtabular}
+%	\caption{Proportion of times that the posterior median log psd obtained via B-spline prior algorithm is encapsulated within the 90\% uniform credible band obtained via the P-spline prior algorithm.}
+%	\label{table:sunsont_knots}
+%\end{table}
 
@@
 
-This dataset contains 1189 visual observations of the S.
-Carinae, a variable star in the southern hemisphere sky.
+This dataset contains 1189 visual observations of the {\it S.\ Carinae}, a variable star in the southern hemisphere sky.
 These are daily observations collected by the Royal Astronomical Society of New Zealand, which correspond to 10 day averages of light intensities over several years.
 This data set has been analysed in the literature previously (see for instance \cite{Cart:1997,Huerta:1999,Kirch:2018}).
-It contains 40 missing observations that in this work have been replaced by the mean.
-This technique does not affect the general features of the data.
+It contains 40 missing observations that in this work have been replaced by the mean, thus not affecting the general features of the data.
 In addition, the data has been squared root transformed and mean centred.
 
-\cite{Kirch:2018} provided a PSD estimate via a nonparametrically corrected (NPC) approach, which is able to detect several peaks in the function.
-These features makes this data set suitable to assess the impact of the knot location on the P-spline prior estimate.
-We test 2 knot location approaches: equally spaced internal knots and distributed according to the periodogram as defined above.
+\cite{Kirch:2018} provided a psd estimate via a nonparametrically corrected (NPC) approach, which is able to detect several peaks in the function.
+These features make this data set suitable to assess the impact of the knot location on the P-spline estimates.
 
-The P-spline algorithm is run in two stages.
+Again, the P-spline algorithm is run in two stages.
 First, a Markov chain of 5,000 samples with a burn-in period of 1,000 and thinning factor of 10, resulting in 400 points, is used to calibrate the proposals of the final run.
-Then, a Markov chain of 10,000 samples with a burn-in period of 2,000 and thinning factor of 10, that is 800 points, is used for the PSD estimation.
-We use 40 B-spline densities, number based on the criterion discussed above, and the second order penalty.
-The whole process takes less than 2 minutes and is replicated for the 2 knot location approaches.
+Then, a Markov chain of 10,000 samples with a burn-in period of 2,000 and thinning factor of 10, that is 800 points, is used for the psd estimation.
+We use 40 B-spline densities,  based on the rule-of-thumb criterion discussed above, and the second order penalty.
+The whole process takes less than 2 minutes to run and is replicated for equidistant and P-spaced knots.
 The results are displayed in Figure~\ref{fig:carinae}.
@@
 	\centering
-	\includegraphics[scale=0.5,clip=true,angle=0]{carinae.pdf}
-	\caption{Log-spectral density estimate for the variable star S.
-Carinae data via the P-spline prior algorithm.
-The continuous gray line represents the periodogram whereas the continuous and dashed black lines stand for the posterior median log-PSD obtained using the periodogram knot location scheme and equally spaced knots, respectively.}
+	\includegraphics[scale=0.4,clip=true,angle=0]{carinae.pdf}
+	\caption{Log-spectral density estimate for the variable star {\it S.\ Carinae} data based on the P-spline prior.
+The continuous gray line represents the periodogram whereas the continuous and dashed black lines stand for the posterior median log psd obtained using P-spaced and equidistant knots, respectively.}
 	\label{fig:carinae}
@@
 
-The periodogram shows several peaks, being the main one located at around 0.42.
-When the knots are equally spaced, the algorithm is unable to capture properly this characteristic, unlike our knot location proposal.
-This is because our approach allocates more knots in those areas which contain significant peaks detected via the periodogram.
-The rest of peaks, which are smaller in comparison to the main one, are slightly better captured by the equally spaced knot approach, since the other scheme spend most of the effort in the main peak, but which can be corrected by increasing the number of knots.
-Allowing the knots to vary according to the periodogram, the PSD estimate is more similar to the one obtained via the NPC approach \citep{Kirch:2018}.
+The periodogram shows several peaks with the main one located at around 0.42.  When the knots are equally spaced, the algorithm is unable to capture the sharpness of this peak,
+in contrast to the P-spaced knots.
+This is because the periodogram-based knot allocation puts more knots in those regions which contain significant peaks detected by the periodogram.
+The rest of peaks, which are smaller in comparison to the main one, are slightly better captured by the equidistant knots, since a large proportion of the P-spaced knots have been allocated around the main peak.
+However, this can easily  be corrected by increasing the number of knots.
+The psd estimate based on P-spaced knots is comparable to the one obtained via the NPC approach \citep{Kirch:2018}.
+
+%REMARK: TO SUPPORT THIS CLAIM, SHOULD WE INCLUDE ANOTHER GRAPH THAT COMPARES THE NPC ESTIMATE WITH P-SPACE P-SPLINES WITH LARGER NUMBER OF KNOTS?
 
@@
 
-As shown by \cite{Edwards2018}, the B-spline prior outperforms the Bernstein polynomial prior in terms of IAE and uniform coverage probabilities for complicated PSDs.
+As shown by \cite{Edwards2018}, the BspDP prior outperforms the BDP prior in terms of IAE and uniform coverage probabilities for spiky psds.
 This is explained by the local support of the B-splines and its better approximation properties \citep{Edwards2018}.
-However, its performance involves a much higher computational cost.
-These characteristics have motivated our approach.
+However, its performance is only achieved at a much higher computational cost.
+These characteristics have motivated our P-spline approach.
 
-The P-spline prior can be considered a particular case of the B-spline prior.
+The P-spline prior can be regarded as a special case of the B-spline prior.
 The latter works with a variable number and location of knots, which are driven by the data.
-The P-spline prior assumes a fixed number of knots, with fixed locations.
+The P-spline prior assumes a fixed number of knots  with fixed locations.
 These simplifications reduce the computational effort significantly, without sacrificing accuracy as shown in the application section.
-To deal with abrupt peaks in the PSD, we propose to locate the knots according to the peaks detected via the periodogram.
+To deal with abrupt peaks in the psd, we propose to locate the knots according to the peaks detected via the periodogram.
 This approach does not affect the computational time and improve the estimates dramatically.
+This also indicates that the BspDP computation time could be reduced by the choice of P-spaced knots as initial values.
 
-In our simulation study, we showed that, in general, the P-spline prior outperforms the B-spline prior in all our assessments, that is in terms of IAE, coverage probabilities, and run-times.
-From this example, it seems that a second order penalty P-spline works reasonably well.
+In our simulation study, we showed that, in general, the P-spline prior outperforms the BspDP prior in all our assessments, that is in terms of IAE, coverage probabilities, and run-times.
+From the simulation study, it seems that a second order penalty P-spline works reasonably well.
 
 We also assessed our proposal in the classic sunspot dataset, in which the B-spline prior algorithm is able to estimate correctly the solar cycle that occurs every 11.07 years.
-The P-spline prior yields almost identical results, but in a significant less computational time.
+The P-spline prior yields almost identical results, but in  significantly  less computational time.
 
-Finally, we assess the P-spline prior algorithm under two different knot location schemes: equally spaced and distributed according to the periodogram.
-For this, we estimated the PSD for the S.
-Carinae data, which function contains several abrupt peaks.
-The equal spaced scheme fails in capturing the peaks of the function.
-On the other hand, our proposal improved significantly the results without affecting the computational time.
+Finally, we assess the P-spline prior algorithm under two different knot location schemes: equally spaced and periodogram-spaced knots.
+For this, we estimated the psd for the {\it S.\ Carinae} time series which has a spectral density with several sharp peaks.
+Equidistant knots failed to capture the peaks of this spectral density whereas  P-spaced knots
+significantly improved  the results without affecting the computational time.
 
 The number of B-spline densities was selected according to the criterion proposed by \cite{Ruppert2002}.
-It seems to work well, at least for the kind of PSDs treated in this work.
+It seems to work well, at least for the spectral densities that we experimented with.
 However, more complex functions could require larger number of B-spline densities.
@@
 
-In conclusion, the P-spline prior represents a more viable alternative to the B-spline prior for spectral density estimation.
-As shown in this work, it can handle quite well complex PSDs in a reasonable time-run.\\
+In conclusion, the P-spline prior represents a more computationally viable alternative to the BspDP prior for spectral density estimation.
+As shown in this work, it can handle spectral densities with several peaks in a reasonable run-time.
 
@@
 
-%\pagebreak
-
-%%%%%%%%%%
-
-\bibliographystyle{apsrev4-1}
-
-\bibliography{pSplinesReferences}
+%\bibliographystyle{spmpsci}
+\bibliographystyle{plainnat}
+\bibliography{pSplinesReferences}  % Name of BibTeX file
 
 \end{document}
-
-%\begin{thebibliography}{10}
-
-%\end{thebibliography}
-%%%%%%%%%%
-
-\end{document}

### 2019-04-18 R calculations/runTimeComparisonSplines.R

@@ -21,6 +21,8 @@ x
 ### P-spline time estimates ###
 ###############################
 
+# Equidistant knots
+
 y = matrix( c(30.872, 34.873, 42.738,
               5.506, 8.185, 10.338, 
               5.490, 8.208, 10.300,
@@ -43,5 +45,32 @@ y[5,]/y[7,]; # B-splines / Psplines d=2
 y[5,]/y[8,]; # B-splines / Psplines d=3
 
 round(y[5,]/apply(y[6:8,],2,mean),3)
+
+# Distributed knots
+
+y = matrix( c(30.872, 34.873, 42.738,
+              5.629, 8.291, 10.507,
+              5.650, 8.374, 10.512,
+              5.587, 8.301, 10.427,
+              31.663, 37.066, 47.604,  
+              5.621, 8.240, 10.462,
+              5.641, 8.372, 10.549,
+              5.611, 8.327, 10.686), ncol = 3, byrow = TRUE) 
+
+
+# AR(1)
+y[1,]/y[2,]; # B-splines / Psplines d=1
+y[1,]/y[3,]; # B-splines / Psplines d=2
+y[1,]/y[4,]; # B-splines / Psplines d=3
+
+round(y[1,]/apply(y[2:4,],2,mean),3)
+
+# AR(4)
+y[5,]/y[6,]; # B-splines / Psplines d=1
+y[5,]/y[7,]; # B-splines / Psplines d=2
+y[5,]/y[8,]; # B-splines / Psplines d=3
+
+round(y[5,]/apply(y[6:8,],2,mean),3)
+
 
 

### 2019-04-30 pSplines.tex
update

@@
 
-This article proposes a Bayesian approach to estimating the spectral density of a stationary time series using a  P-spline prior.
-Our proposal is motivated by the B-spline Dirichlet process prior of \cite{Edwards2018} combined with Whittle's likelihood with the aim of reducing the computational complexity of posterior computations.
-The strength of the B-spline Dirichlet process prior
-over the Bernstein-Dirichlet process prior of \cite{Choudhuri:2004}  lies in its ability to estimate spectral densities with sharp peaks and abrupt changes due to the flexibility of B-splines with variable number and location of knots.
-Here, we fix the number and location of the knots and specify a P-spline prior.
-We demonstrate in a simulation study and two real  case studies that this approach retains the flexiblity of the B-splines, achieves the same ability
-to accurately estimate peaks due to a new data-driven knot allocation scheme but significantly reduces the computational costs.
+This article proposes a Bayesian approach to estimating the spectral density of a stationary time series using a prior based on a mixture of P-spline distributions.
+Our proposal is motivated by the B-spline Dirichlet process prior of \cite{Edwards2018} in combination  with Whittle's likelihood and aims at  reducing the high computational complexity of its posterior computations.
+The strength of the B-spline Dirichlet process prior over the Bernstein-Dirichlet process prior of \cite{Choudhuri:2004}  lies in its ability to estimate spectral densities with sharp peaks and abrupt changes due to the flexibility of B-splines with variable number and location of knots.
+Here, we suggest to use P-splines of \cite{Eilers:1996} that combine a B-spline basis with a discrete penalty on the basis coefficients.
+We demonstrate in a simulation study and two real  case studies that this approach retains the flexibility of the B-splines, achieves the same ability to accurately estimate peaks due to a new data-driven knot allocation scheme but significantly reduces the computational costs.
 
@@
 L(\bm{Y} | f) &= \prod_{l=1}^{\nu}\dfrac{1}{f(\lambda_{l})} e^{-I_n(\lambda_l)/f(\lambda_l)}  \nonumber \\
-&\propto \exp \left( - \sum_{l=1}^{\nu} \log f(\lambda_l) + \dfrac{I_{n}(\lambda_l)}{f(\lambda_l)} \right)\label{eq:whittle_like}
+&\propto \exp \left( - \sum_{l=1}^{\nu} \log f(\lambda_l) + \dfrac{I_{n}(\lambda_l)}{f(\lambda_l)} \right),\label{eq:whittle_like}
 \end{align}
@@
 who used a smoothing prior on the log psd.
-\cite{Rosen:2007} decompose the log psd into a linear and nonlinear component and  put a Gaussian process prior on the nonlinear part, \cite{Pensky:2007} propose Bayesian wavelet-smoothing of the log psd and recently, and \cite{Cadonna2017} modeled the log psd  as a mixture
-of Gaussian distributions with frequency-dependent weights and mean functions.
+\cite{Rosen:2012} decompose the log psd into a linear and nonlinear component and  put a Gaussian process prior on the nonlinear part, \cite{Pensky:2007} propose Bayesian wavelet-smoothing of the log psd and recently, and \cite{Cadonna2017} modelled the log psd  as a mixture of Gaussian distributions with frequency-dependent weights and mean functions.
 %
 \cite{Gangopadhyay:1999} modelled the log-spectral density function by  piecewise polynomials of low order between strategically placed knots on the support of the psd,  allowing the data to select the number and location of the knots.
-A reversible jump Markov chain Monte Carlo (RJMCMC) \citep[RJMCMC;][]{Green:1995} algorithm was implemented to sample from the posterior distribution.
+A reversible jump Markov chain Monte Carlo \citep[RJMCMC;][]{Green:1995} algorithm was implemented to sample from the posterior distribution.
 
@@
 
-With the aim of reducing the computational cost of the BspDP prior approach,  we suggest to use the idea of P-splines \cite{Eilers:1996} that combine a large but fixed number of  B-splines with a simple difference penalty that controls the degree smoothness of the spectral density.
+With the aim of reducing the computational cost of the BspDP prior approach,  we suggest to use the idea of P-splines \cite{Eilers:1996} that combine a large but fixed number of  B-splines with a simple difference penalty that controls the degree of  smoothness of the spectral density.
 P-splines, i.e.\  equally-spaced B-splines,  have been successfully implemented for frequentist nonparametric regression but to the best of our knowledge not yet for Bayesian spectral density estimation.
@@
 
-\section{B-spline--Dirichlet process prior}
+\section{Notation and review of the B-spline--Dirichlet process prior}
 
@@
 \end{align*}
-is the knot vector.  This vector is a non-decreasing sequence that contains $K+r+1$ knots, which can be divided in $2(r+1)$ external and $K-r-1$ internal knots.  The latter must be $\geq r$.
+is the knot vector.  This vector is a non-decreasing sequence that contains $K+r+1$ knots, which can be divided in $2r$ external and $K-r+1$ internal knots.  The latter must be $\geq r$.
+% external and internal knots as in Edwards et al (2018)
+
+%This vector is a non-decreasing sequence that contains $K+r+1$ knots, which can be divided in $2(r+1)$ external and $K-r-1$ internal knots.  The latter must be $\geq r$.
 
@@
 
-\subsection*{Prior description}
+\subsection*{BspDP Prior}
 The BspDP prior is a mixture of B-spline densities defined as
@@
 
-To sum up, the BspDP prior has the following components:
-\begin{itemize}
-	\item $G \sim \text{DP}(M_G, G_0)$, where $M_G>0$ is the concentration parameter and $G_0$ is the base distribution.  It produces the weights for the $K$ B-spline densities.  
-	\item $H \sim \text{DP}(M_H, H_0)$, where $M_H>0$ is the concentration parameter and $H_0$ is the base distribution.
-It produces the knot locations determining the shape and location of the B-spline densities.
-	\item $p(K) \propto \text{exp}(-\theta_K K^2)$ for $K=1,\dots,K_{\text{max}}$.
-$K$ is the number of B-spline densities used in Eq.
-\eqref{eq:bs_prior}.
-It plays the role of smoothing parameter, that is smaller values implies smoother spectral density estimates.
-$K_{\text{max}}$ must be limited  for computational reasons.
-In practice, pilot runs are used to determine it.
-	\item $\tau \sim \text{Inverse-Gamma}(\alpha_{\tau},\beta_{\tau})$ is the normalizing constant.  $\alpha_{\tau}$ and $\beta_{\tau}$ are the shape and scale parameters, respectively.
-\end{itemize}
-
 %\subsection*{Implementation discussion}
@@
 
-$\bm{\xi}$ contains fixed equidistant internal knots on $[0,1]$ in our basic P-spline prior. A different  allotment of knots is described in  section ``knot allocation strategy" .
+$\bm{\xi}$ contains fixed equidistant internal knots on $[0,1]$ in our basic P-spline prior. A different  allotment of knots is described in  section ``knot allocation strategy".
 Therefore, the P-spline prior on the spectral density function is given by
@@
 putting a penalty on the integrated squared second derivative  as a measure of roughness.
-P-splines avoid derivatives by expressing the penalty as the sum of squares of differences of the B-psline coefficients.
+P-splines avoid derivatives by expressing the penalty as the sum of squares of differences of the B-spline coefficients.
 %The roughness penalty in a frequentist context is carried out by penalizing the likelihood function.  
-In a Bayesian context, this penalty can be translated into a prior distribution for the $d^{\text{th}}$ order difference of successive coefficients, see e.g. \citep{Lang:2004}.
+In a Bayesian context, this penalty can be transformed into a prior distribution for the $d^{\text{th}}$ order difference of successive coefficients, see e.g. \cite{Lang:2004}.
 %, known also as B-spline parameters.
@@
 \end{align*}
-where $\bm{v} = (v_1,\dots,v_{K-1})^\top$ is a $K-1$ dimensional parameter vector, $\phi$ is the smoothing or penalty parameter, $\textbf{P} = \textbf{D}^\top \textbf{D} +\epsilon \textbf{I}_{K-1}$ is the penalty matrix (which is a full matrix rank matrix for any small quantity $\epsilon$, for instance, $10^{-6}$) with $\textbf{D}$ the $d^{\text{th}}$ order difference matrix ($d=1,2,3,\ldots)$, $\alpha_{\phi}$ and $\alpha_{\delta}$ are shape parameters, $\delta \beta_{\phi}$ and $\beta_{\delta}$ are rate parameters, and $\text{Gamma}(a,b)$ denotes a gamma distribution with mean $a/b$ and variance $a/b^2$.
+where $\bm{v} = (v_1,\dots,v_{K-1})^\top$ is a $K-1$ dimensional parameter vector, $\phi$ is the smoothing or penalty parameter, $\textbf{P} = \textbf{D}^\top \textbf{D} +\epsilon \textbf{I}_{K-1}$ is the penalty matrix (which is a full matrix rank matrix for any small quantity $\epsilon$, for instance, $10^{-6}$) with $\textbf{D}$ the $d^{\text{th}}$ order difference matrix ($d=1,2,\ldots, K-2)$, $\alpha_{\phi}$ and $\alpha_{\delta}$ are shape parameters, $\delta \beta_{\phi}$ and $\beta_{\delta}$ are rate parameters, and $\text{Gamma}(a,b)$ denotes a gamma distribution with mean $a/b$ and variance $a/b^2$.
 We used first, second and third order difference penalties in the simulation and case study.
@@
 
-We take a robust specification for the penalty prior distribution as suggested by \cite{Jullion:2007}.
-Choosing small values for $\alpha_{\delta}$ and $\beta_{\delta}$, for instance, $10^{-4}$, the choice of $\alpha_{\phi}$ and $\beta_{\phi}$ do not affect the spectral density estimate.
+The parameter vector for the P-spline prior is $\bm{\theta} = (\bm{v}^\top, \phi, \delta, \tau)^\top$.
+We use a robust specification for the  prior distribution of the penalty parameters as suggested by \cite{Jullion:2007} by choosing small values for $\alpha_{\delta}$ and $\beta_{\delta}$, for instance, $10^{-4}$.
+The the choice of $\alpha_{\phi}$ and $\beta_{\phi}$ do not affect the spectral density estimate.
 Here we set $\alpha_{\phi}$ and $\beta_{\phi}$ equal to 1 as used by \cite{Bremhorst:2016}.
-These are the prior specifications used in Section \ref{sec:application}.
-
-The parameter vector for the P-spline prior is $\bm{\theta} = (\bm{v}^\top, \phi, \delta, \tau)^\top$.
+The scale parameter $\tau$ is given by an inverse Gamma($\alpha_{\tau},\beta_{\tau}$) prior.
+These are also the prior specifications used in the simulations and examples in Section \ref{sec:application}.
 
@@
 %Too few could cause an underfitting whereas too many an overfitting.  
-Even though the penalty parameter controls the smoothness of the fit, a large number of B-splines produces rougher estimates and  a low number produces smoother spectral densities.
+Even though the penalty parameter controls the smoothness of the fit, a large number of B-splines yields rougher and  a low number  smoother spectral density estimates.
 There is a smallest sufficient number which results in a psd estimate that fits the features of the data whereas choosing a larger number will have an insignificant effect on the fit.
 Unexpectedly, \cite{Ruppert2002} found some cases in which too many B-splines degrades the fitting in terms of mean square error.
-Unfortunately, there is no  consensus in how to find the optimal value.
+Unfortunately, there is no  consensus in how to find the optimal value of the number of B-splines.
 
 \cite{Eilers2015} consider the use of $K=100$ B-splines a wise choice, unless computational constraints are evident.
-Some algorithms that explore different number of knots $K^*$ in a trial sequence and find a relative optimal value according to certain goodness-of-fit criteria have been proposed.
+Some algorithms that explore different number of knots $K^*$ in a trial sequence and find an optimal value with respect to a certain goodness-of-fit criterion have been proposed.
 \cite{Ruppert2002} suggested a full-search and myopic algorithms which are based on the generalized cross-validation statistic.
-However, their results are not conclusive and this criterion only applies if the penalty parameter is treated as a tuning constant~\cite{Kauermann2011}.
+However, their results are not conclusive and this criterion only applies if the penalty parameter is treated as a tuning constant~\citep{Kauermann2011}.
 \cite{Likhachev2017} also considered a knot sequence and proposed the selection of the number of knots based on  Akaike's, corrected Akaike's and
@@
 The knots are often equally spaced in P-spline methods.
-This scheme works quite well in the P-spline prior algorithm for psds in which there are no abrupt changes.
+This scheme works quite well in the P-spline prior algorithm for smooth psds without any abrupt changes.
 However, for spectral densities with spikes, the algorithm will require a large number of knots to adequately estimate the peaks.
@@
 	\item Let $\bm{x} = \left (\sqrt{I_n(\lambda_1)}, \dots, \sqrt{I_n(\lambda_{\nu})} \right)$ be the square root of the periodogram values for each angular frequency.
-	\item Define $z_i = \left|\dfrac{x_i - \overline{x}}{S_x} \right|$ and calculate its distribution function $F_z = P(z \leq z_i)$.
+	\item Define $z_i = \left|\dfrac{x_i - \bar{x}}{S_x} \right|$ where $\bar{x}$ denotes the average and $S_x$ the standard deviation of ${\bm x}$ and calculate its distribution function $F_z = P(z \leq z_i)$.
 	\item Let $\bm{\xi}^* = \left(\xi_1^*,\dots,\xi_{K^*}^* \right)$ be a vector of $K^*$ equally spaced values in $[0,1]$.
@@
 We first assess the properties of our P-spline spectral density estimates in a simulation study and compare them to estimates based on the BspDP prior. We furthermore apply all approaches to the
-analysis of the classical sunspot dataset  and the {\it S.\ Carinae} variable star light intensities, previously analyzed by \cite{Cart:1997,Huerta:1999,Kirch:2018}).
+analysis of the classical sunspot dataset  and the {\it S.\ Carinae} variable star light intensities, previously analyzed by \cite{Cart:1997,Huerta:1999,Kirch:2018}.
 As the psd of the  {\it S.\ Carinae} time series contains sharp peaks, its analysis allows to assess the impact of the periodogram-spaced knots in contrast to the equidistant knots.
@@
 % We perform the same analysis, varying only the number of iterations used in the analysis.  
-We generated  300 time series autoregressive time series of order 1 and 4 with unit variance Gaussian innovations of length $n = \{128, 256, 512\}$.
+We generated  300  autoregressive time series of order 1 and 4 with unit variance Gaussian innovations of length $n = \{128, 256, 512\}$.
 For the AR(1) model, the  first  order autocorrelation $\rho_1 = 0.9$ was chosen, which has a relatively simple spectral density (see Figure~\ref{fig:ar1}).
@@
 For $n=512$, the third order penalty P-splines has a slightly higher median IAE, but overall, this approach achieves higher accuracy.
-As shown by \cite{Edwards2018}, the BspDP prior produces better results than the BDP in terms of median IAE in this example.  Consequently, our P-spline algorithm outperforms the BDP method as well.
+As shown by \cite{Edwards2018}, the BspDP prior yields better results than the BDP in terms of median IAE in this example.  Consequently, our P-spline algorithm outperforms the BDP method as well.
 
 %REMARK: PLEASE INSERT GRAPH WITH AR(1) AND AR(4) PSD ESTIMATES
-
+\def\arraystretch{1.1}
 % IAE
@@
 The P-spline method is approximately 4-6 times faster than BspDP approach in these examples.
-This is basically because it does not involve as many calculations as the latter per iteration because the B-spline densities are calculated only  once at the start of the algorithm and there are no Dirichlet process approximations via the stick-breaking method.
+This is due to a reduced number of calculations  per iteration because the B-spline densities are calculated only  once at the start of the algorithm and there are no Dirichlet process approximations via the stick-breaking method.
 
@@
 %	\end{ruledtabular}
-	\caption{Median run-times in minutes and relative run-times respect to an average of the P-spline times.}
+	\caption{Median run-times in minutes and relative run-times with respect to an average of the P-spline run-times.}
 	\label{table:sim_time}
@@
 Thus, 5,000 samples are used for the estimation.
-This took around 32 minutes.
+This took about 32 minutes.
 The other specifications are the same as those used in the simulation study.
@@
 Then the algorithm is run for 75,000 iterations with a burn-in period of 25,000 and thinning factor of 10, resulting in 5,000 samples.
-The analysis is replicated for the first, second and third order penalties and for equidistant and P-spaced knots.
+The analysis is performed using the first order penalty for equidistant and P-spaced knots.
 All other specifications are the same as those used in the simulation study.
-A single analysis takes around 6 minutes.
+A single analysis takes about 6 minutes.
 
@@
 in contrast to the P-spaced knots.
-This is because the periodogram-based knot allocation puts more knots in those regions which contain significant peaks detected by the periodogram.
+This is because the periodogram-based knot allocation puts more knots in those regions which contain significant peaks as indicated by the periodogram.
 The rest of peaks, which are smaller in comparison to the main one, are slightly better captured by the equidistant knots, since a large proportion of the P-spaced knots have been allocated around the main peak.
@@
 
-The P-spline prior can be regarded as a special case of the B-spline prior.
+Althought the P-spline prior is very similar to the B-spline prior as both are based on mixtures of B-spline distributions but there is quite a significant difference.
 The latter works with a variable number and location of knots, which are driven by the data.
-The P-spline prior assumes a fixed number of knots  with fixed locations.
-These simplifications reduce the computational effort significantly, without sacrificing accuracy as shown in the application section.
+The P-spline prior assumes a fixed number of knots  with fixed locations, thus almost looks like a special case.
+However, the multivariate prior based on the difference penalty of the basis coefficients controls the smoothness of the spectral density, a feature that can't be achieved with univariate Dirichlet process prior.
+Most importantly,
+ fixing the number and location of knots reduces the computational effort significantly, without sacrificing accuracy as shown in the application section.
 To deal with abrupt peaks in the psd, we propose to locate the knots according to the peaks detected via the periodogram.
-This approach does not affect the computational time and improve the estimates dramatically.
-This also indicates that the BspDP computation time could be reduced by the choice of P-spaced knots as initial values.
+This approach does not affect the computational time and improves the estimates dramatically.
+%This also indicates that the BspDP computation time could be reduced by the choice of P-spaced knots as initial values.
 
-In our simulation study, we showed that, in general, the P-spline prior outperforms the BspDP prior in all our assessments, that is in terms of IAE, coverage probabilities, and run-times.
-From the simulation study, it seems that a second order penalty P-spline works reasonably well.
+In our simulation study, we showed that, in general, the P-spline prior outperforms the BspDP prior in all our assessments  in terms of IAE, coverage probabilities, and run-times.
+From the simulation study, it seems that a {\em second order} penalty P-spline works reasonably well.
 
@@
 
+The \textsf{R} package \texttt{psplinePsd} will be available on CRAN in due course.  In the meanwhile, it can be downloaded from \href{https://github.com/pmat747/psplinePsd}{GitHub}.
+
 \begin{acknowledgements}
@@
 PM's and RM's work is supported by Grant 3714568 from the University of Auckland Faculty Research Development Fund and the DFG Grant KI 1443/3-1.
+RM gratefully acknowledges support by a James Cook Fellowship from Government funding, administered by teh Royal Society Te Ap\={a}rangi.
 All analysis was conducted in \textsf{R}, an open-source statistical software available on \textsf{CRAN} (cran.r-project.org).
@@
 \end{document}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+To sum up, the BspDP prior has the following components:
+\begin{itemize}
+	\item $G \sim \text{DP}(M_G, G_0)$, where $M_G>0$ is the concentration parameter and $G_0$ is the base distribution.  It produces the weights for the $K$ B-spline densities.  
+	\item $H \sim \text{DP}(M_H, H_0)$, where $M_H>0$ is the concentration parameter and $H_0$ is the base distribution.
+It produces the knot locations determining the shape and location of the B-spline densities.
+	\item $p(K) \propto \text{exp}(-\theta_K K^2)$ for $K=1,\dots,K_{\text{max}}$.
+$K$ is the number of B-spline densities used in Eq.
+\eqref{eq:bs_prior}.
+It plays the role of smoothing parameter, that is smaller values implies smoother spectral density estimates.
+$K_{\text{max}}$ must be limited  for computational reasons.
+In practice, pilot runs are used to determine it.
+	\item $\tau \sim \text{Inverse-Gamma}(\alpha_{\tau},\beta_{\tau})$ is the normalizing constant.  $\alpha_{\tau}$ and $\beta_{\tau}$ are the shape and scale parameters, respectively.
+\end{itemize}

### 2019-05-03 pSplines.tex
b-spline result update

@@
 two novel Bayesian  approaches in this paper, both based on the Whittle likelihood and on a mixture of B-spline densities. Both approaches fix the  number of B-spline densities. 
-The first approach uses equally spaced B-spline densities with a P-spline prior which we define in Section III.
-We demonstrate that this approach yields significant savings in computational time without sacrificing accuracy over the BspDP prior approach.
+The first approach uses equally spaced knots with a P-spline prior which we define in Section III.
+We demonstrate that this approach yields significant savings in computational time without sacrificing accuracy over the BspDP prior approach for simple spectral structures.
 The second approach has a particular  focus on spectral densities with sharp peaks.
@@
 where $\bm{v} = (v_1,\dots,v_{K-1})^\top$ is a $K-1$ dimensional parameter vector, $\phi$ is the smoothing or penalty parameter, $\textbf{P} = \textbf{D}^\top \textbf{D} +\epsilon \textbf{I}_{K-1}$ is the penalty matrix (which is a full matrix rank matrix for any small quantity $\epsilon$, for instance, $10^{-6}$) with $\textbf{D}$ the $d^{\text{th}}$ order difference matrix ($d=1,2,\ldots, K-2)$, $\alpha_{\phi}$ and $\alpha_{\delta}$ are shape parameters, $\delta \beta_{\phi}$ and $\beta_{\delta}$ are rate parameters, and $\text{Gamma}(a,b)$ denotes a gamma distribution with mean $a/b$ and variance $a/b^2$.
-We used first, second and third order difference penalties in the simulation and case study.
+We used first, second and third order difference penalties in the simulation study.
 The $1^{\text{st}}$ order difference matrix \textbf{D} is defined as
@@
 A larger penalty order does not necessarily result in more accurate psd estimates.
-In this case of a relatively smooth spectral density, the P-spaced knots do not yield an improvement over the equidistant knots but slightly better or similar accuracy than the BspDP prior.
+In this case of a relatively smooth spectral density, the P-spaced knots do not yield an improvement over the equidistant knots and the BspDP prior.
+
 % yields more accurate results than the B-spline prior for $n=128$.
@@
 It also outperforms the equidistant knot scheme.
-The latter yields more accurate results than the BspDP prior for $n=\{128, 256\}$.
+In general, the BspDP algorithm achieves higher accuracy than the P-spline method with equidistant knots.
+As shown by \cite{Edwards2018}, the BspDP prior yields better results than the BDP in terms of median IAE in this example.
+Consequently, our P-spline algorithm with P-spaced knots outperforms the BDP method as well.
+
+%The latter yields more accurate results than the BspDP prior for $n=\{128, 256\}$.
 For $n=512$, the third order penalty P-splines has a slightly higher median IAE, but overall, this approach achieves higher accuracy.
-As shown by \cite{Edwards2018}, the BspDP prior yields better results than the BDP in terms of median IAE in this example.  Consequently, our P-spline algorithm outperforms the BDP method as well.
+%As shown by \cite{Edwards2018}, the BspDP prior yields better results than the BDP in terms of median IAE in this example.  Consequently, our P-spline algorithm outperforms the BDP method as well.
 
@@
 			AR(1)     &  &  &  \\ 
-			B-spline &  1.257 & 0.738 & 0.605  \\
+			B-spline &  0.825 & 0.769 & 0.609  \\
 			\textit{Equidistant knots}& & &   \\
@@
 			AR(4)     &  &  &  \\ 
-			B-spline & 4.069 & 3.016 & 2.401   \\
+			B-spline & 3.019 & 2.250 & 1.832   \\
 			\textit{Equidistant knots}& & &   \\
@@
 	\includegraphics[scale=0.4,clip=true,angle=0]{ar1.pdf}
-	\caption{Estimated log-spectral density for an AR(1) time series via the P-spline prior algorithm.  The solid line stands for the true psd, whereas gray area is the uniform 90\% credible band.}
+	\caption{Estimated log-spectral density for an AR(1) time series via the P-spline prior algorithm with a first order penalty.
+The solid line stands for the true psd, whereas gray area is the uniform 90\% credible band.}
 	\label{fig:ar1}
@@
 	\includegraphics[scale=0.4,clip=true,angle=0]{ar4.pdf}
-	\caption{Estimated log-spectral density for an AR(4) time series via the P-spline prior algorithm.  The solid line stands for the true psd, whereas gray area is the uniform 90\% credible band.}
+	\caption{Estimated log-spectral density for an AR(4) time series via the P-spline prior algorithm with a first order penalty.
+The solid line stands for the true psd, whereas gray area is the uniform 90\% credible band.}
 	\label{fig:ar4}
@@
 
-The idea is to measure the proportion of times that the true spectral density is contained by the uniform credible band.
-This calculation is made for all 300 replications.
+The idea is to measure the proportion of times that the true spectral density is fully contained by the uniform credible band in the 300 replications.
 The results are displayed in Table~\ref{table:sim_covPbb}.
 
-For the AR(1) model, all 90\% credible bands cover the whole true spectral density function, i.e.\ for all frequencies $\lambda$.
-For the AR(4) model, the P-spline prior with P-spaced knots has the best coverage accross the different specifications.
-% Only one of its coverage probabilities is not equal to 1 (for $d=3$), but it is close.  
+For the AR(1) model, the BspDP and first order penalty P-spline 90\% credible bands cover the whole true spectral density function, i.e.\ for all frequencies $\lambda$, in all the 300 independent analyses.
+In general, the latter method reduces its performance as the penalty degree increases.
+For the AR(4) model, the P-spline method with equidistant knots performs poorly.
+On the other hand, the P-space knot scheme yields evidently much better results, which are comparable (in the case of the first order penalty) to those obtained via the BspDP algorithm.
+For $n=\{128,256\}$, the latter has higher coverage probabilities than our proposal, except for $n=512$.
+For the P-spline priors, the coverage probabilities decrease as the penalty degree order increases, which can be explained by the forced smoothness imposed by the higher penalty order.
+As shown by \cite{Edwards2018}, the BspDP algorithm produces better results than the BDP prior in the AR(4) case, which coverage probability is zero across $n=\{128,256,512\}$.
+Consequently, our P-spline algorithm with P-spaced knots outperforms the BDP method in this case.
+
+%For the AR(4) model, the P-spline prior with P-spaced knots has the best coverage across the different specifications.
 For the equidistant knots the coverage is always greater than 90\%.
@@
 For the P-spline priors, in particular for $n=512$, the coverage probabilities decrease as the penalty degree order increases, which can be explained by the forced smoothness imposed by the higher penalty order.
-In general, P-spline priors have a better coverage probability than the BspDP method which
-as shown by  \cite{Edwards2018} have better coverage probabilities than the BDP prior in the AR(4) case.
+In general, P-spline priors have a better coverage probability than the BspDP method which as shown by  \cite{Edwards2018} have better coverage probabilities than the BDP prior in the AR(4) case.
 
@@
 			B-spline &  1.000 & 1.000 & 1.000 \\
-			\textit{Equidistant \&}& & &   \\
-			\textit{P-spaced knots}& & &   \\
-			\hspace{0.5em}P-spline $d=1,2,3$ & 1.000 & 1.000 & 1.000  \\ \hline
-			%		P-spline $d=2$& 1.000 & 1.000 & 1.000 \\
-			%		P-spline $d=3$& 1.000 & 1.000 & 1.000 \\ \hline		       
-			AR(4)     &  &  &  \\ 
-			B-spline &  0.968 & 0.945 & 0.980  \\
-			\textit{Equidistant knots}& & &   \\
-			\hspace{0.5em}P-spline $d=1$ & 0.984 & 0.992 & 0.984  \\
-			\hspace{0.5em}P-spline $d=2$& 0.984 & 0.984 & 0.980 \\
-			\hspace{0.5em}P-spline $d=3$& 0.968 & 0.976 & 0.969  \\
+			\textit{Equidistant}& & &   \\
+			\hspace{0.5em}P-spline $d=1$& 1.000 & 1.000 & 1.000  \\
+			\hspace{0.5em}P-spline $d=2$& 0.987 & 0.993 & 1.000  \\
+			\hspace{0.5em}P-spline $d=3$& 0.943 & 0.877 & 0.870  \\
 			\textit{P-spaced knots}& & &   \\
 			\hspace{0.5em}P-spline $d=1$ & 1.000 & 1.000 & 1.000 \\
-			\hspace{0.5em}P-spline $d=2$ & 1.000 &  1.000 &  1.000 \\
-			\hspace{0.5em}P-spline $d=3$ & 0.984 &  1.000 &  1.000 \\	
+			\hspace{0.5em}P-spline $d=2$ & 0.993 & 0.980 & 0.990 \\
+			\hspace{0.5em}P-spline $d=3$ & 0.947 & 0.840 & 0.957 \\ \hline		       
+			AR(4)     &  &  &  \\ 
+			B-spline &  0.933 & 0.967 & 0.897  \\
+			\textit{Equidistant knots}& & &   \\
+			\hspace{0.5em}P-spline $d=1$ & 0.400 & 0.353 & 0.010 \\
+			\hspace{0.5em}P-spline $d=2$ & 0.200 & 0.167 & 0.003 \\
+			\hspace{0.5em}P-spline $d=3$ & 0.053 & 0.023 & 0.000 \\
+			\textit{P-spaced knots}& & &   \\	
+			\hspace{0.5em}P-spline $d=1$ & 0.820 & 0.887 & 0.920 \\
+			\hspace{0.5em}P-spline $d=2$ & 0.637 & 0.700 & 0.773 \\
+			\hspace{0.5em}P-spline $d=3$ & 0.470 & 0.600 & 0.583 \\	
 			\bottomrule
@@
 Note that the P-spline method based on equidistant knots has similar run-times to the P-spaced knots.
-The P-spline method is approximately 4-6 times faster than BspDP approach in these examples.
+The P-spline method is approximately 3-6 times faster than BspDP approach in these examples.
 This is due to a reduced number of calculations  per iteration because the B-spline densities are calculated only  once at the start of the algorithm and there are no Dirichlet process approximations via the stick-breaking method.
@@
 			AR(1)     &  &  &  \\ 
-			B-spline &   30.872 & 34.873 & 42.738 \\
+			B-spline &   29.497 & 32.423 & 37.896 \\
 			\textit{P-spaced knots}& & &   \\
@@
 			\hspace{0.5em}P-spline $d=3$& 5.587 & 8.301 & 10.427 \\ 
-			B-spline/P-spline &  5.491 & 4.190 & 4.077 \\ \hline
+			B-spline/P-spline &  5.247 & 3.896 & 3.615 \\ \hline
 			AR(4)     &  &  & \\ 
-			B-spline & 31.663 & 37.066 & 47.604  \\
+			B-spline & 30.111 & 34.126 & 40.875  \\
 			\textit{P-spaced knots}& & &   \\
@@
 			\hspace{0.5em}P-spline $d=3$& 5.611 & 8.327 & 10.686 \\
-			B-spline/P-spline & 5.630 & 4.459 & 4.506 \\	
+			B-spline/P-spline & 5.354 & 4.105 & 3.869 \\	
 			\bottomrule
@@
 
-Overall, the P-spline prior outperforms the B-spline prior in this example in terms of IAE, coverage probability and run-time.
+In general terms, the P-spline algorithm with our novel knot allocation scheme, the periodogram-spaced knots, offers the best trade-off between accuracy and computation cost.
+The  rule-of-thumb criterion,  $K=\min\left\{n/4, 40\right\}$, for the number of B-spline densities in the P-spline prior algorithm worked perfectly well in this example.
+
+%Overall, the P-spline prior outperforms the B-spline prior in this example in terms of IAE, coverage probability and run-time.
 Particularly, our novel knot allocation scheme, the periodogram-spaced knots, yield the best results in the case of peaked spectral densities.
@@
 Then, a Markov chain of 10,000 samples with a burn-in period of 2,000 and thinning factor of 10, that is 800 points, is used for the psd estimation.
-We use 40 B-spline densities,  based on the rule-of-thumb criterion discussed above, and the second order penalty.
+We use 40 B-spline densities,  based on the rule-of-thumb criterion discussed above, and the first order penalty.
 The whole process takes less than 2 minutes to run and is replicated for equidistant and P-spaced knots.
@@
 
-Althought the P-spline prior is very similar to the B-spline prior as both are based on mixtures of B-spline distributions but there is quite a significant difference.
+Although the P-spline prior is very similar to the B-spline prior as both are based on mixtures of B-spline distributions, there is quite a significant difference.
 The latter works with a variable number and location of knots, which are driven by the data.
 The P-spline prior assumes a fixed number of knots  with fixed locations, thus almost looks like a special case.
-However, the multivariate prior based on the difference penalty of the basis coefficients controls the smoothness of the spectral density, a feature that can't be achieved with univariate Dirichlet process prior.
-Most importantly,
- fixing the number and location of knots reduces the computational effort significantly, without sacrificing accuracy as shown in the application section.
+However, the multivariate prior based on the difference penalty of the basis coefficients controls the smoothness of the spectral density, a feature that can't be achieved with a univariate Dirichlet process prior.
+Most importantly, fixing the number and location of knots reduces the computational effort significantly, without sacrificing accuracy as shown in the application section.
 To deal with abrupt peaks in the psd, we propose to locate the knots according to the peaks detected via the periodogram.
@@
 
-In our simulation study, we showed that, in general, the P-spline prior outperforms the BspDP prior in all our assessments  in terms of IAE, coverage probabilities, and run-times.
-From the simulation study, it seems that a {\em second order} penalty P-spline works reasonably well.
+In our simulation study, we showed that the equidistant knot scheme for the P-spline method works quite well in the case of simple spectral structures (AR(1) model).
+However, its performance is surpassed by the P-spaced knot scheme for complex structures (AR(4) model).
+In this scenario particularly, we can highlight the good performance of the P-spaced knot scheme.
+Actually, in general, this method outperforms the BspDP prior in terms of IAE and run-times, particularly for the complex psds.
+However, the results were not conclusive for coverage probabilities.
+From this study, it seems that a {\em first order} penalty P-spline works reasonably well.
+
+%In our simulation study, we showed that, in general, the P-spline prior with P-spaced knots outperforms the BspDP prior in terms of IAE and run-times, particularly for the complex psd.
+However, the results were not conclusive for coverage probabilities.
+From this study, it seems that a {\em first order} penalty P-spline works reasonably well.
 
@@
 For this, we estimated the psd for the {\it S.\ Carinae} time series which has a spectral density with several sharp peaks.
-Equidistant knots failed to capture the peaks of this spectral density whereas  P-spaced knots
-significantly improved  the results without affecting the computational time.
+Equidistant knots failed to capture the peaks of this spectral density whereas  P-spaced knots significantly improved  the results without affecting the computational time.
 
 The number of B-spline densities was selected according to the criterion proposed by \cite{Ruppert2002}.
-It seems to work well, at least for the spectral densities that we experimented with.
+It seems to work well for simple spectral densities.
 However, more complex functions could require larger number of B-spline densities.
@@
 Our knot location proposal is very useful in these situations.
+Future research will explore whether the choice of the number of knots can be related to the entropy of the periodogram.
 
-In conclusion, the P-spline prior represents a more computationally viable alternative to the BspDP prior for spectral density estimation.
+In conclusion, the P-spline prior with the P-spaced knot scheme offers a more computationally viable alternative to the BspDP prior for spectral density estimation.
 As shown in this work, it can handle spectral densities with several peaks in a reasonable run-time.

### 2019-05-06 pSplines.tex
Renate edition

@@
 Instead of an equidistant spacing, the knots are spread based on the distribution of peaks in the periodogram.
-Thus the knot allocation is also data-driven as with the BspDP prior but the knots  remains fixed once
+Thus the knot allocation is also data-driven as with the BspDP prior but the knots  remain fixed once
 allocated.
@@
 In this scenario particularly, we can highlight the good performance of the P-spaced knot scheme.
-Actually, in general, this method outperforms the BspDP prior in terms of IAE and run-times, particularly for the complex psds.
-However, the results were not conclusive for coverage probabilities.
-From this study, it seems that a {\em first order} penalty P-spline works reasonably well.
+This method outperforms the BspDP prior in terms of IAE and run-times, particularly for the complex psds.
+Regarding uniform coverage,  a {\em first order} penalty P-spline coupled with
+P-spaced knots works best. For this combination, the uniform coverage almost reaches the nominal level.
 

### 2019-05-13 pSplines.tex
update table

@@
 Most Bayesian approaches to psd estimation use the so-called Whittle likelihood, introduced by \citep{Whittle:1957} as an approximation to the likelihood of a Gaussian stationary time series.
-Based on the  limiting independent exponential distribution of the periodogram, the Whittle likelihood for a mean centered weakly stationary time series $\bm{Y}$ of length $n$ is defined as
+Based on the  limiting independent exponential distribution of the periodogram, the Whittle likelihood for a mean centred weakly stationary time series $\bm{Y}$ of length $n$ is defined as
 \begin{align}
@@
 P-splines, i.e.\  equally-spaced B-splines,  have been successfully implemented for frequentist nonparametric regression but to the best of our knowledge not yet for Bayesian spectral density estimation.
-We suggest
-two novel Bayesian  approaches in this paper, both based on the Whittle likelihood and on a mixture of B-spline densities. Both approaches fix the  number of B-spline densities. 
+We suggest two novel Bayesian  approaches in this paper, both based on the Whittle likelihood and on a mixture of B-spline densities.
+Both approaches fix the  number of B-spline densities.
 The first approach uses equally spaced knots with a P-spline prior which we define in Section III.
@@
 Instead of an equidistant spacing, the knots are spread based on the distribution of peaks in the periodogram.
-Thus the knot allocation is also data-driven as with the BspDP prior but the knots  remain fixed once
-allocated.
+Thus the knot allocation is also data-driven as with the BspDP prior but the knots  remain fixed once allocated.
 This novel knot allocation scheme avoids the computational complexity of the BspDP prior.
@@
 We first assess the properties of our P-spline spectral density estimates in a simulation study and compare them to estimates based on the BspDP prior. We furthermore apply all approaches to the
-analysis of the classical sunspot dataset  and the {\it S.\ Carinae} variable star light intensities, previously analyzed by \cite{Cart:1997,Huerta:1999,Kirch:2018}.
+analysis of the classical sunspot dataset  and the {\it S.\ Carinae} variable star light intensities, previously analysed by \cite{Cart:1997,Huerta:1999,Kirch:2018}.
 As the psd of the  {\it S.\ Carinae} time series contains sharp peaks, its analysis allows to assess the impact of the periodogram-spaced knots in contrast to the equidistant knots.
@@
 	\caption{Estimated log-spectral density for an AR(1) time series via the P-spline prior algorithm with a first order penalty.
-The solid line stands for the true psd, whereas gray area is the uniform 90\% credible band.}
+The solid line stands for the true psd, whereas grey area is the uniform 90\% credible band.}
 	\label{fig:ar1}
@@
 	\caption{Estimated log-spectral density for an AR(4) time series via the P-spline prior algorithm with a first order penalty.
-The solid line stands for the true psd, whereas gray area is the uniform 90\% credible band.}
+The solid line stands for the true psd, whereas grey area is the uniform 90\% credible band.}
 	\label{fig:ar4}
@@
 The results are displayed in Table~\ref{table:sim_covPbb}.
+In addition, we calculate the proportion of times that the uniform credible band contains the value of the true psd
+at each of the Fourier frequencies.  The median of these pointwise coverage proportions are displayed in Table~\ref{table:sim_covPbb_median}.
 
 For the AR(1) model, the BspDP and first order penalty P-spline 90\% credible bands cover the whole true spectral density function, i.e.\ for all frequencies $\lambda$, in all the 300 independent analyses.
-In general, the latter method reduces its performance as the penalty degree increases.
+In general, the latter method shows diminishing  performance as the penalty degree increases.
 For the AR(4) model, the P-spline method with equidistant knots performs poorly.
-On the other hand, the P-space knot scheme yields evidently much better results, which are comparable (in the case of the first order penalty) to those obtained via the BspDP algorithm.
+On the other hand, the P-spaced knot scheme yields evidently much better results, which are comparable (in the case of the first order penalty) to those obtained via the BspDP algorithm.
 For $n=\{128,256\}$, the latter has higher coverage probabilities than our proposal, except for $n=512$.
@@
 Consequently, our P-spline algorithm with P-spaced knots outperforms the BDP method in this case.
+
+Even though the P-spline method with equidistant knots for the AR(4) case performs poorly in terms of uniform coverage, i.e.
+the proportion of times that the credible band covers  the true psd {\em completely} in the 300 runs, it is important to note that this is only because the uniform band does not cover
+the psd for  just very few frequencies. As is evident from 
+  Table~\ref{table:sim_covPbb_median},  only for a small fraction of the frequencies the psd is not covered by the uniform credible band.
+In this context, BspDP and P-spaced knot P-spline algorithms perform equally well.
 
@@
 %	\end{ruledtabular}
-	\caption{Coverage probabilities of $90\%$ uniform credible bands.}
+	\caption{Proportion of times that the true psd is entirely encapsulated within the 90\% uniform credible band of the 300 estimates.}
 	\label{table:sim_covPbb}
+\end{table}
+
+\begin{table}
+	\centering
+		\begin{tabular}{lccc}
+			\toprule
+			& $n=128$ & $n=256$ & $n=512$ \\ \hline
+			AR(1)     &  &  &  \\ 
+			B-spline &  1.000 & 1.000 & 1.000 \\
+			\textit{Equidistant \&}& & &   \\
+			\textit{P-spaced knots}& & &   \\
+			\hspace{0.5em}P-spline $d=1,2,3$ & 1.000 & 1.000 & 1.000  \\ \hline	 		      
+			AR(4)     &  &  &  \\ 
+			B-spline &  1.000 & 1.000 & 1.000 \\
+			\textit{Equidistant knots}& & &   \\
+			\hspace{0.5em}P-spline $d=1$ & 0.984 & 0.992 & 0.984 \\
+			\hspace{0.5em}P-spline $d=2$& 0.984 & 0.984 & 0.980  \\
+			\hspace{0.5em}P-spline $d=3$& 0.968 & 0.976 & 0.969  \\	
+			\textit{P-spaced knots}& & &   \\	
+			\hspace{0.5em}P-spline $d=1$ & 1.000 &  1.000 &  1.000 \\
+			\hspace{0.5em}P-spline $d=2$ & 1.000 &  1.000 &  1.000 \\
+			\hspace{0.5em}P-spline $d=3$ & 0.984 &  1.000 &  1.000 \\		
+			\bottomrule
+		\end{tabular}
+	\caption{Median of the pointwise  proportions that the true psd at each Fourier frequency is contained within the $90\%$ uniform credible bands.}
+	\label{table:sim_covPbb_median}
 \end{table}
@@
 	\caption{Log-spectral density estimate for the sunspot data.
-The continuous gray line represents the periodogram.
+The continuous grey line represents the periodogram.
 The continuous and dashed lines stand for the posterior median log psd obtained using the B-spline and P-spline priors, respectively.
@@
 	\caption{Log-spectral density estimate for the variable star {\it S.\ Carinae} data based on the P-spline prior.
-The continuous gray line represents the periodogram whereas the continuous and dashed black lines stand for the posterior median log psd obtained using P-spaced and equidistant knots, respectively.}
+The continuous grey line represents the periodogram whereas the continuous and dashed black lines stand for the posterior median log psd obtained using P-spaced and equidistant knots, respectively.}
 	\label{fig:carinae}

### 2019-10-30 pSplines.tex
Update

@@
 Unlike the beta densities  which have global support on the interval [0,1], B-spline densities have local support, thereby increasing their flexibility and allowing better modelling of spectral densities with sharp peaks and abrupt changes.
-As demonstrated in \citep{Edwards2018}, the B-spline--Dirichlet process (BspDP) prior outperforms the BDP prior in estimating complex psd's.
+As demonstrated in \cite{Edwards2018}, the B-spline--Dirichlet process (BspDP) prior outperforms the BDP prior in estimating complex psd's.
 However, this flexibility comes at a high computational price.
@@
 
-With the aim of reducing the computational cost of the BspDP prior approach,  we suggest to use the idea of P-splines \cite{Eilers:1996} that combine a large but fixed number of  B-splines with a simple difference penalty that controls the degree of  smoothness of the spectral density.
+With the aim of reducing the computational cost of the BspDP prior approach,  we suggest to use the idea of P-splines of \cite{Eilers:1996} that combine a large but fixed number of  B-splines with a simple difference penalty that controls the degree of  smoothness of the spectral density.
 P-splines, i.e.\  equally-spaced B-splines,  have been successfully implemented for frequentist nonparametric regression but to the best of our knowledge not yet for Bayesian spectral density estimation.
@@
 \end{align*}	  
-converges uniformly to $G(\omega)$, where $G(u,v] = G(v) - G(u)$ and $\beta(\omega; a,b)$ is the beta density with parameters $a$ and $b$.
+converges uniformly to $G(\omega)$, where $G(u,v] = G(v) - G(u)$, $\beta(\omega; a,b)$ is the beta density with parameters $a$ and $b$, and $w_k=G\left((k-1)/K, k/K \right]$ are the weights of the mixture.
 The Bernstein polynomial prior, which is used to describe a nonparametric prior for probability densities in the unit interval, is based on this approximation \citep{Petrone:1999a,Petrone:199b}.
@@
 Note that even though the proposal is univariate on $\bm{\beta}$, it is multivariate on $\bm{v}^*$.
+Alternatively to the Metropolis step used for \bm{v}, a data augmentation step based on Polya-Gamma auxiliary variables can be used which would allow for a full Gibbs sampler \citep{Polson:2013}.
 
@@
 \end{enumerate}
+
+The P-spaced knot allocation can be regarded as an empirical Bayes prior as the data are used to specify the prior distribution.
+However, only the locations of the B-splines are determined by the data, not the parameters of the prior distribution which in this case are the weights $w_k$ associated with each B-spline $b_{k,r}$.
+The main criticism of an empirical Bayes approach is that the data are being used twice, once to specify the prior and then again to update the prior to the posterior.
+Here, the data are used only once to specify the knot locations but not again to update these through the likelihood.
 

### 2020-03-23 pSplines.tex
Updated figures
Figures have been updated + carinae data

@@
 % Useful packages
-\usepackage{graphicx}  % Package for importing postscript images
+\usepackage[pdftex]{graphicx}  % Package for importing postscript images
+\usepackage{color}
 \usepackage[authoryear, round, numbers]{natbib}  % Package for Bibtex
@@
 \usepackage{amsfonts}
+\usepackage{verbatim}
 
@@
 
-\title{Spectral density estimation using P-spline priors}
+\title{Bayesian spectral density estimation using P-spline priors}
 \author{Patricio Maturana-Russel \and Renate Meyer}
@@
 }
+
+\newcommand{\pmr}{ \color{blue}}
 
@@
 
-This article proposes a Bayesian approach to estimating the spectral density of a stationary time series using a prior based on a mixture of P-spline distributions.
-Our proposal is motivated by the B-spline Dirichlet process prior of \cite{Edwards2018} in combination  with Whittle's likelihood and aims at  reducing the high computational complexity of its posterior computations.
+This article  a Bayesian approach to estimating the spectral density of a stationary time series using a prior based on a mixture of P-spline distributions.
+Our proposal is motivated by the B-spline Dirichlet process prior of \cite{Edwards2019} in combination  with Whittle's likelihood and aims at  reducing the high computational complexity of its posterior computations.
 The strength of the B-spline Dirichlet process prior over the Bernstein-Dirichlet process prior of \cite{Choudhuri:2004}  lies in its ability to estimate spectral densities with sharp peaks and abrupt changes due to the flexibility of B-splines with variable number and location of knots.
@@
 
-\keywords{P-spline prior \and B-spline prior \and spectral density estimation \and Whittle likelihood}
+\keywords{P-splines \and B-splines \and \textcolor{red}{Bernstein-Dirichlet process prior}  \and spectral density estimation \and Whittle likelihood}
 
@@
 in one way or another, where $-\pi < \lambda \leq \pi$ is the angular frequency and $\bm{Y}=(Y_1,\ldots,Y_n)$ is a stationary time series of length $n$.
-The periodogram has interesting features, for instance, it is asymptotically unbiased and follows an exponential distribution with mean $f(\lambda)$.
-However, it is not consistent \citep{Brockwell:1986}.
+The periodogram has interesting features \textcolor{red}{(***)?why is that here?}.
+For instance, the periodogram evaluated at the Fourier frequencies $\lambda_i$  are asymptotically independent and follow an exponential distribution with mean $f(\lambda_i)$.
+Thus, they are unbiased estimates of $f(\lambda_i)$  but not consistent \citep{Brockwell:1986}.
 
 Most Bayesian approaches to psd estimation use the so-called Whittle likelihood, introduced by \citep{Whittle:1957} as an approximation to the likelihood of a Gaussian stationary time series.
-Based on the  limiting independent exponential distribution of the periodogram, the Whittle likelihood for a mean centred weakly stationary time series $\bm{Y}$ of length $n$ is defined as
+Based on the  limiting independent exponential distribution of the periodogram \textcolor{red}{values}, the Whittle likelihood for a mean centred weakly stationary time series $\bm{Y}$ of length $n$ is defined as
 \begin{align}
@@
 where $\lambda_l = 2\pi l / n$ are the Fourier frequencies, $f(\cdot)$ spectral density of $\bm{Y}$, and $\nu = \left\lfloor (n-1)/2 \right\rfloor$ is the greatest integer value less than or equal to $(n-1)/2$.
-Advantages of this approximation is that it depends directly on the spectral density, unlike the true Gaussian likelihood function, and gives a good approximation for non-Gaussian time series under certain conditions as shown by \cite{ShaoXiaofeng2007ASTf}.
+Advantages of this approximation \textcolor{red}{are} that it depends directly on the spectral density, unlike the true Gaussian likelihood function, and gives a good approximation for non-Gaussian time series under certain conditions as shown by \cite{ShaoXiaofeng2007ASTf}.
 
@@
 A reversible jump Markov chain Monte Carlo \citep[RJMCMC;][]{Green:1995} algorithm was implemented to sample from the posterior distribution.
+\textcolor{red}{Similarly, frequentist approaches such as \cite{Rodriguez} and \cite{Wood2017} model the log spectral density by a linear
+combination of splines with variable knots and combine these with adaptive smoothing. }
 
+\textcolor{red}{Penalized splines have been used for a time-domain analysis of time series data e.g.\ to model the mean as a smooth function of time by
+\cite{Krivobokova}  or for fitting and  forecasting univariate  nonlinear time series by \cite{Wegener}, but not in the frequency domain for estimating the spectral density.}
 The P-spline prior on the spectral density function that we are going to propose is related to the 
@@
 The number of components, which is a smoothing parameter, is given a discrete prior.
-\cite{Edwards2018} extended the method by replacing the beta densities with B-spline densities.
+\cite{Edwards2019} extended the method by replacing the beta densities with B-spline densities.
 Similar to
@@
 Unlike the beta densities  which have global support on the interval [0,1], B-spline densities have local support, thereby increasing their flexibility and allowing better modelling of spectral densities with sharp peaks and abrupt changes.
-As demonstrated in \cite{Edwards2018}, the B-spline--Dirichlet process (BspDP) prior outperforms the BDP prior in estimating complex psd's.
+As demonstrated in \cite{Edwards2019}, the B-spline--Dirichlet process (BspDP) prior outperforms the BDP prior in estimating complex psd's.
 However, this flexibility comes at a high computational price.
@@
 
-With the aim of reducing the computational cost of the BspDP prior approach,  we suggest to use the idea of P-splines of \cite{Eilers:1996} that combine a large but fixed number of  B-splines with a simple difference penalty that controls the degree of  smoothness of the spectral density.
+%<<<<<<< HEAD
+With the aim of reducing the computational cost of the BspDP prior approach,  we suggest to use the idea of P-splines of \cite{Eilers:1996} that \textcolor{red}{combines} a large but fixed number of  B-splines with a simple difference penalty that controls the degree of  smoothness of the spectral density.
 P-splines, i.e.\  equally-spaced B-splines,  have been successfully implemented for frequentist nonparametric regression but to the best of our knowledge not yet for Bayesian spectral density estimation.
@@
 We show that with this knot allocation scheme, the estimates of spectral densities with spikes and abrupt changes improves significantly without an increase in computing time.
+=======
+With the aim of reducing the computational cost of the BspDP prior approach,  we suggest to use the idea of P-splines \cite{Eilers:1996} that combine a large but fixed number of  B-splines with a simple difference penalty that controls the degree of  smoothness of the spectral density.
+P-splines, i.e.\  equally-spaced B-splines,  have been successfully implemented for frequentist nonparametric regression but to the best of our knowledge not yet for Bayesian spectral density estimation.
+We suggest
+two novel Bayesian  approaches in this paper, both based on the Whittle likelihood and on a mixture of B-spline densities. Both approaches fix the  number of B-spline densities. 
+The first approach uses equally spaced knots with a P-spline prior which we define in Section III.
+We demonstrate that this approach yields significant savings in computational time without sacrificing accuracy over the BspDP prior approach for simple spectral structures.
+The second approach has a particular  focus on spectral densities with sharp peaks.
+Instead of an equidistant spacing, the knots are spread based on the distribution of peaks in the periodogram.
+Thus the knot allocation is also data-driven as with the BspDP prior but the knots  remain fixed once
+allocated.
+This novel knot allocation scheme avoids the computational complexity of the BspDP prior.
+We show that with this knot allocation scheme, the estimates of spectral densities with spikes and abrupt changes improves significantly without an increase in computing time.
+%>>>>>>> 19c7083b8cf09264b1cfb0f5ba98e32bf3de8990
 
@@
 Let $G$ be a cumulative distribution function (cdf), with continuous density $g$, then
-\begin{align*}
-\widehat{G}(\omega) &= \sum_{k=1}^{K} G \left( \dfrac{k-1}{K} , \dfrac{k}{K} \right] \beta(\omega; k, K-k+1)\\
+\begin{align}
+\label{eq:Bernstein}
+\widehat{g}(\omega) &= \sum_{k=1}^{K} G \left( \dfrac{k-1}{K} , \dfrac{k}{K} \right] \beta(\omega; k, K-k+1)\\
 &= \sum_{k=1}^{K} w_k \beta(\omega; k, K-k+1)
-\end{align*}	  
-converges uniformly to $G(\omega)$, where $G(u,v] = G(v) - G(u)$, $\beta(\omega; a,b)$ is the beta density with parameters $a$ and $b$, and $w_k=G\left((k-1)/K, k/K \right]$ are the weights of the mixture.
+\end{align}	  
+converges uniformly to $g(\omega)$, where $G(u,v] = G(v) - G(u)$, $\beta(\omega; a,b)$ is the beta density with parameters $a$ and $b$, and $w_k=G\left((k-1)/K, k/K \right]$ are the weights of the mixture.
 The Bernstein polynomial prior, which is used to describe a nonparametric prior for probability densities in the unit interval, is based on this approximation \citep{Petrone:1999a,Petrone:199b}.
@@
 The adequacy of the approximation is a function of the number of knots.
-This work motivated \cite{Edwards2018} to propose the BspDP prior on the spectral density function.
+This work motivated \cite{Edwards2019} to propose the BspDP prior on the spectral density function which is based on the representation (\ref{eq:Bernstein}) but with
+the beta densities replaced by B-spline densities as defined in the following.
 
-\subsection*{B-splines}
+\subsection*{B-splines and BspDP Prior}
 
@@
 
+\begin{comment}
 \subsection*{BspDP Prior}
@@
 This approach avoids the potential difficulties caused by  jumps between the parameter spaces of different dimensions in  RJMCMC.
-\cite{Edwards2018} followed this approach, but in addition they assumed that the $K-r$ internal knot differences are induced by a cdf $H$ on $[0,1]$ with further Dirichlet process prior, that is
+\cite{Edwards2019} followed this approach, but in addition they assumed that the $K-r$ internal knot differences are induced by a cdf $H$ on $[0,1]$ with further Dirichlet process prior, that is
 \begin{align*}
@@
 The BspDP prior is a mixture of B-spline densities with local support, unlike the BDP prior which uses beta densities with  support on the whole unit interval.
-This feature allows the B-spline method more flexibility for spectral density estimation \citep{Edwards2018}.
+This feature allows the B-spline method more flexibility for spectral density estimation \citep{Edwards2019}.
 
@@
 %\subsection*{Implementation discussion}
+\end{comment}
 
-The BspDP prior proposed by \cite{Edwards2018} involves two  Dirichlet processes, one  placed on $G$ inducing  the weights of the B-spline densities and the other on  $H$ for the knot locations.
-In practice, these are implemented using   the  stick-breaking representation \citep{Sethuraman:1994}, for details see \cite{Edwards2018}.
+The BspDP prior proposed by \cite{Edwards2019}
+ involves two  Dirichlet processes, one  placed on $G$ inducing  the weights of the B-spline densities and the other on the knot spacings.
+In practice, these are implemented using   the  stick-breaking representation \citep{Sethuraman:1994}, for details see \cite{Edwards2019}.
 % This method consists basically in a representation of the Dirichlet process as a sequential division of the unit interval infinite times according to a beta distribution. 
-For computational reasons, the number of components of this infinite series representation of $G$ and $H$ is limited to a large but finite positive integer, that is $L_G$ and $L_H$, respectively.
-These parameters control the quality of the approximations.
-Larger values improve the approximation, but increase the number of calculations and consequently the computation time.
+For computational reasons,  this infinite series representation of $G$ and $H$ is truncated at a large but finite positive integer, $L_G$ and $L_H$, respectively.  
+These truncations control the quality of the approximations.  Longer series improve the approximation, but increase the number of calculations and consequently the computation time.
 
-As shown by \cite{Edwards2018}, the approach based on the  BspDP prior outperforms the BDP prior in estimating spiky spectral densities.  In the case of smooth
+As shown by \cite{Edwards2019}, the approach based on the  BspDP prior outperforms the BDP prior in estimating spiky spectral densities.  In the case of smooth
 spectral densities, there is no a significant difference in the performance of these two approaches.
@@
 \section{P-spline prior}
+\textcolor{red}{
+As the spectral density function $f(\cdot)$ is defined on the interval $[0,\pi]$, it is reparametrized and defined as
+\begin{align*}
+f(\pi \omega) =\tau \times s_r(\omega;\bm{w}, \bm{\xi}), \quad \omega \in [0,1], %\tau \times s_r(\omega;K,G,H), \quad \omega \in [0,1],
+\end{align*}
+where $\tau = \int_{0}^{1}f(\tau \omega)\text{d}\omega$ is the normalizing constant and
+\begin{align}
+s_r(\omega; \bm{w}, \bm{\xi}) = \sum_{k=1}^{K} w_k b_{k,r}(\omega;\bm{\xi}),
+\label{eq:ps_prior}	
+\end{align}
+with  {\em fixed} number $K$ of B-spline densities  $b_{k,r}(\cdot)$ of fixed degree $r$, weight vector $\bm{w}=(w_1,\cdots, w_K)$, and knot sequence $\bm{\xi}$.}
 
-Fixing the number of B-spline densities $K$ in Eq. \eqref{eq:bs_prior}, the prior for the unit interval $[0,1]$ is defined as
-\begin{align*}
-s_r(\omega;\bm{w}, \bm{\xi}) = \sum_{k=1}^{K} w_{k} b_{k,r}(\omega;\bm{\xi}),	
-\end{align*}
-where $\bm{w} = (w_1,\dots,w_K)$.
+%Fixing the number of B-spline densities $K$ in Eq. \eqref{eq:bs_prior}, the prior for the unit interval $[0,1]$ is defined as
+%\begin{align*}
+%s_r(\omega;\bm{w}, \bm{\xi}) = \sum_{k=1}^{K} w_{k} b_{k,r}(\omega;\bm{\xi}),	
+%\end{align*}
+%where $\bm{w} = (w_1,\dots,w_K)$.
 
-$\bm{\xi}$ contains fixed equidistant internal knots on $[0,1]$ in our basic P-spline prior. A different  allotment of knots is described in  section ``knot allocation strategy".
-Therefore, the P-spline prior on the spectral density function is given by
-\begin{align*}
-f(\pi \omega) = \tau \times s_r(\omega;\bm{w}, \bm{\xi}), \quad \omega \in [0,1],
-\end{align*}
-where $\tau = \int_{0}^{1}f(\tau \omega)\text{d}\omega$ is the normalizing constant.
+$\bm{\xi}$ contains fixed equidistant internal knots on $[0,1]$ in our basic P-spline prior. A different  allotment of knots is described in  section ``\textcolor{red}{Adaptive knots}".
+%Therefore, the P-spline prior on the spectral density function is given by
+%\begin{align*}
+%f(\pi \omega) = \tau \times s_r(\omega;\bm{w}, \bm{\xi}), \quad \omega \in [0,1],
+%\end{align*}
+%where $\tau = \int_{0}^{1}f(\tau \omega)\text{d}\omega$ is the normalizing constant.
 
@@
 \begin{align*}
-\bm{v}&\sim\text{N}_{K-1}(\bm{0}, (\phi \bm{P})^{-1})\\
+\bm{v}|\phi&\sim\text{N}_{K-1}(\bm{0}, (\phi \bm{P})^{-1})\\
 \phi|\delta &\sim \text{Gamma}(\alpha_{\phi}, \delta \beta_{\phi})\\
@@
 We use a robust specification for the  prior distribution of the penalty parameters as suggested by \cite{Jullion:2007} by choosing small values for $\alpha_{\delta}$ and $\beta_{\delta}$, for instance, $10^{-4}$.
-The the choice of $\alpha_{\phi}$ and $\beta_{\phi}$ do not affect the spectral density estimate.
+The  choice of $\alpha_{\phi}$ and $\beta_{\phi}$ do not affect the spectral density estimate.
 Here we set $\alpha_{\phi}$ and $\beta_{\phi}$ equal to 1 as used by \cite{Bremhorst:2016}.
@@
 Note that even though the proposal is univariate on $\bm{\beta}$, it is multivariate on $\bm{v}^*$.
-Alternatively to the Metropolis step used for \bm{v}, a data augmentation step based on Polya-Gamma auxiliary variables can be used which would allow for a full Gibbs sampler \citep{Polson:2013}.
+Alternatively to the Metropolis step used for \bm{v}, a data augmentation step based on P\'olya-Gamma auxiliary variables can be used which would allow for a full Gibbs sampler \citep{Polson:2013}.
 
@@
 
-\subsection*{Knot allocation strategy}
+%\subsection*{Knot allocation strategy}
+\subsection*{Adaptive knots}
 
 The knots are often equally spaced in P-spline methods.
-This scheme works quite well in the P-spline prior algorithm for smooth psds without any abrupt changes.
+This scheme works quite well  for smooth psds without any abrupt changes.
 However, for spectral densities with spikes, the algorithm will require a large number of knots to adequately estimate the peaks.
 The B-spline prior method is able to handle abrupt changes because the knot location is variable and  driven by the data.
-We propose a method that allocates the knots according to the periodogram but once allocated, the knots stay fixed.
+We propose \textcolor{red}{an adaptive} method that allocates the knots according to the periodogram but once allocated, the knots stay fixed.
 The idea is to concentrate more effort in those regions that have potential peaks as detected by the periodogram.
 
-Our knot location proposal works as follow.
+Our\textcolor{red}{adaptive} knot location proposal works as follow.
 We calculate the periodogram and apply a square root transformation in order to have more regular magnitudes and eliminate a potential trend.
@@
 Thus, in those areas in which the periodogram has sharp and abrupt changes, our procedure will assign more knots in proportion to their magnitudes.
-Hereafter, we will refer to this scheme as \textit{periodogram-spaced knots} or short \textit{P-spaced knots}.
+Hereafter, we will refer to this scheme as \textcolor{red}{\em adaptive knots}.%\textit{periodogram-spaced knots} or short \textit{P-spaced knots}.
 %REMARK: HOW ABOUT "PERIODOGRAM-SPACED KNOTS"	OR 'P-SPACED KNOTS'  INSTEAD OF DISTRIBUTED KNOTS?
@@
 	\item Let $\bm{\xi}^* = \left(\xi_1^*,\dots,\xi_{K^*}^* \right)$ be a vector of $K^*$ equally spaced values in $[0,1]$.
-	\item The P-spaced knot vector is given by $\bm{\xi} = \left(F_{z}^{-1}(\xi^*_1),\dots, F_{z}^{-1}(\xi^*_{K^*}) \right)$.
+	\item The \textcolor{red}{adaptive} knot vector is given by $\bm{\xi} = \left(F_{z}^{-1}(\xi^*_1),\dots, F_{z}^{-1}(\xi^*_{K^*}) \right)$.
 
@@
 
-The P-spaced knot allocation can be regarded as an empirical Bayes prior as the data are used to specify the prior distribution.
+The \textcolor{red}{adaptive}  knot allocation can be regarded as an empirical Bayes prior as the data are used to specify the prior distribution.
 However, only the locations of the B-splines are determined by the data, not the parameters of the prior distribution which in this case are the weights $w_k$ associated with each B-spline $b_{k,r}$.
@@
 
-The setup of our simulation study mirrors that  in \cite{Edwards2018} who  compared spectral density estimates based on the BspDP and the BDP priors. 
+The setup of our simulation study mirrors that  in \cite{Edwards2019} who  compared spectral density estimates based on the BspDP and the BDP priors. 
 % We perform the same analysis, varying only the number of iterations used in the analysis.  
@@
 %Therefore, the number of iterations equate the B-spline analysis.  
-This procedure was replicated for both equidistant and P-spaced knots.
+This procedure was replicated for both equidistant and \textcolor{red}{adaptive} knots.
 
@@
 A larger penalty order does not necessarily result in more accurate psd estimates.
-In this case of a relatively smooth spectral density, the P-spaced knots do not yield an improvement over the equidistant knots and the BspDP prior.
+In this case of a relatively smooth spectral density, the \textcolor{red}{adaptive} knots do not yield an improvement over the equidistant knots and the BspDP prior.
 
@@
 
-For the AR(4) time series, the P-spline prior with P-spaced knots categorically outperforms the BspDP algorithm.
-It also outperforms the equidistant knot scheme.
+For the AR(4) time series, the \textcolor{red}{fit based on the P-spline prior with \textcolor{red}{adaptive} knots is better than the fit based on the B-spline Dirichlet process prior and the one based on the P-spline prior with equidistant knots}.
+% categorically outperforms the BspDP algorithm.   It also outperforms the equidistant knot scheme. 
  In general, the BspDP algorithm achieves higher accuracy than the P-spline method with equidistant knots.
-As shown by \cite{Edwards2018}, the BspDP prior yields better results than the BDP in terms of median IAE in this example.
-Consequently, our P-spline algorithm with P-spaced knots outperforms the BDP method as well.
+As shown by \cite{Edwards2019}, the \textcolor{red}{posterior distribution based on the BspDP prior} yields better results than \textcolor{red}{posterior based on } the BDP in terms of median IAE in this example.
+Consequently, our P-spline algorithm with \textcolor{red}{adaptive} knots outperforms the BDP method as well.
 
@@
 			\hspace{0.5em}P-spline $d=3$&  0.727 & 0.728 & 0.769  \\
-			\textit{P-spaced knots}& & &   \\
+			\textit{\textcolor{red}{Adaptive} knots}& & &   \\
 			\hspace{0.5em}P-spline $d=1$ & 0.910 & 0.769 & 0.595 \\
@@
 			\hspace{0.5em}P-spline $d=3$& 3.438 & 2.786 & 2.510  \\	
-			\textit{P-spaced knots}& & &   \\
+			\textit{\textcolor{red}{Adaptive} knots}& & &   \\
 			\hspace{0.5em}P-spline $d=1$ & 2.659 & 1.884 & 1.465 \\
@@
 For the AR(4) model, the P-spline method with equidistant knots performs poorly.
-On the other hand, the P-spaced knot scheme yields evidently much better results, which are comparable (in the case of the first order penalty) to those obtained via the BspDP algorithm.
+On the other hand, the \textcolor{red}{adaptive} knot scheme yields evidently much better results, which are comparable (in the case of the first order penalty) to those obtained via the BspDP algorithm.
 For $n=\{128,256\}$, the latter has higher coverage probabilities than our proposal, except for $n=512$.
 For the P-spline priors, the coverage probabilities decrease as the penalty degree order increases, which can be explained by the forced smoothness imposed by the higher penalty order.
-As shown by \cite{Edwards2018}, the BspDP algorithm produces better results than the BDP prior in the AR(4) case, which coverage probability is zero across $n=\{128,256,512\}$.
-Consequently, our P-spline algorithm with P-spaced knots outperforms the BDP method in this case.
+As shown by \cite{Edwards2019}, the BspDP algorithm produces better results than the BDP prior in the AR(4) case, which coverage probability is zero across $n=\{128,256,512\}$.
+Consequently, our P-spline algorithm with \textcolor{red}{adaptive} knots outperforms the BDP method in this case.
 
@@
   Table~\ref{table:sim_covPbb_median},  only for a small fraction of the frequencies the psd is not covered by the uniform credible band.
-In this context, BspDP and P-spaced knot P-spline algorithms perform equally well.
+In this context, BspDP and \textcolor{red}{adaptive} knot P-spline algorithms perform equally well.
 
@@
 For the P-spline priors, in particular for $n=512$, the coverage probabilities decrease as the penalty degree order increases, which can be explained by the forced smoothness imposed by the higher penalty order.
-In general, P-spline priors have a better coverage probability than the BspDP method which as shown by  \cite{Edwards2018} have better coverage probabilities than the BDP prior in the AR(4) case.
+In general, P-spline priors have a better coverage probability than the BspDP method which as shown by  \cite{Edwards2019} have better coverage probabilities than the BDP prior in the AR(4) case.
 
@@
 			\hspace{0.5em}P-spline $d=3$& 0.943 & 0.877 & 0.870  \\
-			\textit{P-spaced knots}& & &   \\
+			\textit{\textcolor{red}{Adaptive} knots}& & &   \\
 			\hspace{0.5em}P-spline $d=1$ & 1.000 & 1.000 & 1.000 \\
@@
 			\hspace{0.5em}P-spline $d=3$ & 0.053 & 0.023 & 0.000 \\
-			\textit{P-spaced knots}& & &   \\	
+			\textit{\textcolor{red}{Adaptive} knots}& & &   \\	
 			\hspace{0.5em}P-spline $d=1$ & 0.820 & 0.887 & 0.920 \\
@@
 			\textit{Equidistant \&}& & &   \\
-			\textit{P-spaced knots}& & &   \\
+			\textit{\textcolor{red}{Adaptive} knots}& & &   \\
 			\hspace{0.5em}P-spline $d=1,2,3$ & 1.000 & 1.000 & 1.000  \\ \hline	 		      
@@
 			\hspace{0.5em}P-spline $d=3$& 0.968 & 0.976 & 0.969  \\	
-			\textit{P-spaced knots}& & &   \\	
+			\textit{\textcolor{red}{Adaptive} knots}& & &   \\	
 			\hspace{0.5em}P-spline $d=1$ & 1.000 &  1.000 &  1.000 \\
@@
 
-As \cite{Edwards2018} noted, one of the drawbacks of the BspDP prior  is its computational complexity relative to the BDP.
-This is directly reflected in the run-time, which is approximately 2-3 times higher in this example (see Table~3 in \cite{Edwards2018}).
-The median run-time in our analyses for the BspDP prior and the P-spline with P-spaced knots are displayed in Table~\ref{table:sim_time}.
-Note that the P-spline method based on equidistant knots has similar run-times to the P-spaced knots.
+As \cite{Edwards2019} noted, one of the drawbacks of the BspDP prior  is its computational complexity relative to the BDP.
+This is directly reflected in the run-time, which is approximately 2-3 times higher in this example (see Table~3 in \cite{Edwards2019}).
+The median run-time in our analyses for the BspDP prior and the P-spline with \textcolor{red}{adaptive} knots are displayed in Table~\ref{table:sim_time}.
+Note that the P-spline method based on equidistant knots has similar run-times to the \textcolor{red}{adapative} knots.
 The P-spline method is approximately 3-6 times faster than BspDP approach in these examples.
@@
 			B-spline &   29.497 & 32.423 & 37.896 \\
-			\textit{P-spaced knots}& & &   \\
+			\textit{\textcolor{red}{Adaptive} knots}& & &   \\
 			\hspace{0.5em}P-spline $d=1$& 5.629 & 8.291 & 10.507 \\
@@
 			B-spline & 30.111 & 34.126 & 40.875  \\
-			\textit{P-spaced knots}& & &   \\
+			\textit{\textcolor{red}{Adaptive} knots}& & &   \\
 			\hspace{0.5em}P-spline $d=1$ & 5.621 & 8.240 & 10.462 \\
@@
 Particularly, our novel knot allocation scheme, the periodogram-spaced knots, yield the best results in the case of peaked spectral densities.
-The  rule-of-thumb criterion,  $K=\min\left\{n/4, 40\right\}$, for the number of B-spline densities in the P-spline prior algorithm worked perfectly well in this example.
+The  rule-of-thumb criterion,  $K=\min\left\{n/4, 40\right\}$, for the number of
+\textcolor{red}{P-splines} %B-spline densities in the P-spline prior algorithm
+ worked perfectly well in this example.
 
@@
 
-As in \cite{Edwards2018}, the MCMC algorithm based on the BspDP prior  is run for 100,000 iterations, with a burn-in period of 50,000 and thinning factor of 10.
+As in \cite{Edwards2019}, the MCMC algorithm based on the BspDP prior  is run for 100,000 iterations, with a burn-in period of 50,000 and thinning factor of 10.
 Thus, 5,000 samples are used for the estimation.
@@
 Then the algorithm is run for 75,000 iterations with a burn-in period of 25,000 and thinning factor of 10, resulting in 5,000 samples.
-The analysis is performed using the first order penalty for equidistant and P-spaced knots.
+The analysis is performed using the first order penalty for equidistant and \textcolor{red}{adaptive} knots.
 All other specifications are the same as those used in the simulation study.
@@
 The continuous and dashed lines stand for the posterior median log psd obtained using the B-spline and P-spline priors, respectively.
-The dot lines represents the 90\% pointwise credible bands for the latter prior, which consider a first order penalty and P-spaced knots.}
+The dot lines represents the 90\% pointwise credible bands for the latter prior, which consider a first order penalty and \textcolor{red}{adaptive} knots.}
 	\label{fig:sunspot}
@@
 
-The results for the BspDP and P-spline prior  with first order penalty and P-spaced knots are displayed in Figure~\ref{fig:sunspot}.
+The results for the BspDP and P-spline prior  with first order penalty and \textcolor{red}{adaptive}  knots are displayed in Figure~\ref{fig:sunspot}.
 Both methods reveal a large peak at about 0.09.
@@
 In addition, the data has been squared root transformed and mean centred.
+{\pmr The first 150 observations are displayed in Figure~\ref{fig:carinae_data}, where the missing values are replaced by the mean.}
+
+\begin{figure}[]
+	\centering
+	\includegraphics[scale=0.4,clip=true,angle=0]{carinae_data.pdf}
+	\caption{{\pmr First 150 observations from the \textit{S. Carinae} data.}}
+	\label{fig:carinae_data}
+\end{figure}
 
@@
 We use 40 B-spline densities,  based on the rule-of-thumb criterion discussed above, and the first order penalty.
-The whole process takes less than 2 minutes to run and is replicated for equidistant and P-spaced knots.
+The whole process takes less than 2 minutes to run and is replicated for equidistant and \textcolor{red}{adaptive} knots.
 The results are displayed in Figure~\ref{fig:carinae}.
@@
 	\caption{Log-spectral density estimate for the variable star {\it S.\ Carinae} data based on the P-spline prior.
-The continuous grey line represents the periodogram whereas the continuous and dashed black lines stand for the posterior median log psd obtained using P-spaced and equidistant knots, respectively.}
+The continuous grey line represents the periodogram whereas the continuous and dashed black lines stand for the posterior median log psd obtained using \textcolor{red}{adaptive} and equidistant knots, respectively.}
 	\label{fig:carinae}
@@
 The periodogram shows several peaks with the main one located at around 0.42.  When the knots are equally spaced, the algorithm is unable to capture the sharpness of this peak,
-in contrast to the P-spaced knots.
+in contrast to the \textcolor{red}{adaptive} knots.
 This is because the periodogram-based knot allocation puts more knots in those regions which contain significant peaks as indicated by the periodogram.
-The rest of peaks, which are smaller in comparison to the main one, are slightly better captured by the equidistant knots, since a large proportion of the P-spaced knots have been allocated around the main peak.
+The rest of peaks, which are smaller in comparison to the main one, are slightly better captured by the equidistant knots, since a large proportion of the \textcolor{red}{adaptive} knots have been allocated around the main peak.
 However, this can easily  be corrected by increasing the number of knots.
-The psd estimate based on P-spaced knots is comparable to the one obtained via the NPC approach \citep{Kirch:2018}.
+The psd estimate based on \textcolor{red}{adaptive} knots is comparable to the one obtained via the NPC approach \citep{Kirch:2018}.
 
@@
 
-As shown by \cite{Edwards2018}, the BspDP prior outperforms the BDP prior in terms of IAE and uniform coverage probabilities for spiky psds.
-This is explained by the local support of the B-splines and its better approximation properties \citep{Edwards2018}.
+As shown by \cite{Edwards2019}, the BspDP prior outperforms the BDP prior in terms of IAE and uniform coverage probabilities for spiky psds.
+This is explained by the local support of the B-splines and its better approximation properties \citep{Edwards2019}.
 However, its performance is only achieved at a much higher computational cost.
@@
 This approach does not affect the computational time and improves the estimates dramatically.
-%This also indicates that the BspDP computation time could be reduced by the choice of P-spaced knots as initial values.
+%This also indicates that the BspDP computation time could be reduced by the choice of \textcolor{red}{adaptive} knots as initial values.
 
 In our simulation study, we showed that the equidistant knot scheme for the P-spline method works quite well in the case of simple spectral structures (AR(1) model).
-However, its performance is surpassed by the P-spaced knot scheme for complex structures (AR(4) model).
-In this scenario particularly, we can highlight the good performance of the P-spaced knot scheme.
+However, its performance is surpassed by the \textcolor{red}{adaptive} knot scheme for complex structures (AR(4) model).
+In this scenario particularly, we can highlight the good performance of the \textcolor{red}{adaptive} knot scheme.
 This method outperforms the BspDP prior in terms of IAE and run-times, particularly for the complex psds.
 Regarding uniform coverage,  a {\em first order} penalty P-spline coupled with
-P-spaced knots works best. For this combination, the uniform coverage almost reaches the nominal level.
+\textcolor{red}{adaptive} knots works best. For this combination, the uniform coverage almost reaches the nominal level.
 
-%In our simulation study, we showed that, in general, the P-spline prior with P-spaced knots outperforms the BspDP prior in terms of IAE and run-times, particularly for the complex psd.
+%In our simulation study, we showed that, in general, the P-spline prior with \textcolor{red}{adaptive} knots outperforms the BspDP prior in terms of IAE and run-times, particularly for the complex psd.
 However, the results were not conclusive for coverage probabilities.
@@
 
-We also assessed our proposal in the classic sunspot dataset, in which the B-spline prior algorithm is able to estimate correctly the solar cycle that occurs every 11.07 years.
-The P-spline prior yields almost identical results, but in  significantly  less computational time.
+We also assessed our proposal in the classic sunspot dataset, in which the \textcolor{red}{posterior distribution based on the BspDP prior }%B-spline prior algorithm
+ is able to estimate correctly the solar cycle that occurs every 11.07 years.
+The \textcolor{red}{posterior distribution based on the} P-spline prior yields almost identical results, but in  significantly  less computational time.
 
-Finally, we assess the P-spline prior algorithm under two different knot location schemes: equally spaced and periodogram-spaced knots.
+Finally, we assess the \textcolor{red}{ fit of the posterior distribution based on the} P-spline prior a under two different knot location schemes: equally spaced and periodogram-spaced knots.
 For this, we estimated the psd for the {\it S.\ Carinae} time series which has a spectral density with several sharp peaks.
-Equidistant knots failed to capture the peaks of this spectral density whereas  P-spaced knots significantly improved  the results without affecting the computational time.
+Equidistant knots failed to capture the peaks of this spectral density whereas  \textcolor{red}{adpative} knots significantly improved  the results without affecting the computational time.
 
@@
 
-In conclusion, the P-spline prior with the P-spaced knot scheme offers a more computationally viable alternative to the BspDP prior for spectral density estimation.
+\textcolor{red}{Another interesting question is whether the P-spline prior can be extended to enable a Bayesian analysis of multivariate time series. In this case, the spectral density is a 
+Hermitian positive definite  matrix-valued function. A novel Bayesain nonparametric approach to multivariate spectral density estimation was proposed by
+\cite{Meier2020} based on a generalization of the Bernstein-Dirchlet process prior where Bernstein polynomials are used to smooth a matrix-Gamma process.
+An avenue for future research will be to explore whether a computational speed-up might be possible by making use of P-splines in smoothing the elements of the
+Cholesky decomposition of the spectral density matrix.}
+
+In conclusion, the P-spline prior with the \textcolor{red}{adaptive} knot scheme offers a more computationally viable alternative to the BspDP prior for spectral density estimation.
 As shown in this work, it can handle spectral densities with several peaks in a reasonable run-time.

### 2020-03-25 response.tex
changes to pSplines.tex, response.tex

\documentclass{scrartcl}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{color}
\usepackage[english]{babel}
%\usepackage[fixlanguage]{babelbib}
\usepackage{natbib}
\usepackage[pdftex]{graphicx}
\usepackage{dsfont}
\usepackage{color}
%\usepackage{a4wide}
\usepackage[a4paper, left=3cm, right=3cm, top=2cm]{geometry}
%\usepackage{texititel}
\title{Spectral density estimation using P-spline priors}
\subtitle{Response letter to Statistical Papers}
\author{Patricio Maturana-Russel, Renate Meyer}

\newcommand{\refereeQuote}{\textit }
\newcommand{\responseDraft}{\textcolor{red}}
\newcommand{\response}{}

\parindent 0pt
\parskip 4pt

\setlength{\oddsidemargin}{0cm}
\setlength{\evensidemargin}{0cm}
\setlength{\topmargin}{-2cm}

\pagestyle{empty}
\begin{document}
\maketitle\thispagestyle{empty}

% -----------------------------------------------------------------------------
% -----------------------------------------------------------------------------
% -----------------------------------------------------------------------------

\subsection*{Referee \#1}

\response{We would like to thank the referee for a thorough review of our paper. Below, you find a point-by-point response to your points raised. We have now carefully revised the paper taking all of your -- and the other referees' -- comments into account and hope that you will now find the paper acceptable for publication. For your convenience, all changes made are marked in red in the manuscript.}

\begin{enumerate}
\item
\refereeQuote{The authors' terminology is surprising. They compute a P-spline fit to a power spectrum, but they call it a P-spline prior. The P-spline coefficients are the parameters of the model and they have a prior.}\smallskip

\response{We have edited the manuscript and use {\em P-spline prior} only when referring to the prior but use {\em posterior distribution based on the P-spline prior} when referring to the posterior distribution or Bayesian estimate of the spectral density. However, the referee raises a subtle point here in what frequentist and Bayesian regard as unknown parameters. In our view, the unknown parameter in this model is the spectral density function. As a prior on this spectral density function, we use a mixture of P-spline distributions. The mixture weights are hyperparameters with their own hyperprior distribution.}\bigskip

\item
\refereeQuote{To get a better fit to sharp peaks in the spectrum, the authors propose an adaptive scheme for knot placement. Apparently, they forgot that the differences in the penalty must change too, to correct for the variations in the distance between knots. One solution could be to use divided differences, another to use the algorithm of Wood for a derivative-based penalty.}\smallskip

\response{Thank you very much for pointing this out. You are absolutely correct, we did use the same penalty for the unequally spaced knots as for the equidistant knots.
We have rectified this and use derivative-based penalties for the unequally spaced knots. We describe the prior specification in paragraph 4  of the section  "adaptive knots" in Chapter 3 . As third order derivative-based penalties do not make sense for cubic splines and we want to compare the results for the equally-spaced knots and adaptive knots in the simulation study, we only compare results based on first and second order difference and derivative-based penalties, respectively. Therefore we deleted the definition of the third order difference matrix in Chapter 3. }\bigskip

\item
\refereeQuote{A more principle approach is to model the logarithm of the spectrum. That can be combined with adaptive smoothing, as described by Wood and Fasiolo, or Rodrigues-Alvarez et al. R packages are available.}\smallskip

\response{It is debatable whether putting a prior on the log spectral density is more {\em principled} than putting a prior on the spectral density itself as both terms occur in the log-likelihood function (1) and thus the prior can be put on either scale. Our P-spline prior is based on the idea of approximating  a {\em density} function as a mixture of basis distributions on the unit interval.
Thus, the original scale and not the log scale is appropriate for this particular approach even though for many applications, the log-scale might be the more natural one and we cited many papers that put a prior on the log spectral density, see our paragraph 5 of the introduction. This could be an interesting avenue to pursue in future research and see whether putting a P-spline based prior on the log spectral density has advantages over putting a prior on the spectral density. We added a paragraph in the Discussion. We have now also added the references to the alternative frequentist approaches that you suggested. Thank you for pointing these out.
}\bigskip

\item
\refereeQuote{The plots of spectra and fit with logarithmic are popular in spectrum analysis, but they are a bit misleading as the fitting is done on the linear scale.}\smallskip

\response{Please see our response to the previous point. Even though we fit a mixture of B-spline densities to the normalized spectral density, it is still possible to show the fit
on the log scale. Large peaks are better visible using the log scale. We also want to have a direct comparison of our model fits in Figure 1 and figure 2 to the corresponding ones shown  in Edwards et al. (2019) so it makes sense to use the same scale.}\bigskip

\item
\refereeQuote{I took a short look at the repository. It could be made more valuable by adding the data and some example scripts.}\smallskip

\response{Data and example scripts are included in the R package {\tt psplinePsd}.}\bigskip

\item
\refereeQuote{Sunspots counts are relatively familiar, but very few people will have seen a time the light curve of S. Carinae. I propose to add a plot of (a part of) the time series.}\smallskip

\response{We have added a time series plot of the first 150 observations of the \textit{S.\ Carinae} variable star light intensities data in Figure 4}.\bigskip

\item
\refereeQuote{
The term 'P-spaced knots' is confusing in the context of P-splines, P is already used for penalties. An adaptive scheme is proposed that is not exclusive to periodograms. Simply calling them adaptive knot looks adequate to me.}\smallskip

\response{ This could indeed be confusing. We have changed 'P-spaced' to 'adaptive' knots.}\bigskip

\item
\refereeQuote{That there is no consensus on the optimal number of knots for P-splines (page 10) has a simple explanation: it is irrelevant. Ruppert's recommendations have little value.}\smallskip

\response{We see this as a rough guideline only.}\bigskip

\item
\refereeQuote{I noticed some language problems}\smallskip

\response{Thank you for pointing these out. These have been corrected.}

\end{enumerate}

\newpage
% -----------------------------------------------------------------------------
% -----------------------------------------------------------------------------
% -----------------------------------------------------------------------------

\subsection*{Referee \#2}

\response{We would like to thank the referee for a thorough review of our paper. Below, you find a point-by-point response to your points raised. We have now carefully revised the paper taking all of your -- and the other referees' -- comments into account and hope that you will now find the paper acceptable for publication. For your convenience changes are marked in red in the manuscript.
} \smallskip

\begin{enumerate}
\item
\refereeQuote{The entire Section 2 is dedicated to the B-spline prior, which were already effectively introduced in Edwards et al. (2019). I would significantly reduce the length of this section.}\smallskip

\response{We have deleted the detailed description of the BspDP prior but kept the definition of the B-splines as these are required for the definition of the prior based on the  P-splines.}\bigskip

\item 
\refereeQuote{The proposed method deals with estimating one single spectral density. However, many existing paper already deal with this. Can the proposed methodology be extended to multiple or multivariate time series?}\smallskip

\response{
 This could be a
very interesting avenue for future research and we have added a paragraph in the Discussion suggesting a potential approach. However, a multivariate extension is outside the scope of this paper.}\bigskip

\item
\refereeQuote{A Metropolis step is used for v, but a data augmentation step based on Polya-Gamma auxiliary variables will allow for a full Gibbs sampler. See "Bayesian Inference for Logistic Models Using PolyaGamma Latent Variables" (2013) of Polson, Scott and Windle.}\smallskip

\response{We had already pointed this out in section "posterior computation" and had cited the paper by Polson et al. (2013).}

\end{enumerate}

\subsection*{Further Changes}

We have changed the title from {\em Spectral density estimation using P-spline priors} to {\em Bayesian spectral density estimation using P-spline priors}.\smallskip

We have added the keyword {\em Bernstein-Dirichlet process prior}.\smallskip

In the introduction, we have added references to relevant papers published in {\em Statistical Papers} that deal with the penalized spline approach for time series (these are Krivoboka et al (2006) and Wegener and Kauermann (2017)).

\end{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2020-03-25 pSplines.tex
changes to pSplines.tex, response.tex

@@
 in one way or another, where $-\pi < \lambda \leq \pi$ is the angular frequency and $\bm{Y}=(Y_1,\ldots,Y_n)$ is a stationary time series of length $n$.
-The periodogram has interesting features \textcolor{red}{(***)?why is that here?}.
-For instance, the periodogram evaluated at the Fourier frequencies $\lambda_i$  are asymptotically independent and follow an exponential distribution with mean $f(\lambda_i)$.
+The periodogram ordinates evaluated at the Fourier frequencies $\lambda_i$  are asymptotically independent and follow an exponential distribution with mean $f(\lambda_i)$.
 Thus, they are unbiased estimates of $f(\lambda_i)$  but not consistent \citep{Brockwell:1986}.
@@
 =======
-With the aim of reducing the computational cost of the BspDP prior approach,  we suggest to use the idea of P-splines \cite{Eilers:1996} that combine a large but fixed number of  B-splines with a simple difference penalty that controls the degree of  smoothness of the spectral density.
+With the aim of reducing the computational cost of the BspDP prior approach,  we suggest to use the idea of P-splines \cite{Eilers:1996} that combines a large but fixed number of  B-splines with a simple difference penalty that controls the degree of  smoothness of the spectral density.
 P-splines, i.e.\  equally-spaced B-splines,  have been successfully implemented for frequentist nonparametric regression but to the best of our knowledge not yet for Bayesian spectral density estimation.
@@
 where $\bm{v} = (v_1,\dots,v_{K-1})^\top$ is a $K-1$ dimensional parameter vector, $\phi$ is the smoothing or penalty parameter, $\textbf{P} = \textbf{D}^\top \textbf{D} +\epsilon \textbf{I}_{K-1}$ is the penalty matrix (which is a full matrix rank matrix for any small quantity $\epsilon$, for instance, $10^{-6}$) with $\textbf{D}$ the $d^{\text{th}}$ order difference matrix ($d=1,2,\ldots, K-2)$, $\alpha_{\phi}$ and $\alpha_{\delta}$ are shape parameters, $\delta \beta_{\phi}$ and $\beta_{\delta}$ are rate parameters, and $\text{Gamma}(a,b)$ denotes a gamma distribution with mean $a/b$ and variance $a/b^2$.
-We used first, second and third order difference penalties in the simulation study.
-The $1^{\text{st}}$ order difference matrix \textbf{D} is defined as
-\begin{align*}
+We used \textcolor{red}{first and second}  order difference penalties in the simulation study.
+The $1^{\text{st}}$ order difference matrix \textbf{D} $\in \mathbb{R}^{(K-2)\times (K-1)}$  is defined as
+\begin{equation}\label{eq:D1}
+%\begin{align*}
 \begin{bmatrix}
@@
 \end{bmatrix}
-\in \mathbb{R}^{(K-2)\times (K-1)},
-\end{align*}
-for the $2^{\text{nd}}$ order is
-\begin{align*}
+%\end{align*}
+\end{equation}
+\textcolor{red}{and} the $2^{\text{nd}}$ order difference matrix $\in \mathbb{R}^{(K-3)\times (K-1)} $ as
+\begin{equation}\label{eq:D2}
+%\begin{align*}
 \begin{bmatrix}
@@
 0 & &&&& \cdots & 0 & 1 & -2 & 1 
-\end{bmatrix}
-\in \mathbb{R}^{(K-3)\times (K-1)},
-\end{align*}
-and for the $3^{\text{rd}}$ order is
-\begin{align*}
-\setcounter{MaxMatrixCols}{12} % to include more than 10 columns in bmatrix enviroment
-\begin{bmatrix}
-1 & -3 &  3 & -1 &  0 & 0 & \cdots & & & & & 0 \\
-0 &  1 & -3 &  3 & -1 & 0 & \cdots & & & & & 0 \\
-\vdots &    & \ddots &  \ddots & \ddots & \ddots & &  & & & & \vdots \\
-0 &   &  &   &  &  & \cdots & 0 & 1 & -3 & 3 & -1  
-\end{bmatrix}
-\end{align*}
-$\in \mathbb{R}^{(K-4)\times (K-1)}$.
+\end{bmatrix}.
+%\end{align*}
+\end{equation}
+%and for the $3^{\text{rd}}$ order is
+%\begin{align*}
+%\setcounter{MaxMatrixCols}{12} % to include more than 10 columns in bmatrix enviroment
+%\begin{bmatrix}
+%1 & -3 &  3 & -1 &  0 & 0 & \cdots & & & & & 0 \\
+%0 &  1 & -3 &  3 & -1 & 0 & \cdots & & & & & 0 \\
+%\vdots &    & \ddots &  \ddots & \ddots & \ddots & &  & & & & \vdots \\
+%0 &   &  &   &  &  & \cdots & 0 & 1 & -3 & 3 & -1  
+%\end{bmatrix}
+%\end{align*}
+%$\in \mathbb{R}^{(K-4)\times (K-1)}$.  
+\textcolor{red}{The rationale for using discrete second order roughness penalties as an approximation to the usual continuous roughness measure based on the integral over the second derivatives is given
+in \cite{Eilers:1996}.}
 
@@
 
-Our\textcolor{red}{adaptive} knot location proposal works as follow.
+Our \textcolor{red}{adaptive} knot location proposal works as follow.
 We calculate the periodogram and apply a square root transformation in order to have more regular magnitudes and eliminate a potential trend.
@@
 
-The \textcolor{red}{adaptive}  knot allocation can be regarded as an empirical Bayes prior as the data are used to specify the prior distribution.
+\textcolor{red}{For non-equidistant knots, it does not make sense any longer to base the definition of the covariance matrix of the Gaussian prior  on the  difference matrices as defined in (\ref{eq:D1}) and (\ref{eq:D2}) because this is an adequate approximation of the derivative-based roughness measures -- defined as the integral of the first and second order derivatives of the P-splines --
+for equally spaced knots only.
+The definition of the difference matrices could be adjusted using divided differences, but here we use the derivative-based penalties as described by \cite{WoodSimon2017Pwdb}.
+We make use of the implementation of the derivative-based penalty matrix  by the function {\tt bsplinepen} in the R-package {\tt fda}.
+This penalty matrix contains the inner products of the respective first and second order derivatives of the basis functions.}
+
+The \textcolor{red}{adaptive}  knot allocation can be regarded as an empirical Bayes prior since the data are used to specify the prior distribution.
 However, only the locations of the B-splines are determined by the data, not the parameters of the prior distribution which in this case are the weights $w_k$ associated with each B-spline $b_{k,r}$.
@@
 In addition, the data has been squared root transformed and mean centred.
-{\pmr The first 150 observations are displayed in Figure~\ref{fig:carinae_data}, where the missing values are replaced by the mean.}
+\textcolor{red}{ The first 150 observations are displayed in Figure~\ref{fig:carinae_data}, where the missing values are replaced by the mean.}
 
@@
 	\includegraphics[scale=0.4,clip=true,angle=0]{carinae_data.pdf}
-	\caption{{\pmr First 150 observations from the \textit{S. Carinae} data.}}
+	\caption{ \textcolor{red}{First 150 observations of the \textit{S. Carinae} data.}}
 	\label{fig:carinae_data}
@@
 
+\textcolor{red}{As already reviewed in the introduction, many Bayesian nonparametric approaches to spectral density estimation propose to put a prior on the log-spectral density.
+This could be easily implemented by modifying (\ref{eq:ps_prior}) to
+ \[ \log f(\pi \omega)=\sum_{k=1}^K v_k b_{k,r}(\omega,{\bm \xi})\]
+and it could be interesting to examine for what class of spectral densities this prior specification would yield better posterior inference. }
+
 \textcolor{red}{Another interesting question is whether the P-spline prior can be extended to enable a Bayesian analysis of multivariate time series. In this case, the spectral density is a 

### 2020-03-27 pSplines.tex
updated results
Updated results for the simulation study

@@
 We make use of the implementation of the derivative-based penalty matrix  by the function {\tt bsplinepen} in the R-package {\tt fda}.
-This penalty matrix contains the inner products of the respective first and second order derivatives of the basis functions.}
+This penalty matrix contains the inner products of the respective first and second order derivatives of the basis functions.} {\pmr In addition, we normalise this matrix by dividing it by the maximum absolute column sum.}
 
@@
 The P-spline analysis was performed using the \texttt{gibbs\_pspline} function implemented in the \textsf{R} package \texttt{psplinePsd}.  A total of 100,000 MCMC samples were generated
-using  a pilot run of 25,000 iterations with a burn-in period of 5,000 to calibrate the proposals.
-The final sample consisted of 75,000 samples with a burn-in period of 5,000 and thinning factor of 10, resulting in 7,000 samples used for posterior inferences.
+using  a pilot run of 20,000 iterations with a burn-in period of 5,000 and thinning factor of 10 to calibrate the proposals.
+The final sample consisted of 80,000 samples with a burn-in period of 5,000 and thinning factor of 10, resulting in 7,500 samples used for posterior inferences.
 %Therefore, the number of iterations equate the B-spline analysis.  
-This procedure was replicated for both equidistant and \textcolor{red}{adaptive} knots.
+This procedure was replicated for both equidistant and \textcolor{red}{adaptive} knots{\pmr considering first and second order penalties.}
 
@@
 			\textit{Equidistant knots}& & &   \\
-			\hspace{0.5em}P-spline $d=1$ &  0.729 & 0.632 & 0.517  \\
-			\hspace{0.5em}P-spline $d=2$&  0.844 & 0.612 & 0.638  \\
-			\hspace{0.5em}P-spline $d=3$&  0.727 & 0.728 & 0.769  \\
+			\hspace{0.5em}P-spline $d=1$ &  0.772 & 0.644 & 0.527 \\
+			\hspace{0.5em}P-spline $d=2$&  0.708 & 0.604 & 0.629 \\
 			\textit{\textcolor{red}{Adaptive} knots}& & &   \\
-			\hspace{0.5em}P-spline $d=1$ & 0.910 & 0.769 & 0.595 \\
-			\hspace{0.5em}P-spline $d=2$ & 0.994 & 0.867 & 0.731 \\
-			\hspace{0.5em}P-spline $d=3$ & 1.059 & 0.829 & 0.610 \\ \hline
+			\hspace{0.5em}P-spline $d=1$ & 0.874 & 0.783 & 0.610 \\
+			\hspace{0.5em}P-spline $d=2$ & 0.874 & 0.815 & 0.640 \\ \hline
 			AR(4)     &  &  &  \\ 
@@
 			\textit{Equidistant knots}& & &   \\
-			\hspace{0.5em}P-spline $d=1$ & 3.014 & 2.462 & 2.335  \\
-			\hspace{0.5em}P-spline $d=2$& 3.247 & 2.619 & 2.399 \\
-			\hspace{0.5em}P-spline $d=3$& 3.438 & 2.786 & 2.510  \\	
+			\hspace{0.5em}P-spline $d=1$ &  2.914 & 2.389 & 2.307\\
+			\hspace{0.5em}P-spline $d=2$&  3.150 & 2.569 & 2.388\\
 			\textit{\textcolor{red}{Adaptive} knots}& & &   \\
-			\hspace{0.5em}P-spline $d=1$ & 2.659 & 1.884 & 1.465 \\
-			\hspace{0.5em}P-spline $d=2$ & 2.627 & 1.849 & 1.411 \\
-			\hspace{0.5em}P-spline $d=3$ & 2.625 & 1.835 & 1.410 \\
+			\hspace{0.5em}P-spline $d=1$ & 2.495 & 1.830 & 1.489\\
+			\hspace{0.5em}P-spline $d=2$ &  2.696 & 2.003 & 1.677\\
 			\bottomrule
@@
 
-For the AR(1) model, the BspDP and first order penalty P-spline 90\% credible bands cover the whole true spectral density function, i.e.\ for all frequencies $\lambda$, in all the 300 independent analyses.
+For the AR(1) model, the BspDP and first order penalty P-spline 90\% credible bands cover the whole true spectral density function, i.e.\ for all frequencies $\lambda$, in {\pmr almost} all the 300 independent analyses.
 In general, the latter method shows diminishing  performance as the penalty degree increases.
@@
 On the other hand, the \textcolor{red}{adaptive} knot scheme yields evidently much better results, which are comparable (in the case of the first order penalty) to those obtained via the BspDP algorithm.
-For $n=\{128,256\}$, the latter has higher coverage probabilities than our proposal, except for $n=512$.
+%For $n=\{128,256\}$, the latter has higher coverage probabilities than our proposal, except for $n=512$.
 For the P-spline priors, the coverage probabilities decrease as the penalty degree order increases, which can be explained by the forced smoothness imposed by the higher penalty order.
@@
 			\textit{Equidistant}& & &   \\
-			\hspace{0.5em}P-spline $d=1$& 1.000 & 1.000 & 1.000  \\
-			\hspace{0.5em}P-spline $d=2$& 0.987 & 0.993 & 1.000  \\
-			\hspace{0.5em}P-spline $d=3$& 0.943 & 0.877 & 0.870  \\
+			\hspace{0.5em}P-spline $d=1$&  1.000 & 1.000 & 0.997\\
+			\hspace{0.5em}P-spline $d=2$& 0.993 & 1.000 & 0.993  \\
 			\textit{\textcolor{red}{Adaptive} knots}& & &   \\
-			\hspace{0.5em}P-spline $d=1$ & 1.000 & 1.000 & 1.000 \\
-			\hspace{0.5em}P-spline $d=2$ & 0.993 & 0.980 & 0.990 \\
-			\hspace{0.5em}P-spline $d=3$ & 0.947 & 0.840 & 0.957 \\ \hline		       
+			\hspace{0.5em}P-spline $d=1$ & 0.997 & 1.000 & 0.997 \\
+			\hspace{0.5em}P-spline $d=2$ & 0.973 & 0.740 & 0.770 \\ \hline		       
 			AR(4)     &  &  &  \\ 
@@
 			\textit{Equidistant knots}& & &   \\
-			\hspace{0.5em}P-spline $d=1$ & 0.400 & 0.353 & 0.010 \\
-			\hspace{0.5em}P-spline $d=2$ & 0.200 & 0.167 & 0.003 \\
-			\hspace{0.5em}P-spline $d=3$ & 0.053 & 0.023 & 0.000 \\
+			\hspace{0.5em}P-spline $d=1$ &  0.527 & 0.353 & 0.007 \\
+			\hspace{0.5em}P-spline $d=2$ & 0.323 & 0.167 & 0.007\\
 			\textit{\textcolor{red}{Adaptive} knots}& & &   \\	
-			\hspace{0.5em}P-spline $d=1$ & 0.820 & 0.887 & 0.920 \\
-			\hspace{0.5em}P-spline $d=2$ & 0.637 & 0.700 & 0.773 \\
-			\hspace{0.5em}P-spline $d=3$ & 0.470 & 0.600 & 0.583 \\	
+			\hspace{0.5em}P-spline $d=1$ & 0.790 & 0.847 & 0.813 \\
+			\hspace{0.5em}P-spline $d=2$ & 0.337 & 0.387 & 0.270 \\
 			\bottomrule
@@
 			\textit{\textcolor{red}{Adaptive} knots}& & &   \\
-			\hspace{0.5em}P-spline $d=1,2,3$ & 1.000 & 1.000 & 1.000  \\ \hline	 		      
+			\hspace{0.5em}P-spline $d=1,2$ & 1.000 & 1.000 & 1.000  \\ \hline	 		      
 			AR(4)     &  &  &  \\ 
@@
 			\textit{Equidistant knots}& & &   \\
-			\hspace{0.5em}P-spline $d=1$ & 0.984 & 0.992 & 0.984 \\
+			\hspace{0.5em}P-spline $d=1$ &  1.000 & 0.992 & 0.984 \\
 			\hspace{0.5em}P-spline $d=2$&  0.984 & 0.984 & 0.980 \\
-			\hspace{0.5em}P-spline $d=3$& 0.968 & 0.976 & 0.969  \\	
 			\textit{\textcolor{red}{Adaptive} knots}& & &   \\	
 			\hspace{0.5em}P-spline $d=1$ & 1.000 & 1.000 & 1.000 \\
-			\hspace{0.5em}P-spline $d=2$ & 1.000 &  1.000 &  1.000 \\
-			\hspace{0.5em}P-spline $d=3$ & 0.984 &  1.000 &  1.000 \\		
+			\hspace{0.5em}P-spline $d=2$ & 0.984 & 0.992 & 0.992\\	
 			\bottomrule
@@
 Note that the P-spline method based on equidistant knots has similar run-times to the \textcolor{red}{adapative} knots.
-The P-spline method is approximately 3-6 times faster than BspDP approach in these examples.
+The P-spline method is approximately {\pmr 4-7} times faster than BspDP approach in these examples.
 This is due to a reduced number of calculations  per iteration because the B-spline densities are calculated only  once at the start of the algorithm and there are no Dirichlet process approximations via the stick-breaking method.
@@
 			\textit{\textcolor{red}{Adaptive} knots}& & &   \\
-			\hspace{0.5em}P-spline $d=1$& 5.629 & 8.291 & 10.507 \\
-			\hspace{0.5em}P-spline $d=2$& 5.650 & 8.374 & 10.512 \\
-			\hspace{0.5em}P-spline $d=3$& 5.587 & 8.301 & 10.427 \\ 
-			B-spline/P-spline &  5.247 & 3.896 & 3.615 \\ \hline
+			\hspace{0.5em}P-spline $d=1$&  4.859 & 7.450 & 9.606\\
+			\hspace{0.5em}P-spline $d=2$&  4.538 & 6.863 & 8.783\\
+			B-spline/P-spline &  6.278 & 4.531 & 4.122 \\ \hline
 			AR(4)     &  &  & \\ 
@@
 			\textit{\textcolor{red}{Adaptive} knots}& & &   \\
-			\hspace{0.5em}P-spline $d=1$ & 5.621 & 8.240 & 10.462 \\
-			\hspace{0.5em}P-spline $d=2$& 5.641 & 8.372 & 10.549 \\
-			\hspace{0.5em}P-spline $d=3$& 5.611 & 8.327 & 10.686 \\
-			B-spline/P-spline & 5.354 & 4.105 & 3.869 \\	
+			\hspace{0.5em}P-spline $d=1$ & 4.883 & 7.453 & 9.665 \\
+			\hspace{0.5em}P-spline $d=2$& 4.543 & 6.887 & 8.827 \\
+			B-spline/P-spline & 6.389 & 4.760 & 4.421 \\	
 			\bottomrule
@@
 The  rule-of-thumb criterion,  $K=\min\left\{n/4, 40\right\}$, for the number of
-\textcolor{red}{P-splines} %B-spline densities in the P-spline prior algorithm
- worked perfectly well in this example.
+%\textcolor{red}{P-splines} B-spline densities in the P-spline prior algorithm
+% worked perfectly well in this example.
 

### 2020-03-28 response.tex
final version
updated manuscript and references,
updated response letter

@@
 \response{We have edited the manuscript and use {\em P-spline prior} only when referring to the prior but use {\em posterior distribution based on the P-spline prior} when referring to the posterior distribution or Bayesian estimate of the spectral density.
+We use {\em P-spline prior} in analogy to {\em B-spline Dirichlet process prior} and {\em Bernstein-Dirichlet process prior}.
 However, the referee raises a subtle point here in what frequentist and Bayesian regard as unknown parameters.
@@
 \response{Thank you very much for pointing this out. You are absolutely correct, we did use the same penalty for the unequally spaced knots as for the equidistant knots.
-We have rectified this and use derivative-based penalties for the unequally spaced knots.
-We describe the prior specification in paragraph 4  of the section  "adaptive knots" in Chapter 3 .
-As third order derivative-based penalties do not make sense for cubic splines and we want to compare the results for the equally-spaced knots and adaptive knots in the simulation study, we only compare results based on first and second order difference and derivative-based penalties, respectively.
+We have rectified this and now use derivative-based penalties for the unequally spaced knots.
+We describe the prior specification in paragraph 4  of the section  "adaptive knots" in Chapter 3.
+As third order derivative-based penalties do not make sense for cubic splines and also because we want to compare the results for the equally-spaced knots and adaptive knots in the simulation study, we only compare results based on first and second order difference and derivative-based penalties, respectively.
 Therefore we deleted the definition of the third order difference matrix in Chapter 3.
@@
 
-\response{It is debatable whether putting a prior on the log spectral density is more {\em principled} than putting a prior on the spectral density itself as both terms occur in the log-likelihood function (1) and thus the prior can be put on either scale.
+\response{It is debatable whether putting a prior on the log spectral density is more {\em principled} than putting a prior on the spectral density itself as both terms enter the Whittle likelihood function (1) and thus the prior can be put on either scale.
 Our P-spline prior is based on the idea of approximating  a {\em density} function as a mixture of basis distributions on the unit interval.
@@
 Thank you for pointing these out.
+
+We were wondering whether the referee meant "model the logarithm of the periodogram".
+One might consider to model the log-periodogram instead of the periodogram itself which would require using the logarithm of an exponential distribution or approximating
+this via a mixture of Gaussian distributions as in Carter and Kohn (1997).
+However, here we stick to the traditional Whittle likelihood which is contiguous to true likelihood for Gaussian stationary time series (Kirch et al 2019).
 }\bigskip
@@
 Large peaks are better visible using the log scale.
-We also want to have a direct comparison of our model fits in Figure 1 and figure 2 to the corresponding ones shown  in Edwards et al.
+We also want to have a direct comparison of our model fits in Figure 1 and Figure 2 to the corresponding ones shown  in Edwards et al.
 (2019) so it makes sense to use the same scale.}\bigskip
@@
 
-We have changed the title from {\em Spectral density estimation using P-spline priors} to {\em Bayesian spectral density estimation using P-spline priors}.\smallskip
+We have changed the title from {\em Spectral density estimation using P-spline priors} to {\em Bayesian spectral density estimation using P-splines}.\smallskip
 

### 2020-03-28 pSplines.tex
final version
updated manuscript and references,
updated response letter

@@
 
-\title{Bayesian spectral density estimation using P-spline priors}
+\title{\textcolor{red}{Bayesian spectral density estimation using P-splines}}
 \author{Patricio Maturana-Russel \and Renate Meyer}
@@
 We make use of the implementation of the derivative-based penalty matrix  by the function {\tt bsplinepen} in the R-package {\tt fda}.
-This penalty matrix contains the inner products of the respective first and second order derivatives of the basis functions.} {\pmr In addition, we normalise this matrix by dividing it by the maximum absolute column sum.}
+This penalty matrix contains the inner products of the respective first and second order derivatives of the basis functions.
+In addition, we normalise this matrix by dividing it by the maximum absolute column sum.}
 
@@
 
-The P-spline analysis was performed using the \texttt{gibbs\_pspline} function implemented in the \textsf{R} package \texttt{psplinePsd}.  A total of 100,000 MCMC samples were generated
+The P-spline analysis was performed using the \texttt{gibbs\_pspline} function implemented in the \textsf{R} package \texttt{psplinePsd} \cite{psplinepackage}.
+A total of 100,000 MCMC samples were generated
 using  a pilot run of 20,000 iterations with a burn-in period of 5,000 and thinning factor of 10 to calibrate the proposals.
@@
 %Therefore, the number of iterations equate the B-spline analysis.  
-This procedure was replicated for both equidistant and \textcolor{red}{adaptive} knots{\pmr considering first and second order penalties.}
+This procedure was replicated for both equidistant and \textcolor{red}{adaptive} knots \textcolor{red}{ considering first and second order penalties.}
 
@@
 
-For the AR(1) model, the BspDP and first order penalty P-spline 90\% credible bands cover the whole true spectral density function, i.e.\ for all frequencies $\lambda$, in {\pmr almost} all the 300 independent analyses.
+For the AR(1) model, the BspDP and first order penalty P-spline 90\% credible bands cover the whole true spectral density function, i.e.\ for all frequencies $\lambda$, in 
+\textcolor{red}{almost} all the 300 independent analyses.
 In general, the latter method shows diminishing  performance as the penalty degree increases.
@@
 Note that the P-spline method based on equidistant knots has similar run-times to the \textcolor{red}{adapative} knots.
-The P-spline method is approximately {\pmr 4-7} times faster than BspDP approach in these examples.
+The P-spline method is approximately \textcolor{red}{4-7} times faster than BspDP approach in these examples.
 This is due to a reduced number of calculations  per iteration because the B-spline densities are calculated only  once at the start of the algorithm and there are no Dirichlet process approximations via the stick-breaking method.
@@
 
-The \textsf{R} package \texttt{psplinePsd} will be available on CRAN in due course.  In the meanwhile, it can be downloaded from \href{https://github.com/pmat747/psplinePsd}{GitHub}.
+\textcolor{red}{The \textsf{R} package \texttt{psplinePsd} can be downloaded from \href{https://github.com/pmat747/psplinePsd}{GitHub}.}
+%The \textsf{R} package \texttt{psplinePsd} will be available on CRAN in due course.  In the meanwhile, it can be downloaded from \href{https://github.com/pmat747/psplinePsd}{GitHub}.
 

### 2020-06-29 pSplines.tex
update

@@
 \documentclass[twocolumn,final]{svjour3}
+
+% Statistical Papers format
+%\RequirePackage{fix-cm}
+%\documentclass[smallextended]{svjour3} % Statistical Journal
+%\smartqed  % flush right qed marks, e.g. at end of proof
+%
 
@@
 
-\institute{Patricio Maturana-Russel \at Department of Statistics, University of Auckland, Auckland, New Zealand \\ \email{p.russel@auckland.ac.nz} 
+\institute{Patricio Maturana-Russel \at Department of Mathematical Sciences, Auckland
+	University of Technology, Auckland, New Zealand \\ \email{p.maturana.russel@aut.ac.nz} 
 \and Renate Meyer  \at Department of Statistics, University of Auckland, Auckland, New Zealand 
@@
 
-This article  a Bayesian approach to estimating the spectral density of a stationary time series using a prior based on a mixture of P-spline distributions.
+This article  proposes a Bayesian approach to estimating the spectral density of a stationary time series using a prior based on a mixture of P-spline distributions.
 Our proposal is motivated by the B-spline Dirichlet process prior of \cite{Edwards2019} in combination  with Whittle's likelihood and aims at  reducing the high computational complexity of its posterior computations.
@@
 Here, we suggest to use P-splines of \cite{Eilers:1996} that combine a B-spline basis with a discrete penalty on the basis coefficients.
-We demonstrate in a simulation study and two real  case studies that this approach retains the flexibility of the B-splines, achieves the same ability to accurately estimate peaks due to a new data-driven knot allocation scheme but significantly reduces the computational costs.
+We demonstrate in a simulation study and two real  case studies that this approach retains the flexibility of the B-splines, achieves \textcolor{red}{similar} ability to accurately estimate peaks due to a new data-driven knot allocation scheme but significantly reduces the computational costs.
 
@@
 
-Most Bayesian approaches to psd estimation use the so-called Whittle likelihood, introduced by \citep{Whittle:1957} as an approximation to the likelihood of a Gaussian stationary time series.
+Most Bayesian approaches to psd estimation use the so-called Whittle likelihood, introduced by \cite{Whittle:1957} as an approximation to the likelihood of a Gaussian stationary time series.
 Based on the  limiting independent exponential distribution of the periodogram \textcolor{red}{values}, the Whittle likelihood for a mean centred weakly stationary time series $\bm{Y}$ of length $n$ is defined as
@@
 who used a smoothing prior on the log psd.
-\cite{Rosen:2012} decompose the log psd into a linear and nonlinear component and  put a Gaussian process prior on the nonlinear part, \cite{Pensky:2007} propose Bayesian wavelet-smoothing of the log psd and recently, and \cite{Cadonna2017} modelled the log psd  as a mixture of Gaussian distributions with frequency-dependent weights and mean functions.
+\cite{Rosen:2012} decompose the log psd into a linear and nonlinear component and  put a Gaussian process prior on the nonlinear part, \cite{Pensky:2007} propose Bayesian wavelet-smoothing of the log psd and recently, \cite{Cadonna2017} modelled the log psd  as a mixture of Gaussian distributions with frequency-dependent weights and mean functions.
 %
@@
 
-%<<<<<<< HEAD
 With the aim of reducing the computational cost of the BspDP prior approach,  we suggest to use the idea of P-splines of \cite{Eilers:1996} that \textcolor{red}{combines} a large but fixed number of  B-splines with a simple difference penalty that controls the degree of  smoothness of the spectral density.
@@
 We show that with this knot allocation scheme, the estimates of spectral densities with spikes and abrupt changes improves significantly without an increase in computing time.
-=======
-With the aim of reducing the computational cost of the BspDP prior approach,  we suggest to use the idea of P-splines \cite{Eilers:1996} that combines a large but fixed number of  B-splines with a simple difference penalty that controls the degree of  smoothness of the spectral density.
-P-splines, i.e.\  equally-spaced B-splines,  have been successfully implemented for frequentist nonparametric regression but to the best of our knowledge not yet for Bayesian spectral density estimation.
-We suggest
-two novel Bayesian  approaches in this paper, both based on the Whittle likelihood and on a mixture of B-spline densities. Both approaches fix the  number of B-spline densities. 
-The first approach uses equally spaced knots with a P-spline prior which we define in Section III.
-We demonstrate that this approach yields significant savings in computational time without sacrificing accuracy over the BspDP prior approach for simple spectral structures.
-The second approach has a particular  focus on spectral densities with sharp peaks.
-Instead of an equidistant spacing, the knots are spread based on the distribution of peaks in the periodogram.
-Thus the knot allocation is also data-driven as with the BspDP prior but the knots  remain fixed once
-allocated.
-This novel knot allocation scheme avoids the computational complexity of the BspDP prior.
-We show that with this knot allocation scheme, the estimates of spectral densities with spikes and abrupt changes improves significantly without an increase in computing time.
-%>>>>>>> 19c7083b8cf09264b1cfb0f5ba98e32bf3de8990
 
@@
 In Section IV, we test our P-spline-based approaches to spectral density estimation and compare results to those based on the BspDP prior in a simulation study and the application to two real time series.
-The paper concludes with a discussion of the relative merits of the new approaches.
+The paper concludes with a discussion of the relative merits of the new approaches \textcolor{red}{and avenues for future research.}
 
@@
 \widehat{g}(\omega) &= \sum_{k=1}^{K} G \left( \dfrac{k-1}{K} , \dfrac{k}{K} \right] \beta(\omega; k, K-k+1)\\
-&= \sum_{k=1}^{K} w_k \beta(\omega; k, K-k+1)
+&= \sum_{k=1}^{K} w_k \beta(\omega; k, K-k+1) \nonumber
 \end{align}	  
@@
 The definition of the difference matrices could be adjusted using divided differences, but here we use the derivative-based penalties as described by \cite{WoodSimon2017Pwdb}.
-We make use of the implementation of the derivative-based penalty matrix  by the function {\tt bsplinepen} in the R-package {\tt fda}.
+We make use of the implementation of the derivative-based penalty matrix  by the function {\tt bsplinepen} in the R-package {\tt fda} \citep{fda}.
 This penalty matrix contains the inner products of the respective first and second order derivatives of the basis functions.
@@
 
-The P-spline analysis was performed using the \texttt{gibbs\_pspline} function implemented in the \textsf{R} package \texttt{psplinePsd} \cite{psplinepackage}.
+The P-spline analysis was performed using the \texttt{gibbs\_pspline} function implemented in the \textsf{R} package \texttt{psplinePsd} \citep{psplinepackage}.
 A total of 100,000 MCMC samples were generated
-using  a pilot run of 20,000 iterations with a burn-in period of 5,000 and thinning factor of 10 to calibrate the proposals.
-The final sample consisted of 80,000 samples with a burn-in period of 5,000 and thinning factor of 10, resulting in 7,500 samples used for posterior inferences.
+using  a pilot run of \textcolor{red}{20,000} iterations with a burn-in period of 5,000 and thinning factor of 10 to calibrate the proposals.
+The final sample consisted of \textcolor{red}{80,000} samples with a burn-in period of 5,000 and thinning factor of 10, resulting in \textcolor{red}{7,500} samples used for posterior inferences.
 %Therefore, the number of iterations equate the B-spline analysis.  
@@
 
-In general terms, the P-spline algorithm with our novel knot allocation scheme, the periodogram-spaced knots, offers the best trade-off between accuracy and computation cost.
+In general terms, the P-spline algorithm with our novel knot allocation scheme, the \textcolor{red}{adaptive} knots, offers the best trade-off between accuracy and computation cost.
 The  rule-of-thumb criterion,  $K=\min\left\{n/4, 40\right\}$, for the number of B-spline densities in the P-spline prior algorithm worked perfectly well in this example.
@@
 Thus, 5,000 samples are used for the estimation.
-This took about 32 minutes.
+This took about \textcolor{red}{22} minutes.
 The other specifications are the same as those used in the simulation study.
@@
 All other specifications are the same as those used in the simulation study.
-A single analysis takes about 6 minutes.
+A single analysis takes about \textcolor{red}{4.5} minutes.
 
@@
 
-The results for the BspDP and P-spline prior  with first order penalty and \textcolor{red}{adaptive}  knots are displayed in Figure~\ref{fig:sunspot}.
+The results for the BspDP and P-spline prior  with first order penalty and \textcolor{red}{adaptive}  knots are displayed in Figure~\ref{fig:sunspot} \textcolor{red}{(p-spline with equidistant knot approach results are not shown due to their similarity to the adaptive knot estimates)}.
 Both methods reveal a large peak at about 0.09.
@@
 These results are consistent with those obtained via the BDP  prior~\citep{Choudhuri:2004}.
-However, the BspDP estimation takes approximately 32 minutes compared to only 6 minutes for the the P-spline estimation. Thus the P-spline estimation is approximately
+However, the BspDP estimation takes approximately \textcolor{red}{22} minutes compared to only \textcolor{red}{4.5} minutes for the the P-spline estimation.
+Thus the P-spline estimation is approximately
 five times more efficient.
@@
 We use 40 B-spline densities,  based on the rule-of-thumb criterion discussed above, and the first order penalty.
-The whole process takes less than 2 minutes to run and is replicated for equidistant and \textcolor{red}{adaptive} knots.
+The whole process takes less than \textcolor{red}{1.5} minutes to run and is replicated for equidistant and \textcolor{red}{adaptive} knots.
 The results are displayed in Figure~\ref{fig:carinae}.
@@
 The P-spline prior assumes a fixed number of knots  with fixed locations, thus almost looks like a special case.
-However, the multivariate prior based on the difference penalty of the basis coefficients controls the smoothness of the spectral density, a feature that can't be achieved with a univariate Dirichlet process prior.
+However, the multivariate prior based on the difference penalty of the basis coefficients controls the smoothness of the spectral density, a feature that cannot be achieved with a univariate Dirichlet process prior.
 Most importantly, fixing the number and location of knots reduces the computational effort significantly, without sacrificing accuracy as shown in the application section.
@@
 
-Finally, we assess the \textcolor{red}{ fit of the posterior distribution based on the} P-spline prior a under two different knot location schemes: equally spaced and periodogram-spaced knots.
+Finally, we assess the \textcolor{red}{ fit of the posterior distribution based on the} P-spline prior a under two different knot location schemes: equally spaced and \textcolor{red}{adaptive} knots.
 For this, we estimated the psd for the {\it S.\ Carinae} time series which has a spectral density with several sharp peaks.
-Equidistant knots failed to capture the peaks of this spectral density whereas  \textcolor{red}{adpative} knots significantly improved  the results without affecting the computational time.
+Equidistant knots failed to capture the peaks of this spectral density whereas  \textcolor{red}{adaptive} knots significantly improved  the results without affecting the computational time.
 

### 2020-06-30 pSplines.tex
update

@@
 
-\title{\textcolor{red}{Bayesian spectral density estimation using P-splines}}
+\title{Bayesian spectral density estimation using P-splines}
 \author{Patricio Maturana-Russel \and Renate Meyer}
@@
 Here, we suggest to use P-splines of \cite{Eilers:1996} that combine a B-spline basis with a discrete penalty on the basis coefficients.
-We demonstrate in a simulation study and two real  case studies that this approach retains the flexibility of the B-splines, achieves \textcolor{red}{similar} ability to accurately estimate peaks due to a new data-driven knot allocation scheme but significantly reduces the computational costs.
+We demonstrate in a simulation study and two real  case studies that this approach retains the flexibility of the B-splines, achieves similar ability to accurately estimate peaks due to a new data-driven knot allocation scheme but significantly reduces the computational costs.
 
@@
 
-\keywords{P-splines \and B-splines \and \textcolor{red}{Bernstein-Dirichlet process prior}  \and spectral density estimation \and Whittle likelihood}
+\keywords{P-splines \and B-splines \and Bernstein-Dirichlet process prior  \and spectral density estimation \and Whittle likelihood}
 
@@
 Most Bayesian approaches to psd estimation use the so-called Whittle likelihood, introduced by \cite{Whittle:1957} as an approximation to the likelihood of a Gaussian stationary time series.
-Based on the  limiting independent exponential distribution of the periodogram \textcolor{red}{values}, the Whittle likelihood for a mean centred weakly stationary time series $\bm{Y}$ of length $n$ is defined as
+Based on the  limiting independent exponential distribution of the periodogram values, the Whittle likelihood for a mean centred weakly stationary time series $\bm{Y}$ of length $n$ is defined as
 \begin{align}
@@
 where $\lambda_l = 2\pi l / n$ are the Fourier frequencies, $f(\cdot)$ spectral density of $\bm{Y}$, and $\nu = \left\lfloor (n-1)/2 \right\rfloor$ is the greatest integer value less than or equal to $(n-1)/2$.
-Advantages of this approximation \textcolor{red}{are} that it depends directly on the spectral density, unlike the true Gaussian likelihood function, and gives a good approximation for non-Gaussian time series under certain conditions as shown by \cite{ShaoXiaofeng2007ASTf}.
+Advantages of this approximation are that it depends directly on the spectral density, unlike the true Gaussian likelihood function, and gives a good approximation for non-Gaussian time series under certain conditions as shown by \cite{ShaoXiaofeng2007ASTf}.
 
@@
 A reversible jump Markov chain Monte Carlo \citep[RJMCMC;][]{Green:1995} algorithm was implemented to sample from the posterior distribution.
-\textcolor{red}{Similarly, frequentist approaches such as \cite{Rodriguez} and \cite{Wood2017} model the log spectral density by a linear
-combination of splines with variable knots and combine these with adaptive smoothing. }
+Similarly, frequentist approaches such as \cite{Rodriguez} and \cite{Wood2017} model the log spectral density by a linear combination of splines with variable knots and combine these with adaptive smoothing.
 
-\textcolor{red}{Penalized splines have been used for a time-domain analysis of time series data e.g.\ to model the mean as a smooth function of time by
-\cite{Krivobokova}  or for fitting and  forecasting univariate  nonlinear time series by \cite{Wegener}, but not in the frequency domain for estimating the spectral density.}
+Penalized splines have been used for a time-domain analysis of time series data e.g.\ to model the mean as a smooth function of time by \cite{Krivobokova}  or for fitting and  forecasting univariate  nonlinear time series by \cite{Wegener}, but not in the frequency domain for estimating the spectral density.
 The P-spline prior on the spectral density function that we are going to propose is related to the 
@@
 
-With the aim of reducing the computational cost of the BspDP prior approach,  we suggest to use the idea of P-splines of \cite{Eilers:1996} that \textcolor{red}{combines} a large but fixed number of  B-splines with a simple difference penalty that controls the degree of  smoothness of the spectral density.
+With the aim of reducing the computational cost of the BspDP prior approach,  we suggest to use the idea of P-splines of \cite{Eilers:1996} that combines a large but fixed number of  B-splines with a simple difference penalty that controls the degree of  smoothness of the spectral density.
 P-splines, i.e.\  equally-spaced B-splines,  have been successfully implemented for frequentist nonparametric regression but to the best of our knowledge not yet for Bayesian spectral density estimation.
@@
 In Section IV, we test our P-spline-based approaches to spectral density estimation and compare results to those based on the BspDP prior in a simulation study and the application to two real time series.
-The paper concludes with a discussion of the relative merits of the new approaches \textcolor{red}{and avenues for future research.}
+The paper concludes with a discussion of the relative merits of the new approaches and avenues for future research.
 
@@
 \section{P-spline prior}
-\textcolor{red}{
+
 As the spectral density function $f(\cdot)$ is defined on the interval $[0,\pi]$, it is reparametrized and defined as
@@
 \end{align}
-with  {\em fixed} number $K$ of B-spline densities  $b_{k,r}(\cdot)$ of fixed degree $r$, weight vector $\bm{w}=(w_1,\cdots, w_K)$, and knot sequence $\bm{\xi}$.}
+with  {\em fixed} number $K$ of B-spline densities  $b_{k,r}(\cdot)$ of fixed degree $r$, weight vector $\bm{w}=(w_1,\cdots, w_K)$, and knot sequence $\bm{\xi}$.
 
@@
 
-$\bm{\xi}$ contains fixed equidistant internal knots on $[0,1]$ in our basic P-spline prior. A different  allotment of knots is described in  section ``\textcolor{red}{Adaptive knots}".
+$\bm{\xi}$ contains fixed equidistant internal knots on $[0,1]$ in our basic P-spline prior. A different  allotment of knots is described in  section ``Adaptive knots".
 %Therefore, the P-spline prior on the spectral density function is given by
@@
 where $\bm{v} = (v_1,\dots,v_{K-1})^\top$ is a $K-1$ dimensional parameter vector, $\phi$ is the smoothing or penalty parameter, $\textbf{P} = \textbf{D}^\top \textbf{D} +\epsilon \textbf{I}_{K-1}$ is the penalty matrix (which is a full matrix rank matrix for any small quantity $\epsilon$, for instance, $10^{-6}$) with $\textbf{D}$ the $d^{\text{th}}$ order difference matrix ($d=1,2,\ldots, K-2)$, $\alpha_{\phi}$ and $\alpha_{\delta}$ are shape parameters, $\delta \beta_{\phi}$ and $\beta_{\delta}$ are rate parameters, and $\text{Gamma}(a,b)$ denotes a gamma distribution with mean $a/b$ and variance $a/b^2$.
-We used \textcolor{red}{first and second}  order difference penalties in the simulation study.
+We used first and second order difference penalties in the simulation study.
 The $1^{\text{st}}$ order difference matrix \textbf{D} $\in \mathbb{R}^{(K-2)\times (K-1)}$  is defined as
@@
 \end{equation}
-\textcolor{red}{and} the $2^{\text{nd}}$ order difference matrix $\in \mathbb{R}^{(K-3)\times (K-1)} $ as
+and the $2^{\text{nd}}$ order difference matrix $\in \mathbb{R}^{(K-3)\times (K-1)} $ as
 \begin{equation}\label{eq:D2}
@@
 %$\in \mathbb{R}^{(K-4)\times (K-1)}$.  
-\textcolor{red}{The rationale for using discrete second order roughness penalties as an approximation to the usual continuous roughness measure based on the integral over the second derivatives is given
-in \cite{Eilers:1996}.}
+The rationale for using discrete second order roughness penalties as an approximation to the usual continuous roughness measure based on the integral over the second derivatives is given
+in \cite{Eilers:1996}.
 
@@
 The B-spline prior method is able to handle abrupt changes because the knot location is variable and  driven by the data.
-We propose \textcolor{red}{an adaptive} method that allocates the knots according to the periodogram but once allocated, the knots stay fixed.
+We propose an adaptive method that allocates the knots according to the periodogram but once allocated, the knots stay fixed.
 The idea is to concentrate more effort in those regions that have potential peaks as detected by the periodogram.
 
-Our \textcolor{red}{adaptive} knot location proposal works as follow.
+Our adaptive knot location proposal works as follow.
 We calculate the periodogram and apply a square root transformation in order to have more regular magnitudes and eliminate a potential trend.
@@
 Thus, in those areas in which the periodogram has sharp and abrupt changes, our procedure will assign more knots in proportion to their magnitudes.
-Hereafter, we will refer to this scheme as \textcolor{red}{\em adaptive knots}.%\textit{periodogram-spaced knots} or short \textit{P-spaced knots}.
+Hereafter, we will refer to this scheme as {\em adaptive knots}.%\textit{periodogram-spaced knots} or short \textit{P-spaced knots}.
 %REMARK: HOW ABOUT "PERIODOGRAM-SPACED KNOTS"	OR 'P-SPACED KNOTS'  INSTEAD OF DISTRIBUTED KNOTS?
@@
 	\item Let $\bm{\xi}^* = \left(\xi_1^*,\dots,\xi_{K^*}^* \right)$ be a vector of $K^*$ equally spaced values in $[0,1]$.
-	\item The \textcolor{red}{adaptive} knot vector is given by $\bm{\xi} = \left(F_{z}^{-1}(\xi^*_1),\dots, F_{z}^{-1}(\xi^*_{K^*}) \right)$.
+	\item The adaptive knot vector is given by $\bm{\xi} = \left(F_{z}^{-1}(\xi^*_1),\dots, F_{z}^{-1}(\xi^*_{K^*}) \right)$.
 
@@
 
-\textcolor{red}{For non-equidistant knots, it does not make sense any longer to base the definition of the covariance matrix of the Gaussian prior  on the  difference matrices as defined in (\ref{eq:D1}) and (\ref{eq:D2}) because this is an adequate approximation of the derivative-based roughness measures -- defined as the integral of the first and second order derivatives of the P-splines --
+For non-equidistant knots, it does not make sense any longer to base the definition of the covariance matrix of the Gaussian prior  on the  difference matrices as defined in (\ref{eq:D1}) and (\ref{eq:D2}) because this is an adequate approximation of the derivative-based roughness measures -- defined as the integral of the first and second order derivatives of the P-splines --
 for equally spaced knots only.
@@
 This penalty matrix contains the inner products of the respective first and second order derivatives of the basis functions.
-In addition, we normalise this matrix by dividing it by the maximum absolute column sum.}
+In addition, we normalise this matrix by dividing it by the maximum absolute column sum.
 
-The \textcolor{red}{adaptive}  knot allocation can be regarded as an empirical Bayes prior since the data are used to specify the prior distribution.
+The adaptive knot allocation can be regarded as an empirical Bayes prior since the data are used to specify the prior distribution.
 However, only the locations of the B-splines are determined by the data, not the parameters of the prior distribution which in this case are the weights $w_k$ associated with each B-spline $b_{k,r}$.
@@
 A total of 100,000 MCMC samples were generated
-using  a pilot run of \textcolor{red}{20,000} iterations with a burn-in period of 5,000 and thinning factor of 10 to calibrate the proposals.
-The final sample consisted of \textcolor{red}{80,000} samples with a burn-in period of 5,000 and thinning factor of 10, resulting in \textcolor{red}{7,500} samples used for posterior inferences.
+using  a pilot run of 20,000 iterations with a burn-in period of 5,000 and thinning factor of 10 to calibrate the proposals.
+The final sample consisted of 80,000 samples with a burn-in period of 5,000 and thinning factor of 10, resulting in 7,500 samples used for posterior inferences.
 %Therefore, the number of iterations equate the B-spline analysis.  
-This procedure was replicated for both equidistant and \textcolor{red}{adaptive} knots \textcolor{red}{ considering first and second order penalties.}
+This procedure was replicated for both equidistant and adaptive knots considering first and second order penalties.
 
@@
 A larger penalty order does not necessarily result in more accurate psd estimates.
-In this case of a relatively smooth spectral density, the \textcolor{red}{adaptive} knots do not yield an improvement over the equidistant knots and the BspDP prior.
+In this case of a relatively smooth spectral density, the adaptive knots do not yield an improvement over the equidistant knots and the BspDP prior.
 
@@
 
-For the AR(4) time series, the \textcolor{red}{fit based on the P-spline prior with \textcolor{red}{adaptive} knots is better than the fit based on the B-spline Dirichlet process prior and the one based on the P-spline prior with equidistant knots}.
+For the AR(4) time series, the fit based on the P-spline prior with adaptive knots is better than the fit based on the B-spline Dirichlet process prior and the one based on the P-spline prior with equidistant knots.
 % categorically outperforms the BspDP algorithm.   It also outperforms the equidistant knot scheme. 
  In general, the BspDP algorithm achieves higher accuracy than the P-spline method with equidistant knots.
-As shown by \cite{Edwards2019}, the \textcolor{red}{posterior distribution based on the BspDP prior} yields better results than \textcolor{red}{posterior based on } the BDP in terms of median IAE in this example.
-Consequently, our P-spline algorithm with \textcolor{red}{adaptive} knots outperforms the BDP method as well.
+As shown by \cite{Edwards2019}, the posterior distribution based on the BspDP prior yields better results than posterior based on the BDP in terms of median IAE in this example.
+Consequently, our P-spline algorithm with adaptive knots outperforms the BDP method as well.
 
@@
 			\hspace{0.5em}P-spline $d=2$&  0.708 & 0.604 & 0.629 \\
-			\textit{\textcolor{red}{Adaptive} knots}& & &   \\
+			\textit{Adaptive knots}& & &   \\
 			\hspace{0.5em}P-spline $d=1$ & 0.874 & 0.783 & 0.610 \\
@@
 			\hspace{0.5em}P-spline $d=2$&  3.150 & 2.569 & 2.388\\
-			\textit{\textcolor{red}{Adaptive} knots}& & &   \\
+			\textit{Adaptive knots}& & &   \\
 			\hspace{0.5em}P-spline $d=1$ & 2.495 & 1.830 & 1.489\\
@@
 For the AR(1) model, the BspDP and first order penalty P-spline 90\% credible bands cover the whole true spectral density function, i.e.\ for all frequencies $\lambda$, in 
-\textcolor{red}{almost} all the 300 independent analyses.
+almost all the 300 independent analyses.
 In general, the latter method shows diminishing  performance as the penalty degree increases.
 For the AR(4) model, the P-spline method with equidistant knots performs poorly.
-On the other hand, the \textcolor{red}{adaptive} knot scheme yields evidently much better results, which are comparable (in the case of the first order penalty) to those obtained via the BspDP algorithm.
+On the other hand, the adaptive knot scheme yields evidently much better results, which are comparable (in the case of the first order penalty) to those obtained via the BspDP algorithm.
 %For $n=\{128,256\}$, the latter has higher coverage probabilities than our proposal, except for $n=512$.
@@
 As shown by \cite{Edwards2019}, the BspDP algorithm produces better results than the BDP prior in the AR(4) case, which coverage probability is zero across $n=\{128,256,512\}$.
-Consequently, our P-spline algorithm with \textcolor{red}{adaptive} knots outperforms the BDP method in this case.
+Consequently, our P-spline algorithm with adaptive knots outperforms the BDP method in this case.
 
@@
 Table~\ref{table:sim_covPbb_median},  only for a small fraction of the frequencies the psd is not covered by the uniform credible band.
-In this context, BspDP and \textcolor{red}{adaptive} knot P-spline algorithms perform equally well.
+In this context, BspDP and adaptive knot P-spline algorithms perform equally well.
 
@@
 			\hspace{0.5em}P-spline $d=2$& 0.993 & 1.000 & 0.993  \\
-			\textit{\textcolor{red}{Adaptive} knots}& & &   \\
+			\textit{Adaptive knots}& & &   \\
 			\hspace{0.5em}P-spline $d=1$ & 0.997 & 1.000 & 0.997 \\
@@
 			\hspace{0.5em}P-spline $d=2$ & 0.323 & 0.167 & 0.007\\
-			\textit{\textcolor{red}{Adaptive} knots}& & &   \\	
+			\textit{Adaptive knots}& & &   \\	
 			\hspace{0.5em}P-spline $d=1$ & 0.790 & 0.847 & 0.813 \\
@@
 			\textit{Equidistant \&}& & &   \\
-			\textit{\textcolor{red}{Adaptive} knots}& & &   \\
+			\textit{Adaptive knots}& & &   \\
 			\hspace{0.5em}P-spline $d=1,2$ & 1.000 & 1.000 & 1.000  \\ \hline	 		      
@@
 			\hspace{0.5em}P-spline $d=2$&  0.984 & 0.984 & 0.980 \\
-			\textit{\textcolor{red}{Adaptive} knots}& & &   \\	
+			\textit{Adaptive knots}& & &   \\	
 			\hspace{0.5em}P-spline $d=1$ & 1.000 & 1.000 & 1.000 \\
@@
 This is directly reflected in the run-time, which is approximately 2-3 times higher in this example (see Table~3 in \cite{Edwards2019}).
-The median run-time in our analyses for the BspDP prior and the P-spline with \textcolor{red}{adaptive} knots are displayed in Table~\ref{table:sim_time}.
-Note that the P-spline method based on equidistant knots has similar run-times to the \textcolor{red}{adapative} knots.
-The P-spline method is approximately \textcolor{red}{4-7} times faster than BspDP approach in these examples.
+The median run-time in our analyses for the BspDP prior and the P-spline with adaptive knots are displayed in Table~\ref{table:sim_time}.
+Note that the P-spline method based on equidistant knots has similar run-times to the adaptive knots.
+The P-spline method is approximately 4-7 times faster than BspDP approach in these examples.
 This is due to a reduced number of calculations  per iteration because the B-spline densities are calculated only  once at the start of the algorithm and there are no Dirichlet process approximations via the stick-breaking method.
@@
 			B-spline &   29.497 & 32.423 & 37.896 \\
-			\textit{\textcolor{red}{Adaptive} knots}& & &   \\
+			\textit{Adaptive knots}& & &   \\
 			\hspace{0.5em}P-spline $d=1$&  4.859 & 7.450 & 9.606\\
@@
 			B-spline & 30.111 & 34.126 & 40.875  \\
-			\textit{\textcolor{red}{Adaptive} knots}& & &   \\
+			\textit{Adaptive knots}& & &   \\
 			\hspace{0.5em}P-spline $d=1$ & 4.883 & 7.453 & 9.665 \\
@@
 
-In general terms, the P-spline algorithm with our novel knot allocation scheme, the \textcolor{red}{adaptive} knots, offers the best trade-off between accuracy and computation cost.
+In general terms, the P-spline algorithm with our novel knot allocation scheme, the adaptive knots, offers the best trade-off between accuracy and computation cost.
 The  rule-of-thumb criterion,  $K=\min\left\{n/4, 40\right\}$, for the number of B-spline densities in the P-spline prior algorithm worked perfectly well in this example.
@@
 Thus, 5,000 samples are used for the estimation.
-This took about \textcolor{red}{22} minutes.
+This took about 22 minutes.
 The other specifications are the same as those used in the simulation study.
@@
 Then the algorithm is run for 75,000 iterations with a burn-in period of 25,000 and thinning factor of 10, resulting in 5,000 samples.
-The analysis is performed using the first order penalty for equidistant and \textcolor{red}{adaptive} knots.
+The analysis is performed using the first order penalty for equidistant and adaptive knots.
 All other specifications are the same as those used in the simulation study.
-A single analysis takes about \textcolor{red}{4.5} minutes.
+A single analysis takes about 4.5 minutes.
 
@@
 The continuous and dashed lines stand for the posterior median log psd obtained using the B-spline and P-spline priors, respectively.
-The dot lines represents the 90\% pointwise credible bands for the latter prior, which consider a first order penalty and \textcolor{red}{adaptive} knots.}
+The dot lines represents the 90\% pointwise credible bands for the latter prior, which consider a first order penalty and adaptive} knots.
 	\label{fig:sunspot}
@@
 
-The results for the BspDP and P-spline prior  with first order penalty and \textcolor{red}{adaptive}  knots are displayed in Figure~\ref{fig:sunspot} \textcolor{red}{(p-spline with equidistant knot approach results are not shown due to their similarity to the adaptive knot estimates)}.
+The results for the BspDP and P-spline prior  with first order penalty and adaptive knots are displayed in Figure~\ref{fig:sunspot} (p-spline with equidistant knot approach results are not shown due to their similarity to the adaptive knot estimates).
 Both methods reveal a large peak at about 0.09.
@@
 These results are consistent with those obtained via the BDP  prior~\citep{Choudhuri:2004}.
-However, the BspDP estimation takes approximately \textcolor{red}{22} minutes compared to only \textcolor{red}{4.5} minutes for the the P-spline estimation.
-Thus the P-spline estimation is approximately
+However, the BspDP estimation takes approximately 22 minutes compared to only 4.5 minutes for the the P-spline estimation. Thus the P-spline estimation is approximately
 five times more efficient.
@@
 In addition, the data has been squared root transformed and mean centred.
-\textcolor{red}{ The first 150 observations are displayed in Figure~\ref{fig:carinae_data}, where the missing values are replaced by the mean.}
+The first 150 observations are displayed in Figure~\ref{fig:carinae_data}, where the missing values are replaced by the mean.
 
@@
 	\includegraphics[scale=0.4,clip=true,angle=0]{carinae_data.pdf}
-	\caption{ \textcolor{red}{First 150 observations of the \textit{S. Carinae} data.}}
+	\caption{First 150 observations of the \textit{S. Carinae} data.}
 	\label{fig:carinae_data}
@@
 We use 40 B-spline densities,  based on the rule-of-thumb criterion discussed above, and the first order penalty.
-The whole process takes less than \textcolor{red}{1.5} minutes to run and is replicated for equidistant and \textcolor{red}{adaptive} knots.
+The whole process takes less than 1.5 minutes to run and is replicated for equidistant and adaptive knots.
 The results are displayed in Figure~\ref{fig:carinae}.
@@
 	\caption{Log-spectral density estimate for the variable star {\it S.\ Carinae} data based on the P-spline prior.
-The continuous grey line represents the periodogram whereas the continuous and dashed black lines stand for the posterior median log psd obtained using \textcolor{red}{adaptive} and equidistant knots, respectively.}
+The continuous grey line represents the periodogram whereas the continuous and dashed black lines stand for the posterior median log psd obtained using adaptive and equidistant knots, respectively.}
 	\label{fig:carinae}
@@
 The periodogram shows several peaks with the main one located at around 0.42.  When the knots are equally spaced, the algorithm is unable to capture the sharpness of this peak,
-in contrast to the \textcolor{red}{adaptive} knots.
+in contrast to the adaptive knots.
 This is because the periodogram-based knot allocation puts more knots in those regions which contain significant peaks as indicated by the periodogram.
-The rest of peaks, which are smaller in comparison to the main one, are slightly better captured by the equidistant knots, since a large proportion of the \textcolor{red}{adaptive} knots have been allocated around the main peak.
+The rest of peaks, which are smaller in comparison to the main one, are slightly better captured by the equidistant knots, since a large proportion of the adaptive knots have been allocated around the main peak.
 However, this can easily  be corrected by increasing the number of knots.
-The psd estimate based on \textcolor{red}{adaptive} knots is comparable to the one obtained via the NPC approach \citep{Kirch:2018}.
+The psd estimate based on adaptive knots is comparable to the one obtained via the NPC approach \citep{Kirch:2018}.
 
@@
 In our simulation study, we showed that the equidistant knot scheme for the P-spline method works quite well in the case of simple spectral structures (AR(1) model).
-However, its performance is surpassed by the \textcolor{red}{adaptive} knot scheme for complex structures (AR(4) model).
-In this scenario particularly, we can highlight the good performance of the \textcolor{red}{adaptive} knot scheme.
+However, its performance is surpassed by the adaptive knot scheme for complex structures (AR(4) model).
+In this scenario particularly, we can highlight the good performance of the adaptive knot scheme.
 This method outperforms the BspDP prior in terms of IAE and run-times, particularly for the complex psds.
 Regarding uniform coverage,  a {\em first order} penalty P-spline coupled with
-\textcolor{red}{adaptive} knots works best. For this combination, the uniform coverage almost reaches the nominal level.
+adaptive knots works best. For this combination, the uniform coverage almost reaches the nominal level.
 
@@
 
-We also assessed our proposal in the classic sunspot dataset, in which the \textcolor{red}{posterior distribution based on the BspDP prior }%B-spline prior algorithm
+We also assessed our proposal in the classic sunspot dataset, in which the posterior distribution based on the BspDP prior %B-spline prior algorithm
  is able to estimate correctly the solar cycle that occurs every 11.07 years.
-The \textcolor{red}{posterior distribution based on the} P-spline prior yields almost identical results, but in  significantly  less computational time.
+The posterior distribution based on the P-spline prior yields almost identical results, but in  significantly  less computational time.
 
-Finally, we assess the \textcolor{red}{ fit of the posterior distribution based on the} P-spline prior a under two different knot location schemes: equally spaced and \textcolor{red}{adaptive} knots.
+Finally, we assess the fit of the posterior distribution based on the P-spline prior a under two different knot location schemes: equally spaced and adaptive knots.
 For this, we estimated the psd for the {\it S.\ Carinae} time series which has a spectral density with several sharp peaks.
-Equidistant knots failed to capture the peaks of this spectral density whereas  \textcolor{red}{adaptive} knots significantly improved  the results without affecting the computational time.
+Equidistant knots failed to capture the peaks of this spectral density whereas adaptive knots significantly improved  the results without affecting the computational time.
 
@@
 
-\textcolor{red}{As already reviewed in the introduction, many Bayesian nonparametric approaches to spectral density estimation propose to put a prior on the log-spectral density.
+As already reviewed in the introduction, many Bayesian nonparametric approaches to spectral density estimation propose to put a prior on the log-spectral density.
 This could be easily implemented by modifying (\ref{eq:ps_prior}) to
  \[ \log f(\pi \omega)=\sum_{k=1}^K v_k b_{k,r}(\omega,{\bm \xi})\]
-and it could be interesting to examine for what class of spectral densities this prior specification would yield better posterior inference. }
+and it could be interesting to examine for what class of spectral densities this prior specification would yield better posterior inference.
 
-\textcolor{red}{Another interesting question is whether the P-spline prior can be extended to enable a Bayesian analysis of multivariate time series. In this case, the spectral density is a 
+Another interesting question is whether the P-spline prior can be extended to enable a Bayesian analysis of multivariate time series. In this case, the spectral density is a 
 Hermitian positive definite  matrix-valued function. A novel Bayesain nonparametric approach to multivariate spectral density estimation was proposed by
@@
 An avenue for future research will be to explore whether a computational speed-up might be possible by making use of P-splines in smoothing the elements of the
-Cholesky decomposition of the spectral density matrix.}
+Cholesky decomposition of the spectral density matrix.
 
-In conclusion, the P-spline prior with the \textcolor{red}{adaptive} knot scheme offers a more computationally viable alternative to the BspDP prior for spectral density estimation.
+In conclusion, the P-spline prior with the adaptive knot scheme offers a more computationally viable alternative to the BspDP prior for spectral density estimation.
 As shown in this work, it can handle spectral densities with several peaks in a reasonable run-time.
 
-\textcolor{red}{The \textsf{R} package \texttt{psplinePsd} can be downloaded from \href{https://github.com/pmat747/psplinePsd}{GitHub}.}
+The \textsf{R} package \texttt{psplinePsd} can be downloaded from \href{https://github.com/pmat747/psplinePsd}{GitHub}.
 %The \textsf{R} package \texttt{psplinePsd} will be available on CRAN in due course.  In the meanwhile, it can be downloaded from \href{https://github.com/pmat747/psplinePsd}{GitHub}.
@@
 \begin{acknowledgements}
-	We thank the New Zealand eScience Infrastructure (NeSI) for their high performance computing facilities, and the Centre for eResearch at the University of Auckland for their technical support.
+We thank Thomas Yee for helpful discussions on p-splines.	
+We also thank the New Zealand eScience Infrastructure (NeSI) for their high performance computing facilities, and the Centre for eResearch at the University of Auckland for their technical support.
 PM's and RM's work is supported by Grant 3714568 from the University of Auckland Faculty Research Development Fund and the DFG Grant KI 1443/3-1.

### 2020-07-01 pSplines.tex
some edits

@@
 
-\title{Bayesian spectral density estimation using P-splines}
+\title{Bayesian spectral density estimation using P-splines with quantile-based knot placement}
 \author{Patricio Maturana-Russel \and Renate Meyer}
@@
 Our proposal is motivated by the B-spline Dirichlet process prior of \cite{Edwards2019} in combination  with Whittle's likelihood and aims at  reducing the high computational complexity of its posterior computations.
-The strength of the B-spline Dirichlet process prior over the Bernstein-Dirichlet process prior of \cite{Choudhuri:2004}  lies in its ability to estimate spectral densities with sharp peaks and abrupt changes due to the flexibility of B-splines with variable number and location of knots.
-Here, we suggest to use P-splines of \cite{Eilers:1996} that combine a B-spline basis with a discrete penalty on the basis coefficients.
-We demonstrate in a simulation study and two real  case studies that this approach retains the flexibility of the B-splines, achieves similar ability to accurately estimate peaks due to a new data-driven knot allocation scheme but significantly reduces the computational costs.
+The strength of the B-spline Dirichlet process prior over the Bernstein-Dirichlet process prior of \cite{Choudhuri:2004}  lies in its ability to estimate spectral densities with sharp peaks and abrupt changes due to the flexibility of B-splines with variable number and location of knots.Here, we suggest to use P-splines of \cite{Eilers:1996} that combine a B-spline basis with a discrete penalty on the basis coefficients.
+\textcolor{red}{In addition to equidistant knots, a novel strategy for a more expedient placement of  knots is proposed that makes use of the information provided by the periodogram about the steepness of the spectral power distribution.}
+We demonstrate in a simulation study and two real  case studies that this approach retains the flexibility of the B-splines, achieves similar ability to accurately estimate peaks due to the new data-driven knot allocation scheme but significantly reduces the computational costs.
 
@@
 who used a smoothing prior on the log psd.
-\cite{Rosen:2012} decompose the log psd into a linear and nonlinear component and  put a Gaussian process prior on the nonlinear part, \cite{Pensky:2007} propose Bayesian wavelet-smoothing of the log psd and recently, \cite{Cadonna2017} modelled the log psd  as a mixture of Gaussian distributions with frequency-dependent weights and mean functions.
+\cite{Rosen:2012} segmented a nonstationary time series into stationary segments.
+\textcolor{red}{For each stationary segment, they decomposed the log psd into a linear and nonlinear component and  put a linear smoothig spline prior on the nonlinear part.
+}\cite{Pensky:2007} propose Bayesian wavelet-smoothing of the log psd and recently, \cite{Cadonna2017} modelled the log psd  as a mixture of Gaussian distributions with frequency-dependent weights and mean functions.
 %
@@
 \cite{Gangopadhyay:1999}, the specification of the number and location of the B-spline densities is data-driven, however it avoids a RJMCMC algorithm by putting a second Dirichlet process prior  on the cdf that induces the inter-knot distances.
+\textcolor{red}{ \cite{Choudhuri:2004} already demonstrated in a comprehensive simulation study  that the Bayesian psd estimates based on the BDP prior outperforms in terms of the $L_1$-error the smoothed periodogram estimates based on Bartlett-Priestley quadratic kernel, the penalized MLE of \cite{Pawitan} and autoregressive spectral estimates in most scenarios but noted that the BDP-based estimates detect peaks correctly but often underestimate the magnitude of sharp peaks.}
 Unlike the beta densities  which have global support on the interval [0,1], B-spline densities have local support, thereby increasing their flexibility and allowing better modelling of spectral densities with sharp peaks and abrupt changes.
@@
 
-$\bm{\xi}$ contains fixed equidistant internal knots on $[0,1]$ in our basic P-spline prior. A different  allotment of knots is described in  section ``Adaptive knots".
+$\bm{\xi}$ contains fixed equidistant internal knots on $[0,1]$ in our basic P-spline prior.
+\textcolor{red}{A more judicious scheme for the placement  of knots that not only speeds up posterior computations but improves the fit for peaked spectral densities is described in  section ``Quantile-based knot placement".}
 %Therefore, the P-spline prior on the spectral density function is given by
@@
 %\subsection*{Knot allocation strategy}
-\subsection*{Adaptive knots}
+\subsection*{Quantile-based knot placement}
 
@@
 The B-spline prior method is able to handle abrupt changes because the knot location is variable and  driven by the data.
-We propose an adaptive method that allocates the knots according to the periodogram but once allocated, the knots stay fixed.
+We propose a new selective method that allocates the knots according to quantiles of a periodogram-related distribution function but once allocated, the knots stay fixed.
 The idea is to concentrate more effort in those regions that have potential peaks as detected by the periodogram.
+\textcolor{red}{It is similar in spirit to the knot placement of O-splines \cite{WandOrmerod} in the regression context that make use of the empirical quantiles of the covariates.
+}
 
-Our adaptive knot location proposal works as follow.
+Our knot location proposal works as follow.
 We calculate the periodogram and apply a square root transformation in order to have more regular magnitudes and eliminate a potential trend.
@@
 Thus, in those areas in which the periodogram has sharp and abrupt changes, our procedure will assign more knots in proportion to their magnitudes.
-Hereafter, we will refer to this scheme as {\em adaptive knots}.%\textit{periodogram-spaced knots} or short \textit{P-spaced knots}.
+Hereafter, we will refer to this scheme as {\em  quantile-spaced (Q-spaced) knots}.%\textit{periodogram-spaced knots} or short \textit{P-spaced knots}.
 %REMARK: HOW ABOUT "PERIODOGRAM-SPACED KNOTS"	OR 'P-SPACED KNOTS'  INSTEAD OF DISTRIBUTED KNOTS?
@@
 	\item Define $z_i = \left|\dfrac{x_i - \bar{x}}{S_x} \right|$ where $\bar{x}$ denotes the average and $S_x$ the standard deviation of ${\bm x}$ and calculate its distribution function $F_z = P(z \leq z_i)$.
-	\item Let $\bm{\xi}^* = \left(\xi_1^*,\dots,\xi_{K^*}^* \right)$ be a vector of $K^*$ equally spaced values in $[0,1]$.
-	\item The adaptive knot vector is given by $\bm{\xi} = \left(F_{z}^{-1}(\xi^*_1),\dots, F_{z}^{-1}(\xi^*_{K^*}) \right)$.
+	\item Let $\bm{\xi}^* = \left(\xi_1^*,\dots,\xi_{K^*}^* \right)$ be a vector of $K^*$ quantiles, equally spaced in $[0,1]$.
+	\item The Q-spaced knot vector is given by $\bm{\xi} = \left(F_{z}^{-1}(\xi^*_1),\dots, F_{z}^{-1}(\xi^*_{K^*}) \right)$.
 
@@
 
-The adaptive knot allocation can be regarded as an empirical Bayes prior since the data are used to specify the prior distribution.
+The quantile based knot allocation can be regarded as an empirical Bayes prior since the data are used to specify the prior distribution.
 However, only the locations of the B-splines are determined by the data, not the parameters of the prior distribution which in this case are the weights $w_k$ associated with each B-spline $b_{k,r}$.
@@
 %Therefore, the number of iterations equate the B-spline analysis.  
-This procedure was replicated for both equidistant and adaptive knots considering first and second order penalties.
+This procedure was replicated for both equidistant and Q-spaced knots considering first and second order penalties.
 
@@
 A larger penalty order does not necessarily result in more accurate psd estimates.
-In this case of a relatively smooth spectral density, the adaptive knots do not yield an improvement over the equidistant knots and the BspDP prior.
+In this case of a relatively smooth spectral density, the Q-spaced knots do not yield an improvement over the equidistant knots and the BspDP prior.
 
@@
 
-For the AR(4) time series, the fit based on the P-spline prior with adaptive knots is better than the fit based on the B-spline Dirichlet process prior and the one based on the P-spline prior with equidistant knots.
+For the AR(4) time series, the fit based on the P-spline prior with Q-spaced knots is better than the fit based on the B-spline Dirichlet process prior and the one based on the P-spline prior with equidistant knots.
 % categorically outperforms the BspDP algorithm.   It also outperforms the equidistant knot scheme. 
@@
 As shown by \cite{Edwards2019}, the posterior distribution based on the BspDP prior yields better results than posterior based on the BDP in terms of median IAE in this example.
-Consequently, our P-spline algorithm with adaptive knots outperforms the BDP method as well.
+Consequently, our P-spline algorithm with Q-spaced knots outperforms the BDP method as well.
 
@@
 			\hspace{0.5em}P-spline $d=2$&  0.708 & 0.604 & 0.629 \\
-			\textit{Adaptive knots}& & &   \\
+			\textit{Q-spaced knots}& & &   \\
 			\hspace{0.5em}P-spline $d=1$ & 0.874 & 0.783 & 0.610 \\
@@
 			\hspace{0.5em}P-spline $d=2$&  3.150 & 2.569 & 2.388\\
-			\textit{Adaptive knots}& & &   \\
+			\textit{Q-spaced knots}& & &   \\
 			\hspace{0.5em}P-spline $d=1$ & 2.495 & 1.830 & 1.489\\
@@
 For the AR(4) model, the P-spline method with equidistant knots performs poorly.
-On the other hand, the adaptive knot scheme yields evidently much better results, which are comparable (in the case of the first order penalty) to those obtained via the BspDP algorithm.
+On the other hand, the quantile-based knot scheme yields evidently much better results, which are comparable (in the case of the first order penalty) to those obtained via the BspDP algorithm.
 %For $n=\{128,256\}$, the latter has higher coverage probabilities than our proposal, except for $n=512$.
@@
 As shown by \cite{Edwards2019}, the BspDP algorithm produces better results than the BDP prior in the AR(4) case, which coverage probability is zero across $n=\{128,256,512\}$.
-Consequently, our P-spline algorithm with adaptive knots outperforms the BDP method in this case.
+Consequently, our P-spline algorithm with Q-spaced knots outperforms the BDP method in this case.
 
@@
 Table~\ref{table:sim_covPbb_median},  only for a small fraction of the frequencies the psd is not covered by the uniform credible band.
-In this context, BspDP and adaptive knot P-spline algorithms perform equally well.
+In this context, BspDP and Q-spaced  knot P-spline algorithms perform equally well.
 
@@
 			\hspace{0.5em}P-spline $d=2$& 0.993 & 1.000 & 0.993  \\
-			\textit{Adaptive knots}& & &   \\
+			\textit{Q-spaced knots}& & &   \\
 			\hspace{0.5em}P-spline $d=1$ & 0.997 & 1.000 & 0.997 \\
@@
 			\hspace{0.5em}P-spline $d=2$ & 0.323 & 0.167 & 0.007\\
-			\textit{Adaptive knots}& & &   \\	
+			\textit{Q-spaced knots}& & &   \\	
 			\hspace{0.5em}P-spline $d=1$ & 0.790 & 0.847 & 0.813 \\
@@
 			\textit{Equidistant \&}& & &   \\
-			\textit{Adaptive knots}& & &   \\
+			\textit{Q-spaced knots}& & &   \\
 			\hspace{0.5em}P-spline $d=1,2$ & 1.000 & 1.000 & 1.000  \\ \hline	 		      
@@
 			\hspace{0.5em}P-spline $d=2$&  0.984 & 0.984 & 0.980 \\
-			\textit{Adaptive knots}& & &   \\	
+			\textit{Q-spaced knots}& & &   \\	
 			\hspace{0.5em}P-spline $d=1$ & 1.000 & 1.000 & 1.000 \\
@@
 This is directly reflected in the run-time, which is approximately 2-3 times higher in this example (see Table~3 in \cite{Edwards2019}).
-The median run-time in our analyses for the BspDP prior and the P-spline with adaptive knots are displayed in Table~\ref{table:sim_time}.
-Note that the P-spline method based on equidistant knots has similar run-times to the adaptive knots.
+The median run-time in our analyses for the BspDP prior and the P-spline with Q-spaced knots are displayed in Table~\ref{table:sim_time}.
+Note that the P-spline method based on equidistant knots has similar run-times to the Q-spaced knots.
 The P-spline method is approximately 4-7 times faster than BspDP approach in these examples.
@@
 			B-spline &   29.497 & 32.423 & 37.896 \\
-			\textit{Adaptive knots}& & &   \\
+			\textit{Q-spaced knots}& & &   \\
 			\hspace{0.5em}P-spline $d=1$&  4.859 & 7.450 & 9.606\\
@@
 			B-spline & 30.111 & 34.126 & 40.875  \\
-			\textit{Adaptive knots}& & &   \\
+			\textit{Q-spaced knots}& & &   \\
 			\hspace{0.5em}P-spline $d=1$ & 4.883 & 7.453 & 9.665 \\
@@
 
-In general terms, the P-spline algorithm with our novel knot allocation scheme, the adaptive knots, offers the best trade-off between accuracy and computation cost.
+In general terms, the P-spline algorithm with our novel quantile-based  knot allocation scheme offers the best trade-off between accuracy and computation cost.
 The  rule-of-thumb criterion,  $K=\min\left\{n/4, 40\right\}$, for the number of B-spline densities in the P-spline prior algorithm worked perfectly well in this example.
@@
 Then the algorithm is run for 75,000 iterations with a burn-in period of 25,000 and thinning factor of 10, resulting in 5,000 samples.
-The analysis is performed using the first order penalty for equidistant and adaptive knots.
+The analysis is performed using the first order penalty for equidistant and Q-spaced knots.
 All other specifications are the same as those used in the simulation study.
@@
 The continuous and dashed lines stand for the posterior median log psd obtained using the B-spline and P-spline priors, respectively.
-The dot lines represents the 90\% pointwise credible bands for the latter prior, which consider a first order penalty and adaptive} knots.
+The dot lines represents the 90\% pointwise credible bands for the latter prior, which consider a first order penalty and Q-spaced knots.}
 	\label{fig:sunspot}
@@
 
-The results for the BspDP and P-spline prior  with first order penalty and adaptive knots are displayed in Figure~\ref{fig:sunspot} (p-spline with equidistant knot approach results are not shown due to their similarity to the adaptive knot estimates).
+The results for the BspDP and P-spline prior  with first order penalty and Q-spaced knots are displayed in Figure~\ref{fig:sunspot} (p-spline with equidistant knot approach results are not shown due to their similarity to the Q-spaced knot estimates).
 Both methods reveal a large peak at about 0.09.
@@
 We use 40 B-spline densities,  based on the rule-of-thumb criterion discussed above, and the first order penalty.
-The whole process takes less than 1.5 minutes to run and is replicated for equidistant and adaptive knots.
+The whole process takes less than 1.5 minutes to run and is replicated for equidistant and Q-spaced knots.
 The results are displayed in Figure~\ref{fig:carinae}.
@@
 	\caption{Log-spectral density estimate for the variable star {\it S.\ Carinae} data based on the P-spline prior.
-The continuous grey line represents the periodogram whereas the continuous and dashed black lines stand for the posterior median log psd obtained using adaptive and equidistant knots, respectively.}
+The continuous grey line represents the periodogram whereas the continuous and dashed black lines stand for the posterior median log psd obtained using Q-spaced and equidistant knots, respectively.}
 	\label{fig:carinae}
@@
 The periodogram shows several peaks with the main one located at around 0.42.  When the knots are equally spaced, the algorithm is unable to capture the sharpness of this peak,
-in contrast to the adaptive knots.
+in contrast to the Q-spaced knots.
 This is because the periodogram-based knot allocation puts more knots in those regions which contain significant peaks as indicated by the periodogram.
-The rest of peaks, which are smaller in comparison to the main one, are slightly better captured by the equidistant knots, since a large proportion of the adaptive knots have been allocated around the main peak.
+The rest of peaks, which are smaller in comparison to the main one, are slightly better captured by the equidistant knots, since a large proportion of the Q-spaced knots have been allocated around the main peak.
 However, this can easily  be corrected by increasing the number of knots.
-The psd estimate based on adaptive knots is comparable to the one obtained via the NPC approach \citep{Kirch:2018}.
+The psd estimate based on Q-spaced knots is comparable to the one obtained via the NPC approach \citep{Kirch:2018}.
 
@@
 This approach does not affect the computational time and improves the estimates dramatically.
-%This also indicates that the BspDP computation time could be reduced by the choice of \textcolor{red}{adaptive} knots as initial values.
+%This also indicates that the BspDP computation time could be reduced by the choice of \textcolor{red}{Q-spaced} knots as initial values.
 
 In our simulation study, we showed that the equidistant knot scheme for the P-spline method works quite well in the case of simple spectral structures (AR(1) model).
-However, its performance is surpassed by the adaptive knot scheme for complex structures (AR(4) model).
-In this scenario particularly, we can highlight the good performance of the adaptive knot scheme.
+However, its performance is surpassed by the quantile-based  knot location scheme for complex structures (AR(4) model).
+In this scenario particularly, we can highlight the good performance of the quantile-based knot scheme.
 This method outperforms the BspDP prior in terms of IAE and run-times, particularly for the complex psds.
 Regarding uniform coverage,  a {\em first order} penalty P-spline coupled with
-adaptive knots works best. For this combination, the uniform coverage almost reaches the nominal level.
+Q-spaced knots works best. For this combination, the uniform coverage almost reaches the nominal level.
 
-%In our simulation study, we showed that, in general, the P-spline prior with \textcolor{red}{adaptive} knots outperforms the BspDP prior in terms of IAE and run-times, particularly for the complex psd.
+%In our simulation study, we showed that, in general, the P-spline prior with \textcolor{red}{Q-spaced} knots outperforms the BspDP prior in terms of IAE and run-times, particularly for the complex psd.
 However, the results were not conclusive for coverage probabilities.
@@
 
-Finally, we assess the fit of the posterior distribution based on the P-spline prior a under two different knot location schemes: equally spaced and adaptive knots.
+Finally, we assess the fit of the posterior distribution based on the P-spline prior a under two different knot location schemes: equally spaced and Q-spaced knots.
 For this, we estimated the psd for the {\it S.\ Carinae} time series which has a spectral density with several sharp peaks.
-Equidistant knots failed to capture the peaks of this spectral density whereas adaptive knots significantly improved  the results without affecting the computational time.
+Equidistant knots failed to capture the peaks of this spectral density whereas Q-spaced knots significantly improved  the results without affecting the computational time.
 
@@
 
-In conclusion, the P-spline prior with the adaptive knot scheme offers a more computationally viable alternative to the BspDP prior for spectral density estimation.
+In conclusion, the P-spline prior with the quantile-based knot placement scheme offers a more computationally viable alternative to the BspDP prior for spectral density estimation.
 As shown in this work, it can handle spectral densities with several peaks in a reasonable run-time.
@@
 \begin{acknowledgements}
-We thank Thomas Yee for helpful discussions on p-splines.	
+We thank Thomas Yee for helpful discussions on P-splines.	
 We also thank the New Zealand eScience Infrastructure (NeSI) for their high performance computing facilities, and the Centre for eResearch at the University of Auckland for their technical support.
 PM's and RM's work is supported by Grant 3714568 from the University of Auckland Faculty Research Development Fund and the DFG Grant KI 1443/3-1.
-RM gratefully acknowledges support by a James Cook Fellowship from Government funding, administered by teh Royal Society Te Ap\={a}rangi.
+RM gratefully acknowledges support by a James Cook Fellowship from Government funding, administered by the Royal Society Te Ap\={a}rangi.
 All analysis was conducted in \textsf{R}, an open-source statistical software available on \textsf{CRAN} (cran.r-project.org).

### 2020-07-05 pSplines.tex
starting value w text

@@
 Our proposal is motivated by the B-spline Dirichlet process prior of \cite{Edwards2019} in combination  with Whittle's likelihood and aims at  reducing the high computational complexity of its posterior computations.
-The strength of the B-spline Dirichlet process prior over the Bernstein-Dirichlet process prior of \cite{Choudhuri:2004}  lies in its ability to estimate spectral densities with sharp peaks and abrupt changes due to the flexibility of B-splines with variable number and location of knots.Here, we suggest to use P-splines of \cite{Eilers:1996} that combine a B-spline basis with a discrete penalty on the basis coefficients.
-\textcolor{red}{In addition to equidistant knots, a novel strategy for a more expedient placement of  knots is proposed that makes use of the information provided by the periodogram about the steepness of the spectral power distribution.}
+The strength of the B-spline Dirichlet process prior over the Bernstein-Dirichlet process prior of \cite{Choudhuri:2004}  lies in its ability to estimate spectral densities with sharp peaks and abrupt changes due to the flexibility of B-splines with variable number and location of knots.
+Here, we suggest to use P-splines of \cite{Eilers:1996} that combine a B-spline basis with a discrete penalty on the basis coefficients.
+In addition to equidistant knots, a novel strategy for a more expedient placement of  knots is proposed that makes use of the information provided by the periodogram about the steepness of the spectral power distribution.
 We demonstrate in a simulation study and two real  case studies that this approach retains the flexibility of the B-splines, achieves similar ability to accurately estimate peaks due to the new data-driven knot allocation scheme but significantly reduces the computational costs.
@@
 \cite{Rosen:2012} segmented a nonstationary time series into stationary segments.
-\textcolor{red}{For each stationary segment, they decomposed the log psd into a linear and nonlinear component and  put a linear smoothig spline prior on the nonlinear part.
-}\cite{Pensky:2007} propose Bayesian wavelet-smoothing of the log psd and recently, \cite{Cadonna2017} modelled the log psd  as a mixture of Gaussian distributions with frequency-dependent weights and mean functions.
+For each stationary segment, they decomposed the log psd into a linear and nonlinear component and  put a linear smoothig spline prior on the nonlinear part.
+\cite{Pensky:2007} propose Bayesian wavelet-smoothing of the log psd and recently, \cite{Cadonna2017} modelled the log psd  as a mixture of Gaussian distributions with frequency-dependent weights and mean functions.
 %
@@
 \cite{Gangopadhyay:1999}, the specification of the number and location of the B-spline densities is data-driven, however it avoids a RJMCMC algorithm by putting a second Dirichlet process prior  on the cdf that induces the inter-knot distances.
-\textcolor{red}{ \cite{Choudhuri:2004} already demonstrated in a comprehensive simulation study  that the Bayesian psd estimates based on the BDP prior outperforms in terms of the $L_1$-error the smoothed periodogram estimates based on Bartlett-Priestley quadratic kernel, the penalized MLE of \cite{Pawitan} and autoregressive spectral estimates in most scenarios but noted that the BDP-based estimates detect peaks correctly but often underestimate the magnitude of sharp peaks.}
+\cite{Choudhuri:2004} already demonstrated in a comprehensive simulation study  that the Bayesian psd estimates based on the BDP prior outperforms in terms of the $L_1$-error the smoothed periodogram estimates based on Bartlett-Priestley quadratic kernel, the penalized MLE of \cite{Pawitan} and autoregressive spectral estimates in most scenarios but noted that the BDP-based estimates detect peaks correctly but often underestimate the magnitude of sharp peaks.
 Unlike the beta densities  which have global support on the interval [0,1], B-spline densities have local support, thereby increasing their flexibility and allowing better modelling of spectral densities with sharp peaks and abrupt changes.
@@
 $\bm{\xi}$ contains fixed equidistant internal knots on $[0,1]$ in our basic P-spline prior.
-\textcolor{red}{A more judicious scheme for the placement  of knots that not only speeds up posterior computations but improves the fit for peaked spectral densities is described in  section ``Quantile-based knot placement".}
+A more judicious scheme for the placement  of knots that improves the fit for peaked spectral densities is described in  section ``Quantile-based knot placement".
 %Therefore, the P-spline prior on the spectral density function is given by
@@
 However, to sample from the full conditional distribution of $\bm{v}$, we use the Metropolis algorithm.
+\textcolor{red}{Since the weights in general are larger in those areas of the frequency domain where the psd has peaks, we specify their starting value proportionally to the periodogram.
+Applying the corresponding transformation we obtain the starting value for the vector $\bm{v}$.
+This strategy speeds up the MCMC process by shorting the burn-in period.}
+
 In order to improve the mixing of the chains, we apply the Metropolis algorithm on a reparametrized posterior distribution \citep{Lambert:2007}.
@@
 The idea is to concentrate more effort in those regions that have potential peaks as detected by the periodogram.
-\textcolor{red}{It is similar in spirit to the knot placement of O-splines \cite{WandOrmerod} in the regression context that make use of the empirical quantiles of the covariates.
-}
+It is similar in spirit to the knot placement of O-splines \citep{WandOrmerod} in the regression context that make use of the empirical quantiles of the covariates.
 

### 2020-10-06 responses/responses2/response.tex
revision

--- 2020-03-28 response.tex
@@
 We have now carefully revised the paper taking all of your -- and the other referees' -- comments into account and hope that you will now find the paper acceptable for publication.
-For your convenience, all changes made are marked in red in the manuscript.}
+For your convenience, all changes made are marked in red in the manuscript.
+All tables and figures have been modified.}
 
@@
 We have rectified this and now use derivative-based penalties for the unequally spaced knots.
-We describe the prior specification in paragraph 4  of the section  "adaptive knots" in Chapter 3.
+We describe the prior specification in paragraph 4  of the section  ``adaptive knots" in Chapter 3.
 As third order derivative-based penalties do not make sense for cubic splines and also because we want to compare the results for the equally-spaced knots and adaptive knots in the simulation study, we only compare results based on first and second order difference and derivative-based penalties, respectively.
@@
 
-We were wondering whether the referee meant "model the logarithm of the periodogram".
+We were wondering whether the referee meant ``model the logarithm of the periodogram".
 One might consider to model the log-periodogram instead of the periodogram itself which would require using the logarithm of an exponential distribution or approximating
@@
 
-\response{Data and example scripts are included in the R package {\tt psplinePsd}.}\bigskip
+\response{Data and example scripts are included in the R package \texttt{psplinePsd}.}\bigskip
 
@@
 \refereeQuote{
-The term 'P-spaced knots' is confusing in the context of P-splines, P is already used for penalties.
+The term `P-spaced knots' is confusing in the context of P-splines, P is already used for penalties.
 An adaptive scheme is proposed that is not exclusive to periodograms.
@@
 
-\response{ This could indeed be confusing. We have changed 'P-spaced' to 'adaptive' knots.}\bigskip
+\response{ This could indeed be confusing. We have changed `P-spaced' to `adaptive' knots.}\bigskip
 
@@
 \refereeQuote{A Metropolis step is used for v, but a data augmentation step based on Polya-Gamma auxiliary variables will allow for a full Gibbs sampler.
-See "Bayesian Inference for Logistic Models Using PolyaGamma Latent Variables" (2013) of Polson, Scott and Windle.}\smallskip
+See ``Bayesian Inference for Logistic Models Using PolyaGamma Latent Variables" (2013) of Polson, Scott and Windle.}\smallskip
 
-\response{We had already pointed this out in section "posterior computation" and had cited the paper by Polson et al. (2013).}
+\response{We had already pointed this out in section ``posterior computation" and had cited the paper by Polson et al. (2013).}
 

### 2020-10-06 responses/responses3/response.tex
revision

\documentclass{scrartcl}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{color}
\usepackage[english]{babel}
%\usepackage[fixlanguage]{babelbib}
\usepackage{natbib}
\usepackage[pdftex]{graphicx}
\usepackage{dsfont}
\usepackage{color}
\usepackage{hyperref}  % Package for hyperlinking
%\usepackage{a4wide}
\usepackage[a4paper, left=3cm, right=3cm, top=2cm]{geometry}
%\usepackage{texititel}
\title{Bayesian spectral density estimation using P-splines with quantile-based knot placement}
\subtitle{Response letter to Statistical Papers}
\author{Patricio Maturana-Russel, Renate Meyer}

\newcommand{\refereeQuote}{\textit }
\newcommand{\responseDraft}{\textcolor{red}}
\newcommand{\response}{}

\parindent 0pt
\parskip 4pt

\setlength{\oddsidemargin}{0cm}
\setlength{\evensidemargin}{0cm}
\setlength{\topmargin}{-2cm}

\pagestyle{empty}
\begin{document}
\maketitle\thispagestyle{empty}

% -----------------------------------------------------------------------------
% -----------------------------------------------------------------------------
% -----------------------------------------------------------------------------

\subsection*{Referee \#1}

\response{We would like to thank the referee for a thorough review of our paper. Below, you find a point-by-point response to your points raised. We have now carefully revised the paper taking all of your -- and the other referees' -- comments into account and hope that you will now find the paper acceptable for publication. For your convenience, all changes made are marked in red in the manuscript.}

\textbf{Major comment:}

\refereeQuote{p.8, Quantile-based knot placement\\
I could not understand the meaning of ``calculate its distribution function $F_z = P(z \leq z_i)$" in the second item of the procedure.  The values of the transformed periodogram $z_i$ are regarded as the values of the density $\tilde{f}$, which is non-monotonic, taken at equally-spaced frequencies.  According to item 4, the distribution function $F_z$, and hence the density function $\tilde{f}$, should be defined on the frequency domain. \\
I had a look at the R codes of the package psplinePsd. I think that the density function $\tilde{f}$ is a linear spline that interpolates as
\begin{align*}
\tilde{f} \left(\dfrac{\lambda_k}{2\pi}\right) = \dfrac{z_k}{\sum_{i=1}^{\nu}z_i}, \quad k=1,\dots,\nu.
\end{align*}
}

\response{That's correct!  The $\widetilde{f}$ is actually a probability mass function (not a probability density function), which is interpolated.  We have corrected this in the text.  We have also changed the description of the algorithm, following now the exact steps implemented in the \texttt{psplinePsd} package.}\bigskip

\textbf{Minor comments:}
\begin{enumerate}
\item
\refereeQuote{p.4, line 10\\
A spline of degree $r+1\implies$ A spline of degree $r$ (or, of order $r+1$)}\smallskip

\response{Fixed.}\bigskip

\item
\refereeQuote{p.5 line 13\\
In the definition of $\tau, f(\tau \omega)\implies f(\pi \omega)$}\smallskip

\response{That is correct.  Fixed. }\bigskip

\end{enumerate}

% -----------------------------------------------------------------------------
% -----------------------------------------------------------------------------
% -----------------------------------------------------------------------------

\subsection*{Referee \#2}

\response{We would like to thank the referee for a thorough review of our paper. Below, you find a point-by-point response to your points raised. We have now carefully revised the paper taking all of your -- and the other referees' -- comments into account and hope that you will now find the paper acceptable for publication. For your convenience changes are marked in red in the manuscript.
} \smallskip

\textbf{Specific comments:}
\begin{enumerate}
	\item \refereeQuote{ Page 5, how to choose penalty or prior $\phi$?}

	\response{In a Bayesian setting, the roughness penalty is translated into the prior distribution for the spline parameters via a normal distribution.  See for instance \cite{Bremhorst:2016}.  In order to make the posterior less sensitive to the prior, we suggest to follow their proposal and use their robust specifications for the hierarchical priors on $\phi$, which is also described in our paper. }

	\item \refereeQuote{To select the number of B-spline densities, the authors employ a heuristic rule of thumb $K^*$ in the application Section, it is of interest to compare it with other criteria such as AIC and BIC.}\smallskip

	\response{This could be a very interesting avenue for future research, however, we think it is beyond the scope of this paper.  An analysis of those characteristics can be quite extensive and could constitute a paper in itself, see for instance \cite{Likhachev2017}.  We would like to highlight that the number of knots is considered non-important by many, since the penalty parameter plus the penalty order difference penalties specified in the penalty matrix $\textbf{P}$ penalise the inclusion of B-spline densities, controlling the smoothness of the estimate.  This means that we only need to select a minimum value, the penalty terms take care of the rest.}\bigskip

	Other minor comments: \bigskip

	\item \refereeQuote{Page 2, line 17, what is $\lambda_i$?}\smallskip

	\response{$\lambda_i$ represents any angular frequency. Since we only defined the angular frequency $\lambda$ in that paragraph, we deleted its subscript ``$i$" and added $\lambda\in[-\pi,\pi]$. } \bigskip

	\item \refereeQuote{ Page 4, lines 5-7, ``In particular, the approximation error of the mixture of B-spline distributions can be made arbitrarily small by finding suitable knot locations and increasing the number of knots", what does ``made arbitrarily small" mean? Do you mean the approach with the beta densities the error cannot be made arbitrarily small?}\smallskip

	\response{Yes, the loss associated to the approximation given by the beta densities with respect to a loss function cannot be made arbitrarily small. This was shown by \cite{Perron:2001} who also showed that if one replaces the beta densities by B-spline densities of fixed order but with variable knots, the loss can be made arbitrarily small by increasing the number of knots.  We have added the corresponding reference to our manuscript.} \bigskip

	\item \refereeQuote{Page 4, line 15, ``A spline of degree r + 1 is a function", I think degree" should be order", we use degree for polynomials.}\smallskip

	\response{Fixed.} \bigskip

	\item \refereeQuote{Page 5, line 20, ``If $\tau = \int_0^1f(\tau \omega)d \omega$ is the normalizing constant", it implies $s_r$ is a density, is it correct?}

	\response{Yes, $s_r$ is a pdf.  Note that it is actually a mixture of B-spline densities.}
\end{enumerate}

\subsection*{Further Changes}
Here some extra changes:
\begin{enumerate}
	\item We have specified the prior on $\tau$. 
	\item We have added $s_r(\omega)$ in Eq(3), short notation used in the posterior distribution of~$\tau$.	
\end{enumerate}

\bibliographystyle{plainnat}
\bibliography{pSplinesReferences}  % Name of BibTeX file

%\begin{thebibliography}{38}
%	\providecommand{\natexlab}[1]{#1}
%	\providecommand{\url}[1]{\texttt{#1}}
%	\expandafter\ifx\csname urlstyle\endcsname\relax
%	\providecommand{\doi}[1]{doi: #1}\else
%	\providecommand{\doi}{doi: \begingroup \urlstyle{rm}\Url}\fi
%	\bibitem[Bremhorst and Lambert(2016)]{Bremhorst:2016}
%	Vincent Bremhorst and Philippe Lambert.
%	\newblock Flexible estimation in cure survival models using {B}ayesian
%	p-splines.
%	\newblock \emph{Computational Statistics \& Data Analysis}, 93:\penalty0 270 --
%	284, 2016.
%	\newblock ISSN 0167-9473.
%	\newblock \doi{https://doi.org/10.1016/j.csda.2014.05.009}.
%	\newblock URL
%	\url{http://www.sciencedirect.com/science/article/pii/S0167947314001492}.
%\end{thebibliography}

\end{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2020-10-06 pSplines.tex
revision

@@
 
-\newcommand{\pmr}{ \color{blue}}
+\newcommand{\pmr}{ \color{red}}
 
@@
 in one way or another, where $-\pi < \lambda \leq \pi$ is the angular frequency and $\bm{Y}=(Y_1,\ldots,Y_n)$ is a stationary time series of length $n$.
-The periodogram ordinates evaluated at the Fourier frequencies $\lambda_i$  are asymptotically independent and follow an exponential distribution with mean $f(\lambda_i)$.
-Thus, they are unbiased estimates of $f(\lambda_i)$  but not consistent \citep{Brockwell:1986}.
+{\pmr The periodogram ordinates evaluated at the Fourier frequencies $\lambda \in [-\pi,\pi]$  are asymptotically independent and follow exponential distributions with mean $f(\lambda)$.
+Thus, they are unbiased estimates of $f(\lambda)$  but not consistent \citep{Brockwell:1986}.}
 
@@
 L(\bm{Y} | f) &= \prod_{l=1}^{\nu}\dfrac{1}{f(\lambda_{l})} e^{-I_n(\lambda_l)/f(\lambda_l)}  \nonumber \\
-&\propto \exp \left( - \sum_{l=1}^{\nu} \log f(\lambda_l) + \dfrac{I_{n}(\lambda_l)}{f(\lambda_l)} \right),\label{eq:whittle_like}
+&\propto \exp \left\{ -\! \sum_{l=1}^{\nu} \Bigg( \log f(\lambda_l) + \dfrac{I_{n}(\lambda_l)}{f(\lambda_l)} \Bigg) \right\},\label{eq:whittle_like}
 \end{align}
@@
 As shown by \cite{Perron:2001} for distributions on the interval [0,1], the mixture of B-spline distributions has better approximation properties than the mixture of beta distributions for a fixed number of components.
-In particular, the approximation error of the mixture of B-spline distributions
-can be made arbitrarily small by finding  suitable knot locations and increasing the number of knots.
+In particular, the approximation error of the mixture of B-spline distributions can be made arbitrarily small by finding  suitable knot locations and increasing the number of knots {\pmr\citep{Perron:2001}}.
 %the mixture of beta distributions is unable to provide a full coverage of the space of cdfs on $[0,1]$.
@@
 
-A spline of degree $r+1$ is a function defined piecewise by polynomials of degree $\leq r$, which meet at points called \textit{knots} where the function is continuous.
+A spline of {\pmr order} $r+1$ is a function defined piecewise by polynomials of degree $\leq r$, which meet at points called \textit{knots} where the function is continuous.
 Any of these spline functions can be described by basis functions, known as \textit{B-splines}, in other words, all spline functions can be represented as a unique combination of B-splines with the same order and over the same partition.
@@
 \end{align*}
-where $\tau = \int_{0}^{1}f(\tau \omega)\text{d}\omega$ is the normalizing constant and
+where $\tau = \int_{0}^{1}f(\pi \omega)\text{d}\omega$ is the normalizing constant and
 \begin{align}
-s_r(\omega; \bm{w}, \bm{\xi}) = \sum_{k=1}^{K} w_k b_{k,r}(\omega;\bm{\xi}),
+s_r(\omega; \bm{w}, \bm{\xi}) = {\pmr s_r(\omega)} = \sum_{k=1}^{K} w_k b_{k,r}(\omega;\bm{\xi}),
 \label{eq:ps_prior}	
@@
 Here we set $\alpha_{\phi}$ and $\beta_{\phi}$ equal to 1 as used by \cite{Bremhorst:2016}.
-The scale parameter $\tau$ is given by an inverse Gamma($\alpha_{\tau},\beta_{\tau}$) prior.
-These are also the prior specifications used in the simulations and examples in Section \ref{sec:application}.
+The scale parameter $\tau$ is given by an inverse Gamma($\alpha_{\tau},\beta_{\tau}$) prior, {\pmr which can be considered as a noninformative prior for $\alpha_{\tau} = \beta_{\tau} = 0.001$ \citep{Edwards2019}.}  These are also the prior specifications used in the simulations and examples in Section \ref{sec:application}.
 
@@
 \delta|\bm{Y},\phi  &\sim \text{Gamma}\left(\alpha_{\phi} + \alpha_{\delta}, \beta_{\phi} \phi + \beta_{\delta}\right),\:\: \text{and}\\
-\tau|\bm{Y},\delta, \phi,\bm{v}   &\sim \text{Inv-Gamma}\!\left(\alpha_{\tau}+ \nu,\:  {\textstyle\sum\limits_{l=1}^{\nu}} \tfrac{I_n(\lambda_l)}{f(\lambda_l)} + \beta_{\tau} \right)\!.
+\tau|\bm{Y},\delta, \phi,\bm{v}   &\sim \text{Inv-Gamma}\!\left(\alpha_{\tau}+ \nu,\:  {\textstyle\sum\limits_{l=1}^{\nu}} \tfrac{I_n(\lambda_l)}{{\pmr s_r(\lambda_l/\pi)}} + \beta_{\tau} \right)\!.
 \end{align*}
@@
 However, to sample from the full conditional distribution of $\bm{v}$, we use the Metropolis algorithm.
-\textcolor{red}{Since the weights in general are larger in those areas of the frequency domain where the psd has peaks, we specify their starting value proportionally to the periodogram.
+Since the weights in general are larger in those areas of the frequency domain where the psd has peaks, we specify their starting value proportionally to the periodogram.
 Applying the corresponding transformation we obtain the starting value for the vector $\bm{v}$.
-This strategy speeds up the MCMC process by shorting the burn-in period.}
+This strategy speeds up the MCMC process by shortening the burn-in period.
 
@@
 	\begin{align*}
-	\text{Inv-Gamma}\!\left(\alpha_{\tau}\!+\!\nu, {\textstyle\sum\limits_{l=1}^{\nu}} \tfrac{I_n(\lambda_l)}{f(\lambda_l)} + \beta_{\tau} \right) .
+	\text{Inv-Gamma}\!\left(\alpha_{\tau}\!+\!\nu, {\textstyle\sum\limits_{l=1}^{\nu}} \tfrac{I_n(\lambda_l)}{{\pmr s_r(\lambda_l/\pi)}} + \beta_{\tau} \right) .
 	\end{align*}
@@
 We calculate the periodogram and apply a square root transformation in order to have more regular magnitudes and eliminate a potential trend.
-Second, we standardize these values and apply the absolute value function.
-We treat this transformed periodogram as a density function $\widetilde{f}$ and calculate its distribution function $F$.
-Finally, the knots are allocated according to the quantiles of the inverse distribution function of $\widetilde{f}$.
-Thus, in those areas in which the periodogram has sharp and abrupt changes, our procedure will assign more knots in proportion to their magnitudes.
+Second, we standardize these values, apply the absolute value function, {\pmr and normalise them}.
+We treat this transformed periodogram as a {\pmr mass probability} function (pmf) $\widetilde{f}$,{\pmr which is then linearly interpolated in order to define a pmf in a more refined grid.}  Then its distribution function $\widetilde{F}$ is {\pmr approximated by interpolation}.
+Finally, the knots are allocated according to the quantiles of the inverse distribution function $\widetilde{F}^{-1}$, {\pmr which is approximated by interpolation.}  Thus, in those areas in which the periodogram has sharp and abrupt changes, our procedure will assign more knots in proportion to their magnitudes.
 Hereafter, we will refer to this scheme as {\em  quantile-spaced (Q-spaced) knots}.%\textit{periodogram-spaced knots} or short \textit{P-spaced knots}.
@@
 
-The procedure can be summarized as follows:
+%The procedure can be summarized as follows:
+%\begin{enumerate}
+%	\item Let $\bm{x} = \left (\sqrt{I_n(\lambda_1)}, \dots, \sqrt{I_n(\lambda_{\nu})} \right)$ be the square root of the periodogram values for each angular frequency.
+%	\item Define $z_i = \left|\dfrac{x_i - \bar{x}}{S_x} \right|$ where $\bar{x}$ denotes the average and $S_x$ the standard deviation of ${\bm x}$ and calculate its distribution function $F_z = P(z \leq z_i)$.
+%	\item Let $\bm{\xi}^* = \left(\xi_1^*,\dots,\xi_{K^*}^* \right)$ be a vector of $K^*$ quantiles, equally spaced in $[0,1]$.
+%	\item The Q-spaced knot vector is given by $\bm{\xi} = \left(F_{z}^{-1}(\xi^*_1),\dots, F_{z}^{-1}(\xi^*_{K^*}) \right)$.	
+%\end{enumerate}
+
+{\pmr
+The procedure for the frequencies in the unit interval $\omega_l = 2 l/n  \in[0,1]$, which can be easily translated to the frequency domain, can be summarized as follows:	
 \begin{enumerate}
-	\item Let $\bm{x} = \left (\sqrt{I_n(\lambda_1)}, \dots, \sqrt{I_n(\lambda_{\nu})} \right)$ be the square root of the periodogram values for each angular frequency.
-	\item Define $z_i = \left|\dfrac{x_i - \bar{x}}{S_x} \right|$ where $\bar{x}$ denotes the average and $S_x$ the standard deviation of ${\bm x}$ and calculate its distribution function $F_z = P(z \leq z_i)$.
+	\item Let $\bm{x} = \left (\sqrt{I_n(\omega_1)}, \dots, \sqrt{I_n(\omega_{\nu})} \right)$ be the square root of the periodogram values.
+	\item Define $y_l = \left|\dfrac{x_l - \bar{x}}{S_x} \right|$ where $\bar{x}$ denotes the average and $S_x$ the standard deviation of ${\bm x}$.
+	\item Define $z_l = \frac{y_l}{\sum_{l=1}^{\nu} y_l}$, normalised $y_l$ values, and calculate $\widetilde{f}$ by interpolating these values.
+	\item For a large sequence of values $\bm{\varepsilon}\in[0,1]$, calculate $\widetilde{f}(\bm{\varepsilon})$ and normalise them.
+	\item Interpolate the cumulative normalised $\widetilde{f}(\bm{\varepsilon})$ values, which will produce an approximated distribution function $\widetilde{F}$.	
 	\item Let $\bm{\xi}^* = \left(\xi_1^*,\dots,\xi_{K^*}^* \right)$ be a vector of $K^*$ quantiles, equally spaced in $[0,1]$.
-	\item The Q-spaced knot vector is given by $\bm{\xi} = \left(F_{z}^{-1}(\xi^*_1),\dots, F_{z}^{-1}(\xi^*_{K^*}) \right)$.
+	\item The Q-spaced knot vector is given by $\bm{\xi} = \left(\widetilde{F}^{-1}(\xi^*_1),\dots, \widetilde{F}^{-1}(\xi^*_{K^*}) \right)$, where the inverse function is also obtained by interpolation.
+\end{enumerate}
 
-\end{enumerate}
+Note that this knot allocation algorithm is applied only once at the beginning of the MCMC process.  Therefore, its computational cost is negligible.
+
+}
 

### 2020-10-16 responses/responses3/response.tex
Update

@@
 Below, you find a point-by-point response to your points raised.
-We have now carefully revised the paper taking all of your -- and the other referees' -- comments into account and hope that you will now find the paper acceptable for publication.
+We have now carefully revised the paper taking all of your -- and the other referee's -- comments into account and hope that you will now find the paper acceptable for publication.
 For your convenience, all changes made are marked in red in the manuscript.}
@@
 The $\widetilde{f}$ is actually a probability mass function (not a probability density function), which is interpolated.
-We have corrected this in the text.
-We have also changed the description of the algorithm, following now the exact steps implemented in the \texttt{psplinePsd} package.}\bigskip
+However, we have simplified a bit our knot allocation algorithm avoiding that step.
+The new version of the algorithm is described step by step in our manuscript.
+This change leads to small changes in our application section, which has been consequently updated.
+We have also updated \texttt{psplinePsd} package.}\bigskip
 
@@
 Below, you find a point-by-point response to your points raised.
-We have now carefully revised the paper taking all of your -- and the other referees' -- comments into account and hope that you will now find the paper acceptable for publication.
+We have now carefully revised the paper taking all of your -- and the other referee's -- comments into account and hope that you will now find the paper acceptable for publication.
 For your convenience changes are marked in red in the manuscript.
@@
 
-	\response{This could be a very interesting avenue for future research, however, we think it is beyond the scope of this paper.
+	\response{This could be a potential avenue for future research, however, we think it is beyond the scope of this paper.
 An analysis of those characteristics can be quite extensive and could constitute a paper in itself, see for instance \cite{Likhachev2017}.
 We would like to highlight that the number of knots is considered non-important by many, since the penalty parameter plus the penalty order difference penalties specified in the penalty matrix $\textbf{P}$ penalise the inclusion of B-spline densities, controlling the smoothness of the estimate.
-This means that we only need to select a minimum value, the penalty terms take care of the rest.}\bigskip
+This means that we only need to select a minimum value, the penalty terms take care of the rest.
+To quote \cite{Eilers2015}: ``With the number of B-splines in the basis we can tune the smoothness of a curve to
+the data at hand. A smaller number of splines gives a smoother result. However, this is not the only possibility. We can also use a large basis and additionally constrain the
+coefficients of the B-splines, to achieve as much smoothness as desired. A properly
+chosen penalty achieves this"; and also ``The number of B-splines can be (much) larger than the number of observations. The
+penalty makes the fitting procedure well-conditioned. This should be taken literally:
+even a thousand splines will fit ten observations without problems. Such is the power
+of the penalty."}\bigskip
 
@@
 
-	\response{Yes, the loss associated to the approximation given by the beta densities with respect to a loss function cannot be made arbitrarily small.
-This was shown by \cite{Perron:2001} who also showed that if one replaces the beta densities by B-spline densities of fixed order but with variable knots, the loss can be made arbitrarily small by increasing the number of knots.
+	\response{Arbitrarily small means that the approximation error converges to zero with increasing number of knots.
+We have  cited the paper by \cite{Perron:2001} that make this statement precise.
+But we do not mean to imply that this is not the case for the mixture of beta densities.
+Of course, according to the well-known Weierstrass approximation theorem, any continuous function can be approximated in the supnorm by a linear combination of Bernstein polynomials.
+If the continuous function is a density, the approximant becomes a mixture of beta distributions.
+By increasing the number of beta distributions, the approximation error can be made arbitrarily small.
+But as shown in Perron and Mengersen (2001) the rate of approximation is faster for Bsplines than for Bernstein polynomials.
+We have changed this paragraph in the manuscript accordingly.}
+
+%\textcolor{red}{I would change the text in the paper to:
+%" As shown by Perron and Mengersen (2001) for distributions on the interval [0,1], both mixtures of B-spline and beta distributions can approximate these arbitrarily well by increasing the number of mixture components, i.e.\ increasing the polynomial order for the beta distributions and the number of knots for suitable know locations for the B-splines.
+But the rate of approximation is faster for B-splines than for beta distributions.}
+%	\response{Yes, the loss associated to the approximation given by the beta densities with respect to a loss function cannot be made arbitrarily small.
+This was shown by %\cite{Perron:2001} who also showed that if one replaces the beta densities by B-spline densities of fixed order but with variable knots, the loss can be made arbitrarily small by increasing %the number of knots.
 We have added the corresponding reference to our manuscript.} \bigskip

### 2020-10-16 pSplines.tex
Update

@@
 
-As shown by \cite{Perron:2001} for distributions on the interval [0,1], the mixture of B-spline distributions has better approximation properties than the mixture of beta distributions for a fixed number of components.
+{\pmr As shown by \cite{Perron:2001} for distributions on the interval [0,1], both mixtures of B-spline and beta distributions can approximate these arbitrarily well by increasing the number of mixture components, i.e.\ increasing the polynomial order for the beta distributions and the number of knots for suitable know locations for the B-splines.
+But the rate of approximation is faster for B-splines than for beta distributions.}
+%As shown by \cite{Perron:2001} for distributions on the interval [0,1], the mixture of B-spline distributions has better approximation properties than the mixture of beta distributions for a fixed number of components.
 In particular, the approximation error of the mixture of B-spline distributions can be made arbitrarily small by finding  suitable knot locations and increasing the number of knots {\pmr\citep{Perron:2001}}.
-%the mixture of beta distributions is unable to provide a full coverage of the space of cdfs on $[0,1]$.
-They defined a loss function and showed that this cannot be arbitrarily small under this %approach.
-Alternatively, they showed that if the beta distributions are replaced by B-spline distributions of fixed order (shown for order~2) with variable knots, the loss function can be %arbitrarily small by increasing the number of knots.
-The adequacy of the approximation is a function of the number of knots.
 This work motivated \cite{Edwards2019} to propose the BspDP prior on the spectral density function which is based on the representation (\ref{eq:Bernstein}) but with
@@
 We calculate the periodogram and apply a square root transformation in order to have more regular magnitudes and eliminate a potential trend.
-Second, we standardize these values, apply the absolute value function, {\pmr and normalise them}.
-We treat this transformed periodogram as a {\pmr mass probability} function (pmf) $\widetilde{f}$,{\pmr which is then linearly interpolated in order to define a pmf in a more refined grid.}  Then its distribution function $\widetilde{F}$ is {\pmr approximated by interpolation}.
+Second, we standardize these values, apply the absolute value function,{\pmr and normalise them}.
+We treat this transformed periodogram as a {\pmr mass probability} function (pmf) and interpolate its distribution function $\widetilde{F}$.
 Finally, the knots are allocated according to the quantiles of the inverse distribution function $\widetilde{F}^{-1}$, {\pmr which is approximated by interpolation.}  Thus, in those areas in which the periodogram has sharp and abrupt changes, our procedure will assign more knots in proportion to their magnitudes.
@@
 	\item Define $y_l = \left|\dfrac{x_l - \bar{x}}{S_x} \right|$ where $\bar{x}$ denotes the average and $S_x$ the standard deviation of ${\bm x}$.
-	\item Define $z_l = \frac{y_l}{\sum_{l=1}^{\nu} y_l}$, normalised $y_l$ values, and calculate $\widetilde{f}$ by interpolating these values.
-	\item For a large sequence of values $\bm{\varepsilon}\in[0,1]$, calculate $\widetilde{f}(\bm{\varepsilon})$ and normalise them.
-	\item Interpolate the cumulative normalised $\widetilde{f}(\bm{\varepsilon})$ values, which will produce an approximated distribution function $\widetilde{F}$.	
+	\item Define $z_l = \frac{y_l}{\sum_{l=1}^{\nu} y_l}$, and calculate its cumulative distribution $\widetilde{F}$.
+	\item Interpolate $\widetilde{F}$.
 	\item Let $\bm{\xi}^* = \left(\xi_1^*,\dots,\xi_{K^*}^* \right)$ be a vector of $K^*$ quantiles, equally spaced in $[0,1]$.
-	\item The Q-spaced knot vector is given by $\bm{\xi} = \left(\widetilde{F}^{-1}(\xi^*_1),\dots, \widetilde{F}^{-1}(\xi^*_{K^*}) \right)$, where the inverse function is also obtained by interpolation.
+	\item The Q-spaced knot vector is given by $\bm{\xi} = \left(\widetilde{F}^{-1}(\xi^*_1),\dots, \widetilde{F}^{-1}(\xi^*_{K^*}) \right)$. %where the inverse function is obtained by interpolation.	
 \end{enumerate}
@@
 			\textit{Equidistant knots}& & &   \\
-			\hspace{0.5em}P-spline $d=1$ &  0.772 & 0.644 & 0.527 \\
-			\hspace{0.5em}P-spline $d=2$&  0.708 & 0.604 & 0.629 \\
+			\hspace{0.5em}P-spline $d=1$ &  0.769 & 0.655 & 0.527 \\
+			\hspace{0.5em}P-spline $d=2$&  0.698 & 0.609 & 0.629 \\
 			\textit{Q-spaced knots}& & &   \\
-			\hspace{0.5em}P-spline $d=1$ & 0.874 & 0.783 & 0.610 \\
-			\hspace{0.5em}P-spline $d=2$ & 0.874 & 0.815 & 0.640 \\ \hline
+			\hspace{0.5em}P-spline $d=1$ & 0.848 & 0.771 & 0.614 \\
+			\hspace{0.5em}P-spline $d=2$ & 0.920 & 0.840 & 0.648 \\ \hline
 			AR(4)     &  &  &  \\ 
@@
 			\textit{Equidistant knots}& & &   \\
-			\hspace{0.5em}P-spline $d=1$ &  2.914 & 2.389 & 2.307\\
-			\hspace{0.5em}P-spline $d=2$&  3.150 & 2.569 & 2.388\\
+			\hspace{0.5em}P-spline $d=1$ &  2.905 & 2.389 & 2.306\\
+			\hspace{0.5em}P-spline $d=2$&  3.149 & 2.566 & 2.387\\
 			\textit{Q-spaced knots}& & &   \\
-			\hspace{0.5em}P-spline $d=1$ & 2.495 & 1.830 & 1.489\\
-			\hspace{0.5em}P-spline $d=2$ &  2.696 & 2.003 & 1.677\\
+			\hspace{0.5em}P-spline $d=1$ & 2.517 & 1.843 & 1.486\\
+			\hspace{0.5em}P-spline $d=2$ &  2.750 & 1.989 & 1.673\\
 			\bottomrule
@@
 			\textit{Q-spaced knots}& & &   \\
-			\hspace{0.5em}P-spline $d=1$ & 0.997 & 1.000 & 0.997 \\
-			\hspace{0.5em}P-spline $d=2$ & 0.973 & 0.740 & 0.770 \\ \hline		       
+			\hspace{0.5em}P-spline $d=1$ & 1.000 & 1.000 & 0.997 \\
+			\hspace{0.5em}P-spline $d=2$ & 0.933 & 0.527 & 0.713 \\ \hline		       
 			AR(4)     &  &  &  \\ 
@@
 			\textit{Equidistant knots}& & &   \\
-			\hspace{0.5em}P-spline $d=1$ &  0.527 & 0.353 & 0.007 \\
-			\hspace{0.5em}P-spline $d=2$ & 0.323 & 0.167 & 0.007\\
+			\hspace{0.5em}P-spline $d=1$ &  0.547 & 0.343 & 0.010 \\
+			\hspace{0.5em}P-spline $d=2$ & 0.320 & 0.170 & 0.007\\
 			\textit{Q-spaced knots}& & &   \\	
-			\hspace{0.5em}P-spline $d=1$ & 0.790 & 0.847 & 0.813 \\
-			\hspace{0.5em}P-spline $d=2$ & 0.337 & 0.387 & 0.270 \\
+			\hspace{0.5em}P-spline $d=1$ & 0.800 & 0.833 & 0.803 \\
+			\hspace{0.5em}P-spline $d=2$ & 0.313 & 0.363 & 0.277 \\
 			\bottomrule

### 2020-10-27 responses/responses3/response.tex
update

@@
 This change leads to small changes in our application section, which has been consequently updated.
-We have also updated \texttt{psplinePsd} package.}\bigskip
+We have also updated the \texttt{psplinePsd} package.}\bigskip
 
@@
 \begin{enumerate}
-	\item We have specified the prior on $\tau$. 
-	\item We have added $s_r(\omega)$ in Eq(3), short notation used in the posterior distribution of~$\tau$.	
+	\item We have defined $K^*$ as the number of internal knots (B-splines and BspCP prior section).
+	\item We have specified the prior on $\tau$ (P-spline prior section). 
+	\item We have added $s_r(\omega)$ in Eq(3), short notation used in the posterior distribution of~$\tau$ (Posterior computation section).	
+	\item We have simplified a bit our knot location approach.
+This has led to minor variations in the results shown in the application section, which has been updated.
+This change has not impacted the conclusions.
+This version of the knot location algorithm is described step by step in the Quantile-based knot placement section.
 \end{enumerate}

### 2020-10-27 pSplines.tex
update

@@
 
-{\pmr As shown by \cite{Perron:2001} for distributions on the interval [0,1], both mixtures of B-spline and beta distributions can approximate these arbitrarily well by increasing the number of mixture components, i.e.\ increasing the polynomial order for the beta distributions and the number of knots for suitable know locations for the B-splines.
+{\pmr As shown by \cite{Perron:2001} for distributions on the interval [0,1], both mixtures of B-spline and beta distributions can approximate these arbitrarily well by increasing the number of mixture components, i.e.\ increasing the polynomial order for the beta distributions and the number of knots for suitable knot locations for the B-splines.
 But the rate of approximation is faster for B-splines than for beta distributions.}
@@
 \end{align*}
-is the knot vector.  This vector is a non-decreasing sequence that contains $K+r+1$ knots, which can be divided in $2r$ external and $K-r+1$ internal knots.  The latter must be $\geq r$.
+is the knot vector.  This vector is a non-decreasing sequence that contains $K+r+1$ knots, which can be divided in $2r$ external and ${\pmr K^*}=K-r+1$ internal knots.  The latter must be $\geq r$.
 % external and internal knots as in Edwards et al (2018)
@@
 \cite{Eilers2015} consider the use of $K=100$ B-splines a wise choice, unless computational constraints are evident.
-Some algorithms that explore different number of knots $K^*$ in a trial sequence and find an optimal value with respect to a certain goodness-of-fit criterion have been proposed.
+Some algorithms that explore different number of knots in a trial sequence and find an optimal value with respect to a certain goodness-of-fit criterion have been proposed.
 \cite{Ruppert2002} suggested a full-search and myopic algorithms which are based on the generalized cross-validation statistic.
@@
 
-\cite{Ruppert2002} discussed a heuristic rule of thumb $K^* = \min\left\{n/4, 40\right\}$, which is simpler and, in our experience, seems to work very well in general.
+\cite{Ruppert2002} discussed a heuristic rule of thumb $\min\left\{n/4, 40\right\}$, which is simpler and, in our experience, seems to work very well in general.
 This criterion allocates a knot between $\max\left\{ 4, \lfloor n/40 \rfloor \right\}$ observations, where $\lfloor \cdot \rfloor$ stands for the floor function.
@@
 	\item Define $y_l = \left|\dfrac{x_l - \bar{x}}{S_x} \right|$ where $\bar{x}$ denotes the average and $S_x$ the standard deviation of ${\bm x}$.
-	\item Define $z_l = \frac{y_l}{\sum_{l=1}^{\nu} y_l}$, and calculate its cumulative distribution $\widetilde{F}$.
-	\item Interpolate $\widetilde{F}$.
-	\item Let $\bm{\xi}^* = \left(\xi_1^*,\dots,\xi_{K^*}^* \right)$ be a vector of $K^*$ quantiles, equally spaced in $[0,1]$.
-	\item The Q-spaced knot vector is given by $\bm{\xi} = \left(\widetilde{F}^{-1}(\xi^*_1),\dots, \widetilde{F}^{-1}(\xi^*_{K^*}) \right)$. %where the inverse function is obtained by interpolation.	
+	\item Define $z_l = \frac{y_l}{\sum_{l=1}^{\nu} y_l}$, calculate its cumulative distribution $\widetilde{F}$, and interpolate it.
+	\item Let $\bm{q} = \left(q_1,\dots,q_{K^*} \right)$ be a vector of $K^*$ quantiles, equally spaced in $[0,1]$.
+	\item The Q-spaced internal knot vector is given by $\bm{\xi}^* = \left(\widetilde{F}^{-1}(q_1),\dots, \widetilde{F}^{-1}(q_{K^*}) \right)$. %where the inverse function is obtained by interpolation.	
 \end{enumerate}
@@
 We estimated the spectral density functions using the BspDP and P-spline priors.
-The BspDP-based  analysis was performed via the \texttt{gibbs\_bspline} function in the \textsf{R} package \texttt{bsplinePsd} \cite{Edwards:bsplinePsd:2018}.
+The BspDP-based  analysis was performed via the \texttt{gibbs\_bspline} function in the \textsf{R} package \texttt{bsplinePsd{\pmr 0.6.0}} \citep{Edwards:bsplinePsd:2018}.
 The algorithm was executed for 100,000 iterations with a burn-in period of 25,000 and thinning factor of 10, resulting in 7,500 samples used for posterior inference.
@@
 			AR(1)     &  &  &  \\ 
-			B-spline &  0.825 & 0.769 & 0.609  \\
+			B-spline &  0.870 & 0.714 & 0.594  \\
 			\textit{Equidistant knots}& & &   \\
@@
 			AR(4)     &  &  &  \\ 
-			B-spline & 3.019 & 2.250 & 1.832   \\
+			B-spline & 2.990 & 2.202 & 1.800   \\
 			\textit{Equidistant knots}& & &   \\
@@
 			AR(1)     &  &  &  \\ 
-			B-spline &  1.000 & 1.000 & 1.000 \\
+			B-spline &  1.000 & 1.000 & 0.997 \\
 			\textit{Equidistant}& & &   \\
@@
 			AR(4)     &  &  &  \\ 
-			B-spline &  0.933 & 0.967 & 0.897  \\
+			B-spline &  0.907 & 0.970 & 0.897  \\
 			\textit{Equidistant knots}& & &   \\
@@
 Note that the P-spline method based on equidistant knots has similar run-times to the Q-spaced knots.
-The P-spline method is approximately 4-7 times faster than BspDP approach in these examples.
+The P-spline method is approximately 4-6 times faster than BspDP approach in these examples.
 This is due to a reduced number of calculations  per iteration because the B-spline densities are calculated only  once at the start of the algorithm and there are no Dirichlet process approximations via the stick-breaking method.
@@
 			AR(1)     &  &  &  \\ 
-			B-spline &   29.497 & 32.423 & 37.896 \\
+			B-spline &   24.323 & 28.258 & 34.093 \\
 			\textit{Q-spaced knots}& & &   \\
@@
 			\hspace{0.5em}P-spline $d=2$&  4.538 & 6.863 & 8.783\\
-			B-spline/P-spline &  6.278 & 4.531 & 4.122 \\ \hline
+			B-spline/P-spline &  6.090 & 4.368 & 4.493 \\ \hline
 			AR(4)     &  &  & \\ 
-			B-spline & 30.111 & 34.126 & 40.875  \\
+			B-spline & 24.983 & 29.796 & 36.913  \\
 			\textit{Q-spaced knots}& & &   \\
@@
 			\hspace{0.5em}P-spline $d=2$& 4.543 & 6.887 & 8.827 \\
-			B-spline/P-spline & 6.389 & 4.760 & 4.421 \\	
+			B-spline/P-spline & 6.062 & 4.591 & 5.523 \\	
 			\bottomrule

### 2020-10-31 response.tex
Add files via upload

--- 2020-03-25 response.tex
@@
 \usepackage{color}
+\usepackage{hyperref}  % Package for hyperlinking
 %\usepackage{a4wide}
@@
 %\usepackage{texititel}
-\title{Spectral density estimation using P-spline priors}
+\title{Bayesian spectral density estimation using P-splines with quantile-based knot placement}
 \subtitle{Response letter to Statistical Papers}
@@
 Below, you find a point-by-point response to your points raised.
-We have now carefully revised the paper taking all of your -- and the other referees' -- comments into account and hope that you will now find the paper acceptable for publication.
+We have now carefully revised the paper taking all of your -- and the other referee's -- comments into account and hope that you will now find the paper acceptable for publication.
 For your convenience, all changes made are marked in red in the manuscript.}
 
+\textbf{Major comment:}
+
+\refereeQuote{p.8, Quantile-based knot placement\\
+I could not understand the meaning of ``calculate its distribution function $F_z = P(z \leq z_i)$" in the second item of the procedure.
+The values of the transformed periodogram $z_i$ are regarded as the values of the density $\tilde{f}$, which is non-monotonic, taken at equally-spaced frequencies.
+According to item 4, the distribution function $F_z$, and hence the density function $\tilde{f}$, should be defined on the frequency domain.
+\\
+I had a look at the R codes of the package psplinePsd. I think that the density function $\tilde{f}$ is a linear spline that interpolates as
+\begin{align*}
+\tilde{f} \left(\dfrac{\lambda_k}{2\pi}\right) = \dfrac{z_k}{\sum_{i=1}^{\nu}z_i}, \quad k=1,\dots,\nu.
+\end{align*}
+}
+
+\response{That's correct!
+The $\widetilde{f}$ is actually a probability mass function (not a probability density function), which is interpolated.
+However, we have simplified a bit our knot allocation algorithm avoiding that step.
+The new version of the algorithm is described step by step in our manuscript.
+This change leads to small changes in our application section, which has been consequently updated.
+We have also updated the \texttt{psplinePsd} package.}\bigskip
+
+\textbf{Minor comments:}
 \begin{enumerate}
 \item
-\refereeQuote{The authors' terminology is surprising.
-They compute a P-spline fit to a power spectrum, but they call it a P-spline prior.
-The P-spline coefficients are the parameters of the model and they have a prior.}\smallskip
+\refereeQuote{p.4, line 10\\
+A spline of degree $r+1\implies$ A spline of degree $r$ (or, of order $r+1$)}\smallskip
 
-\response{We have edited the manuscript and use {\em P-spline prior} only when referring to the prior but use {\em posterior distribution based on the P-spline prior} when referring to the posterior distribution or Bayesian estimate of the spectral density.
-However, the referee raises a subtle point here in what frequentist and Bayesian regard as unknown parameters.
-In our view, the unknown parameter in this model is the spectral density function.
-As a prior on this spectral density function, we use a mixture of P-spline distributions.
-The mixture weights are hyperparameters with their own hyperprior distribution.}\bigskip
+\response{Fixed.}\bigskip
 
 \item
-\refereeQuote{To get a better fit to sharp peaks in the spectrum, the authors propose an adaptive scheme for knot placement.
-Apparently, they forgot that the differences in the penalty must change too, to correct for the variations in the distance between knots.
-One solution could be to use divided differences, another to use the algorithm of Wood for a derivative-based penalty.}\smallskip
+\refereeQuote{p.5 line 13\\
+In the definition of $\tau, f(\tau \omega)\implies f(\pi \omega)$}\smallskip
 
-\response{Thank you very much for pointing this out. You are absolutely correct, we did use the same penalty for the unequally spaced knots as for the equidistant knots.
-We have rectified this and use derivative-based penalties for the unequally spaced knots.
-We describe the prior specification in paragraph 4  of the section  "adaptive knots" in Chapter 3 .
-As third order derivative-based penalties do not make sense for cubic splines and we want to compare the results for the equally-spaced knots and adaptive knots in the simulation study, we only compare results based on first and second order difference and derivative-based penalties, respectively.
-Therefore we deleted the definition of the third order difference matrix in Chapter 3.
-}\bigskip
-
-\item
-\refereeQuote{A more principle approach is to model the logarithm of the spectrum.
-That can be combined with adaptive smoothing, as described by Wood and Fasiolo, or Rodrigues-Alvarez et al.
-R packages are available.}\smallskip
-
-\response{It is debatable whether putting a prior on the log spectral density is more {\em principled} than putting a prior on the spectral density itself as both terms occur in the log-likelihood function (1) and thus the prior can be put on either scale.
-Our P-spline prior is based on the idea of approximating  a {\em density} function as a mixture of basis distributions on the unit interval.
-Thus, the original scale and not the log scale is appropriate for this particular approach even though for many applications, the log-scale might be the more natural one and we cited many papers that put a prior on the log spectral density, see our paragraph 5 of the introduction.
-This could be an interesting avenue to pursue in future research and see whether putting a P-spline based prior on the log spectral density has advantages over putting a prior on the spectral density.
-We added a paragraph in the Discussion.
-We have now also added the references to the alternative frequentist approaches that you suggested.
-Thank you for pointing these out.
-}\bigskip
-
-\item
-\refereeQuote{The plots of spectra and fit with logarithmic are popular in spectrum analysis, but they are a bit misleading as the fitting is done on the linear scale.}\smallskip
-
-\response{Please see our response to the previous point. Even though we fit a mixture of B-spline densities to the normalized spectral density, it is still possible to show the fit
-on the log scale.
-Large peaks are better visible using the log scale.
-We also want to have a direct comparison of our model fits in Figure 1 and figure 2 to the corresponding ones shown  in Edwards et al.
-(2019) so it makes sense to use the same scale.}\bigskip
-
-\item
-\refereeQuote{I took a short look at the repository. It could be made more valuable by adding the data and some example scripts.}\smallskip
-
-\response{Data and example scripts are included in the R package {\tt psplinePsd}.}\bigskip
-
-\item
-\refereeQuote{Sunspots counts are relatively familiar, but very few people will have seen a time the light curve of S. Carinae. I propose to add a plot of (a part of) the time series.}\smallskip
-
-\response{We have added a time series plot of the first 150 observations of the \textit{S.\ Carinae} variable star light intensities data in Figure 4}.\bigskip
-
-\item
-\refereeQuote{
-The term 'P-spaced knots' is confusing in the context of P-splines, P is already used for penalties.
-An adaptive scheme is proposed that is not exclusive to periodograms.
-Simply calling them adaptive knot looks adequate to me.}\smallskip
-
-\response{ This could indeed be confusing. We have changed 'P-spaced' to 'adaptive' knots.}\bigskip
-
-\item
-\refereeQuote{That there is no consensus on the optimal number of knots for P-splines (page 10) has a simple explanation: it is irrelevant. Ruppert's recommendations have little value.}\smallskip
-
-\response{We see this as a rough guideline only.}\bigskip
-
-\item
-\refereeQuote{I noticed some language problems}\smallskip
-
-\response{Thank you for pointing these out. These have been corrected.}
+\response{That is correct.  Fixed. }\bigskip
 
@@
 
-\newpage
 % -----------------------------------------------------------------------------
@@
 
+\textbf{Specific comments:}
 \begin{enumerate}
-\item
-\refereeQuote{The entire Section 2 is dedicated to the B-spline prior, which were already effectively introduced in Edwards et al.
-(2019).
-I would significantly reduce the length of this section.}\smallskip
+	\item \refereeQuote{ Page 5, how to choose penalty or prior $\phi$?}
 
-\response{We have deleted the detailed description of the BspDP prior but kept the definition of the B-splines as these are required for the definition of the prior based on the  P-splines.}\bigskip
+	\response{In a Bayesian setting, the roughness penalty is translated into the prior distribution for the spline parameters via a normal distribution.
+See for instance \cite{Bremhorst:2016}.
+In order to make the posterior less sensitive to the prior, we suggest to follow their proposal and use their robust specifications for the hierarchical priors on $\phi$, which is also described in our paper.
+}
 
-\item 
-\refereeQuote{The proposed method deals with estimating one single spectral density.
-However, many existing paper already deal with this.
-Can the proposed methodology be extended to multiple or multivariate time series?}\smallskip
+	\item \refereeQuote{To select the number of B-spline densities, the authors employ a heuristic rule of thumb $K^*$ in the application Section, it is of interest to compare it with other criteria such as AIC and BIC.}\smallskip
 
-\response{
- This could be a
-very interesting avenue for future research and we have added a paragraph in the Discussion suggesting a potential approach.
-However, a multivariate extension is outside the scope of this paper.}\bigskip
+	\response{This could be a potential avenue for future research, however, we think it is beyond the scope of this paper.
+An analysis of those characteristics can be quite extensive and could constitute a paper in itself, see for instance \cite{Likhachev2017}.
+We would like to highlight that the number of knots is considered non-important by many, since the penalty parameter plus the penalty order difference penalties specified in the penalty matrix $\textbf{P}$ penalise the inclusion of B-spline densities, controlling the smoothness of the estimate.
+This means that we only need to select a minimum value, the penalty terms take care of the rest.
+To quote \cite{Eilers2015}: ``With the number of B-splines in the basis we can tune the smoothness of a curve to
+the data at hand. A smaller number of splines gives a smoother result. However, this is not the only possibility. We can also use a large basis and additionally constrain the
+coefficients of the B-splines, to achieve as much smoothness as desired. A properly
+chosen penalty achieves this"; and also ``The number of B-splines can be (much) larger than the number of observations. The
+penalty makes the fitting procedure well-conditioned. This should be taken literally:
+even a thousand splines will fit ten observations without problems. Such is the power
+of the penalty."}\bigskip
 
-\item
-\refereeQuote{A Metropolis step is used for v, but a data augmentation step based on Polya-Gamma auxiliary variables will allow for a full Gibbs sampler.
-See "Bayesian Inference for Logistic Models Using PolyaGamma Latent Variables" (2013) of Polson, Scott and Windle.}\smallskip
+	Other minor comments: \bigskip
 
-\response{We had already pointed this out in section "posterior computation" and had cited the paper by Polson et al. (2013).}
+	\item \refereeQuote{Page 2, line 17, what is $\lambda_i$?}\smallskip
 
+	\response{$\lambda_i$ denote the Fourier frequencies. These have now been defined just after the first equation on page 2.} \bigskip
+
+	\item \refereeQuote{ Page 4, lines 5-7, ``In particular, the approximation error of the mixture of B-spline distributions can be made arbitrarily small by finding suitable knot locations and increasing the number of knots", what does ``made arbitrarily small" mean?
+Do you mean the approach with the beta densities the error cannot be made arbitrarily small?}\smallskip
+
+	\response{Arbitrarily small means that the approximation error converges to zero with increasing number of knots.
+We have  cited the paper by \cite{Perron:2001} that make this statement precise.
+But we do not mean to imply that this is not the case for the mixture of beta densities.
+Of course, according to the well-known Weierstrass approximation theorem, any continuous function can be approximated in the supnorm by a linear combination of Bernstein polynomials.
+If the continuous function is a density, the approximant becomes a mixture of beta distributions.
+By increasing the number of beta distributions, the approximation error can be made arbitrarily small.
+But as shown in Perron and Mengersen (2001) the rate of approximation is faster for Bsplines than for Bernstein polynomials.
+We have changed this paragraph in the manuscript accordingly.}
+
+%\textcolor{red}{I would change the text in the paper to:
+%" As shown by Perron and Mengersen (2001) for distributions on the interval [0,1], both mixtures of B-spline and beta distributions can approximate these arbitrarily well by increasing the number of mixture components, i.e.\ increasing the polynomial order for the beta distributions and the number of knots for suitable know locations for the B-splines.
+But the rate of approximation is faster for B-splines than for beta distributions.}
+%	\response{Yes, the loss associated to the approximation given by the beta densities with respect to a loss function cannot be made arbitrarily small.
+This was shown by %\cite{Perron:2001} who also showed that if one replaces the beta densities by B-spline densities of fixed order but with variable knots, the loss can be made arbitrarily small by increasing %the number of knots.
+We have added the corresponding reference to our manuscript.} \bigskip
+
+	\item \refereeQuote{Page 4, line 15, ``A spline of degree r + 1 is a function", I think degree" should be order", we use degree for polynomials.}\smallskip
+
+	\response{Fixed.} \bigskip
+
+	\item \refereeQuote{Page 5, line 20, ``If $\tau = \int_0^1f(\tau \omega)d \omega$ is the normalizing constant", it implies $s_r$ is a density, is it correct?}
+
+	\response{Yes, $s_r$ is a pdf.  Note that it is actually a mixture of B-spline densities.}
 \end{enumerate}
@@
 \subsection*{Further Changes}
+Here some extra changes:
+\begin{enumerate}
+	\item We have defined $K^*$ as the number of internal knots (B-splines and BspDP prior section).
+	\item We have specified the prior on $\tau$ (P-spline prior subsection of section 2). 
+	\item We have added $s_r(\omega)$ in Eq(3), short notation used in the posterior distribution of~$\tau$ (Posterior computation section).	
+	\item We have slightly simplified our knot location approach.
+This has led to minor variations in the results shown in the application section, which has been updated.
+This change has not impacted the conclusions.
+This version of the knot location algorithm is described step by step in the Quantile-based knot placement section.
+\end{enumerate}
 
-We have changed the title from {\em Spectral density estimation using P-spline priors} to {\em Bayesian spectral density estimation using P-spline priors}.\smallskip
+\bibliographystyle{plainnat}
+\bibliography{pSplinesReferences}  % Name of BibTeX file
 
-We have added the keyword {\em Bernstein-Dirichlet process prior}.\smallskip
-
-In the introduction, we have added references to relevant papers published in {\em Statistical Papers} that deal with the penalized spline approach for time series (these are Krivoboka et al (2006) and Wegener and Kauermann (2017)).
+%\begin{thebibliography}{38}
+%	\providecommand{\natexlab}[1]{#1}
+%	\providecommand{\url}[1]{\texttt{#1}}
+%	\expandafter\ifx\csname urlstyle\endcsname\relax
+%	\providecommand{\doi}[1]{doi: #1}\else
+%	\providecommand{\doi}{doi: \begingroup \urlstyle{rm}\Url}\fi
+%	\bibitem[Bremhorst and Lambert(2016)]{Bremhorst:2016}
+%	Vincent Bremhorst and Philippe Lambert.
+%	\newblock Flexible estimation in cure survival models using {B}ayesian
+%	p-splines.
+%	\newblock \emph{Computational Statistics \& Data Analysis}, 93:\penalty0 270 --
+%	284, 2016.
+%	\newblock ISSN 0167-9473.
+%	\newblock \doi{https://doi.org/10.1016/j.csda.2014.05.009}.
+%	\newblock URL
+%	\url{http://www.sciencedirect.com/science/article/pii/S0167947314001492}.
+%\end{thebibliography}
 

### 2020-10-31 pSplines.tex
Add files via upload

@@
 \begin{align*}
-I_n(\lambda) = \dfrac{1}{2 \pi n} \left| \sum_{t=1}^{n} Y_t \exp \left( -i t \lambda\right)\right|^2,
+I_n(\lambda_l) = \dfrac{1}{2 \pi n} \left| \sum_{t=1}^{n} Y_t \exp \left( -i t \lambda_l\right)\right|^2,
 \end{align*}
-in one way or another, where $-\pi < \lambda \leq \pi$ is the angular frequency and $\bm{Y}=(Y_1,\ldots,Y_n)$ is a stationary time series of length $n$.
-{\pmr The periodogram ordinates evaluated at the Fourier frequencies $\lambda \in [-\pi,\pi]$  are asymptotically independent and follow exponential distributions with mean $f(\lambda)$.
-Thus, they are unbiased estimates of $f(\lambda)$  but not consistent \citep{Brockwell:1986}.}
+in one way or another, where $ \lambda_l=2\pi/n$ for $l=0,\ldots,\nu = \left\lfloor (n-1)/2 \right\rfloor$ are the Fourier frequencies and $\bm{Y}=(Y_1,\ldots,Y_n)$ is a stationary time series of length $n$.
+{\pmr The periodogram ordinates evaluated at the Fourier frequencies  are asymptotically independent and follow exponential distributions with mean $f(\lambda_l)$.
+Thus, they are asymptotically  unbiased estimates of $f(\lambda_l)$  but not consistent \citep{Brockwell:1986}.}
 
@@
 L(\bm{Y} | f) &= \prod_{l=1}^{\nu}\dfrac{1}{f(\lambda_{l})} e^{-I_n(\lambda_l)/f(\lambda_l)}  \nonumber \\
-&\propto \exp \left\{ -\! \sum_{l=1}^{\nu} \Bigg( \log f(\lambda_l) + \dfrac{I_{n}(\lambda_l)}{f(\lambda_l)} \Bigg) \right\},\label{eq:whittle_like}
+&\propto \exp \left\{ -\! \sum_{l=1}^{\nu} \Bigg( \log f(\lambda_l) + \dfrac{I_{n}(\lambda_l)}{f(\lambda_l)} \Bigg) \right\}.\label{eq:whittle_like}
 \end{align}
-where $\lambda_l = 2\pi l / n$ are the Fourier frequencies, $f(\cdot)$ spectral density of $\bm{Y}$, and $\nu = \left\lfloor (n-1)/2 \right\rfloor$ is the greatest integer value less than or equal to $(n-1)/2$.
+%where $\lambda_l = 2\pi l / n$ are the Fourier frequencies, $f(\cdot)$ spectral density of $\bm{Y}$, and $\nu = \left\lfloor (n-1)/2 \right\rfloor$ is the %greatest integer value less than or equal to $(n-1)/2$.
  Advantages of this approximation are that it depends directly on the spectral density, unlike the true Gaussian likelihood function, and gives a good approximation for non-Gaussian time series under certain conditions as shown by \cite{ShaoXiaofeng2007ASTf}.
@@
 Second, we standardize these values, apply the absolute value function,{\pmr and normalise them}.
-We treat this transformed periodogram as a {\pmr mass probability} function (pmf) and interpolate its distribution function $\widetilde{F}$.
-Finally, the knots are allocated according to the quantiles of the inverse distribution function $\widetilde{F}^{-1}$, {\pmr which is approximated by interpolation.}  Thus, in those areas in which the periodogram has sharp and abrupt changes, our procedure will assign more knots in proportion to their magnitudes.
+We treat this transformed periodogram as a {\pmr probability mass} function (pmf) and interpolate its distribution function \textcolor{red}{ by the continuous cumulative distribution function} $\widetilde{F}$.
+Finally, the knots are allocated according to quantiles of $\widetilde{F}$.
+ Thus, in those areas in which the periodogram has sharp and abrupt changes, our procedure will assign more knots in proportion to their magnitudes.
 Hereafter, we will refer to this scheme as {\em  quantile-spaced (Q-spaced) knots}.%\textit{periodogram-spaced knots} or short \textit{P-spaced knots}.
@@
 \begin{enumerate}
-	\item Let $\bm{x} = \left (\sqrt{I_n(\omega_1)}, \dots, \sqrt{I_n(\omega_{\nu})} \right)$ be the square root of the periodogram values.
+	\item Let $\bm{x} = \left (\sqrt{I_n(\pi\omega_1)}, \dots, \sqrt{I_n(\pi\omega_{\nu})} \right)$ be the square root of the periodogram values.
 	\item Define $y_l = \left|\dfrac{x_l - \bar{x}}{S_x} \right|$ where $\bar{x}$ denotes the average and $S_x$ the standard deviation of ${\bm x}$.
-	\item Define $z_l = \frac{y_l}{\sum_{l=1}^{\nu} y_l}$, calculate its cumulative distribution $\widetilde{F}$, and interpolate it.
-	\item Let $\bm{q} = \left(q_1,\dots,q_{K^*} \right)$ be a vector of $K^*$ quantiles, equally spaced in $[0,1]$.
-	\item The Q-spaced internal knot vector is given by $\bm{\xi}^* = \left(\widetilde{F}^{-1}(q_1),\dots, \widetilde{F}^{-1}(q_{K^*}) \right)$. %where the inverse function is obtained by interpolation.	
+	\item Define $z_l = \frac{y_l}{\sum_{l=1}^{\nu} y_l}$, $Z_l= \sum_{j\leq l} z_j$, and let $\widetilde{F}$ denote the continuous cdf that interpolates the $Z_l$, $l=1,\ldots,\nu$.
+	\item Let $\bm{q} =\left( \frac{1}{K^*+1}, \frac{2}{K^*+1}\dots, \frac{K^*}{K^*+1} \right)$ be a vector of $K^*$ equally spaced points  in  $[0,1]$.
+	\item The Q-spaced internal knot vector is given by the corresponding quantiles $\bm{\xi}^* = \left(\widetilde{F}^{-1}(q_1),\dots, \widetilde{F}^{-1}(q_{K^*}) \right)$.
+%where the inverse function is obtained by interpolation.
 \end{enumerate}
@@
 \begin{align*}
-\text{IAE} = || \widehat{f} - f ||_1 = \int_{0}^{\pi} |\widehat{f}(\omega)-f(\omega)|\text{d}\omega,
+\text{IAE} = || \widehat{f} - f ||_1 = \int_{0}^{\pi} |\widehat{f}(\textcolor{red}{\lambda})-f(\textcolor{red}{\lambda})|\text{d}\textcolor{red}{\lambda},
 \end{align*}

### 2020-11-04 responses/responses3/response.tex
update

@@
 
-	\response{$\lambda_i$ represents any angular frequency.
-Since we only defined the angular frequency $\lambda$ in that paragraph, we deleted its subscript ``$i$" and added $\lambda\in[-\pi,\pi]$.
-} \bigskip
+	\response{$\lambda_i$ denote the Fourier frequencies. These have now been defined just after the first equation on page 2.} \bigskip
 
@@
 By increasing the number of beta distributions, the approximation error can be made arbitrarily small.
-But as shown in Perron and Mengersen (2001) the rate of approximation is faster for Bsplines than for Bernstein polynomials.
+But as shown in \cite{Perron:2001} the rate of approximation is faster for Bsplines than for Bernstein polynomials.
 We have changed this paragraph in the manuscript accordingly.}
@@
 \begin{enumerate}
-	\item We have defined $K^*$ as the number of internal knots (B-splines and BspCP prior section).
-	\item We have specified the prior on $\tau$ (P-spline prior section). 
+	\item We have defined $K^*$ as the number of internal knots (B-splines and BspDP prior section).
+	\item We have specified the prior on $\tau$ (P-spline prior subsection of section 2). 
 	\item We have added $s_r(\omega)$ in Eq(3), short notation used in the posterior distribution of~$\tau$ (Posterior computation section).	
-	\item We have simplified a bit our knot location approach.
+	\item We have slightly simplified our knot location approach.
 This has led to minor variations in the results shown in the application section, which has been updated.

### 2020-11-04 pSplines.tex
update

@@
 \end{align*}
-in one way or another, where $ \lambda_l=2\pi/n$ for $l=0,\ldots,\nu = \left\lfloor (n-1)/2 \right\rfloor$ are the Fourier frequencies and $\bm{Y}=(Y_1,\ldots,Y_n)$ is a stationary time series of length $n$.
-{\pmr The periodogram ordinates evaluated at the Fourier frequencies  are asymptotically independent and follow exponential distributions with mean $f(\lambda_l)$.
-Thus, they are asymptotically  unbiased estimates of $f(\lambda_l)$  but not consistent \citep{Brockwell:1986}.}
+in one way or another,{\pmr where $ \lambda_l=2\pi l/n$ for $l=0,\ldots,\nu = \left\lfloor (n-1)/2 \right\rfloor$ are the Fourier frequencies} and $\bm{Y}=(Y_1,\ldots,Y_n)$ is a stationary time series of length $n$.
+The periodogram ordinates evaluated at the Fourier frequencies  are asymptotically independent and follow exponential distributions with mean{\pmr$f(\lambda_l)$}.
+Thus, they are asymptotically  unbiased estimates of{\pmr$f(\lambda_l)$}  but not consistent \citep{Brockwell:1986}.
 
@@
 Second, we standardize these values, apply the absolute value function,{\pmr and normalise them}.
-We treat this transformed periodogram as a {\pmr probability mass} function (pmf) and interpolate its distribution function \textcolor{red}{ by the continuous cumulative distribution function} $\widetilde{F}$.
-Finally, the knots are allocated according to quantiles of $\widetilde{F}$.
+We treat this transformed periodogram as a{\pmr probability mass} function (pmf) and interpolate its distribution function \textcolor{red}{ by the continuous cumulative distribution function} $F$.
+Finally, the knots are allocated according to the quantiles of $F$.
  Thus, in those areas in which the periodogram has sharp and abrupt changes, our procedure will assign more knots in proportion to their magnitudes.
@@
 	\item Define $y_l = \left|\dfrac{x_l - \bar{x}}{S_x} \right|$ where $\bar{x}$ denotes the average and $S_x$ the standard deviation of ${\bm x}$.
-	\item Define $z_l = \frac{y_l}{\sum_{l=1}^{\nu} y_l}$, $Z_l= \sum_{j\leq l} z_j$, and let $\widetilde{F}$ denote the continuous cdf that interpolates the $Z_l$, $l=1,\ldots,\nu$.
-	\item Let $\bm{q} =\left( \frac{1}{K^*+1}, \frac{2}{K^*+1}\dots, \frac{K^*}{K^*+1} \right)$ be a vector of $K^*$ equally spaced points  in  $[0,1]$.
-	\item The Q-spaced internal knot vector is given by the corresponding quantiles $\bm{\xi}^* = \left(\widetilde{F}^{-1}(q_1),\dots, \widetilde{F}^{-1}(q_{K^*}) \right)$.
+	\item Define $z_l = \frac{y_l}{\sum_{l=1}^{\nu} y_l}$, $Z_l= \sum_{j\leq l} z_j$, and let $F$ denote the continuous cdf that interpolates the $Z_l$, $l=1,\ldots,\nu$.
+	\item Let $\bm{q} =\left( q_1, \dots, q_{K^*} \right)$ be a vector of $K^*$ equally spaced points  in  $[0,1]$, that is $q_j = \frac{j-1}{K^*-1}$, for $j=1,\dots,K^*$.
+	\item The Q-spaced internal knot vector is given by the corresponding quantiles $\bm{\xi}^* = \left(F^{-1}(q_1),\dots, F^{-1}(q_{K^*}) \right)$.
 %where the inverse function is obtained by interpolation.

### 2020-11-05 responses/responses3/response.tex
update

@@
 The new version of the algorithm is described step by step in our manuscript.
-This change leads to small changes in our application section, which has been consequently updated.
-We have also updated the \texttt{psplinePsd} package.}\bigskip
+This change led to small changes in our application section, which has been consequently updated.
+We have also updated the \texttt{psplinePsd} package on GitHub.
+
+We'd like to clarify that the knot allocation procedure is described in the frequency domain in our manuscript.
+However, the functions in our R implementation are defined in the interval [0,1] for convenience.
+}\bigskip
 

### 2020-11-05 pSplines.tex
update

@@
 {\pmr
-The procedure for the frequencies in the unit interval $\omega_l = 2 l/n  \in[0,1]$, which can be easily translated to the frequency domain, can be summarized as follows:	
+The procedure can be summarized as follows:	%for the frequencies in the unit interval $\omega_l = 2 l/n  \in[0,1]$, which can be easily translated to the frequency domain, can be summarized as follows:
 \begin{enumerate}
-	\item Let $\bm{x} = \left (\sqrt{I_n(\pi\omega_1)}, \dots, \sqrt{I_n(\pi\omega_{\nu})} \right)$ be the square root of the periodogram values.
-	\item Define $y_l = \left|\dfrac{x_l - \bar{x}}{S_x} \right|$ where $\bar{x}$ denotes the average and $S_x$ the standard deviation of ${\bm x}$.
-	\item Define $z_l = \frac{y_l}{\sum_{l=1}^{\nu} y_l}$, $Z_l= \sum_{j\leq l} z_j$, and let $F$ denote the continuous cdf that interpolates the $Z_l$, $l=1,\ldots,\nu$.
+	\item Let $\bm{x} = \left (\sqrt{I_n(\lambda_1)}, \dots, \sqrt{I_n(\lambda_{\nu})} \right)$ be the square root of the periodogram values.
+	\item Define $y_l = \left|\dfrac{x_l - \bar{x}}{S_x} \right|$, for $l=1,\ldots,\nu$, where $\bar{x}$ denotes the average and $S_x$ the standard deviation of ${\bm x}$.
+	\item Define $z_l = \frac{y_l}{\sum_{l=1}^{\nu} y_l}$, $Z_l= \sum_{j\leq l} z_j$, and let $F$ denote the continuous cdf that interpolates the $Z_l$.
 	\item Let $\bm{q} =\left( q_1, \dots, q_{K^*} \right)$ be a vector of $K^*$ equally spaced points  in  $[0,1]$, that is $q_j = \frac{j-1}{K^*-1}$, for $j=1,\dots,K^*$.
